### 2015-08-24 poster/poster.Rnw
add start of poster

\documentclass[a0paper,landscape]{tikzposter}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{acronym}
\usepackage[backend=biber,natbib=true,hyperref=true,style=numeric-comp]{biblatex}
\addbibresource{references.bib}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.  Use showboxes option to show outlines of textboxes.
% \usepackage[absolute]{textpos}
\usepackage[absolute,showboxes]{textpos}

\usepackage{multirow}
\usepackage{array}

\title{Identifying the Tissue of Origin of Extracellular Vesicles Using RNA Expression Signatures}
\author{Don L. Armstrong, Monica Uddin, Derek Wildman}
\usetheme{Basic}

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
#opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
@

\maketitle

\begin{columns}

  \column{0.3}

  \block{Abstract}{ Extracellular vesicles (EVs) are small (40-1000nm)
    compartments that can be found in blood, saliva, and urine. EVs
    originate from the plasma membrane or multivesicular endosome of
    many cell types and contain proteins, lipids, mRNA, miRNA, and
    many other non-coding RNAs. Because EVs contain mRNA from the cell
    of origin and can be obtained without invasive procedures, they
    are a promising remote sensor of the transcriptome of tissues
    which would otherwise be inaccessible. One requirement for
    remotely sensing the transcriptome is identifying the source
    tissue of a specific EV. To identify the source tissue of a
    specific EV, we have identified tissue-specific miRNA and mRNA
    markers for multiple organs and tissues from existing
    publicly-available RNAseq transcriptomes by calculating a tissue
    specificity index. Using machine learning techniques we trained a
    multiclass support vector machine to classify the tissue of
    origin. Our approach identified MIR1914, MIR126, SNORD12, and
    SNORD5 as uterus specific. CGB, LGALS13, and members of the PSG
    family were identified as placenta specific. The accuracy of the
    classifier was verified using RNAseq transcriptome data not in the
    training set. We show the applicability of the classifier to
    simulated EVs by down sampling reads to $10^5$ reads and biasing
    the reads to account for preferential association of RNA motifs
    with EV packaging machinery. When coupled with single-EV
    sequencing, our classifier enables clinicians to monitor
    disease-specific markers over the course of treatment without
    biopsies, increasing the ability of clinicians to adapt treatment
    modalities to the individual patient. }

  \block{Methods}{ Nothing here yet. }

\end{columns}

\end{document}

### 2015-08-26 tissue_specific_expression/R/subset_reads.R
subset reads

@@ -1,0 +1,143 @@
+library(data.table)
+library(reshape2)
+
+args <- c("categorized_samples",
+          "combined_read_counts",
+          "interesting_gene_reads",
+          "interesting_isoform_reads"
+          )
+
+args <- commandArgs(trailingOnly=TRUE)
+
+load(args[1])
+load(args[2])
+
+setkey(categorized.samples,"SRX")
+setnames(gene.counts,"srx","SRX")
+setkey(gene.counts,"SRX")
+setnames(isoform.counts,"srx","SRX")
+setkey(isoform.counts,"SRX")
+
+### combine sample name and srx into the gene counts file
+combined.gene.reads <-
+    categorized.samples[,list(Sample_Group,SRX)][gene.counts]
+
+### do the same for the isoforms
+combined.isoform.reads <-
+    categorized.samples[,list(Sample_Group,SRX)][isoform.counts]
+
+.narm.mean <- function(x){
+    if(length(x)==0) {return(0)} else {return(mean(x,na.rm=TRUE))}
+}
+
+c.gene.reads.wide <-
+    dcast(combined.gene.reads,
+          gene_short_name~Sample_Group,
+          fun.aggregate=.narm.mean,
+          value.var="FPKM"
+          )
+
+c.isoform.reads.wide <-
+    dcast(combined.isoform.reads,
+          tracking_id~Sample_Group,
+          fun.aggregate=.narm.mean,
+          value.var="FPKM")
+
+calculate.entropy <- function(x,bins=10,method="MM"){
+    x <- as.numeric(x)
+    x <- x[is.finite(x)]
+    result <- 0
+    try(result <- entropy(discretize(log(x+1),numBins=bins),method=method),
+        silent=TRUE)
+    return(result)
+}
+
+gene.entropy <-
+    dcast(combined.gene.reads,
+          gene_short_name~.,
+          fun.aggregate=calculate.entropy,
+          value.var="FPKM")
+colnames(gene.entropy)[2] <- "entropy"
+
+gene.max <-
+    dcast(combined.gene.reads,
+          gene_short_name~.,
+          fun.aggregate=function(x){max(x,na.rm=TRUE)},
+          value.var="FPKM")
+colnames(gene.max)[2] <- "max"
+
+isoform.entropy <-
+    dcast(combined.isoform.reads,
+          tracking_id~.,
+          fun.aggregate=calculate.entropy,
+          value.var="FPKM")
[…]

### 2015-08-28 tissue_specific_expression/R/classification_svm.R
add start of SVM classification

@@ -1,0 +1,64 @@
+library(data.table)
+library(caret)
+library(reshape2)
+library(doMC)
+registerDoMC(cores=12)
+
+args <- c("interesting_gene_reads",
+          "svm_classification"
+          )
+
+args <- commandArgs(trailingOnly=TRUE)
+
+load(args[1])
+
+
+gene.fpkm.wide <-
+    dcast(interesting.gene.reads,
+          Sample_Group+SRX~gene_short_name,
+          value.var="FPKM",
+          fun.aggregate=function(x){x[1]})
+
+### for the time being, we're interested in uterus, placenta, or neither.
+gene.fpkm.wide$ut.pla.none <- as.character(gene.fpkm.wide$Sample_Group)
+gene.fpkm.wide$ut.pla.none[!(gene.fpkm.wide$ut.pla.none %in% c("uterus","placenta"))] <-
+    "neither"
+gene.fpkm.wide$ut.pla.none <-
+    factor(gene.fpkm.wide$ut.pla.none)
+
+## set NA to zero
+gene.fpkm.wide[is.na(gene.fpkm.wide)] <- 0
+
+set.seed(10609)
+in_training <-
+    createDataPartition(gene.fpkm.wide$ut.pla.none,
+                        p = .75,
+                        list = FALSE)
+
+genes.training <-
+    gene.fpkm.wide[in_training,]
+
+genes.testing <-
+    gene.fpkm.wide[-in_training,]
+
+knn.train <-
+    train(form=ut.pla.none~.,
+          data=genes.training[,-(1:2)],
+          method="knn",
+          trControl=trainControl(method="cv"),
+          tuneLength=5
+          )
+
+
+svm.train <-
+    train(form=ut.pla.none~.,
+          data=genes.training[,-(1:2)],
+          method="svmRadial",
+          trControl=trainControl(method="cv",number=10)
+          )
+
+
+save(knn.train,
+     svm.train,
+     file=args[length(args)])
+

### 2015-08-28 tissue_specific_expression/R/subset_reads.R
subset reads

@@ -123,12 +123,22 @@ interesting.isoforms <-
                  1]
 setkey(combined.gene.reads,"gene_short_name")
 interesting.gene.reads <-
-    combined.gene.reads[interesting.genes,]
+    combined.gene.reads[interesting.genes,list(Sample_Group,SRX,gene_id,
+                                               gene_short_name,
+                                               FPKM,FPKM_conf_lo,FPKM_conf_hi,
+                                               FPKM_status)]
 
 setkey(combined.isoform.reads,"tracking_id")
 
 interesting.isoform.reads <-
-    combined.isoform.reads[interesting.isoforms,]
+    combined.isoform.reads[interesting.isoforms,
+                           list(Sample_Group,SRX,
+                                tracking_id,
+                                gene_id,
+                                gene_short_name,
+                                FPKM,FPKM_conf_lo,FPKM_conf_hi,
+                                FPKM_status)
+                           ]
 
 save(interesting.gene.reads,
      gene.entropy,

### 2015-08-28 tissue_specific_expression/R/tissue_specific_expression.R

@@ -1,8 +1,8 @@
 library(data.table)
 library(reshape2)
 
-args <- c("categorized_samples",
-          "combined_read_counts",
+args <- c("interesting_gene_reads",
+          "interesting_isoform_reads",
           "tissue_specific_markers")
 
 args <- commandArgs(trailingOnly=TRUE)
@@ -10,42 +10,38 @@ args <- commandArgs(trailingOnly=TRUE)
 load(args[1])
 load(args[2])
 
-setkey(categorized.samples,"SRX")
-setnames(gene.counts,"srx","SRX")
-setkey(gene.counts,"SRX")
-setnames(isoform.counts,"srx","SRX")
-setkey(isoform.counts,"SRX")
-
 ### combine sample name and srx into the gene counts file
-combined.gene.reads <-
-    categorized.samples[,list(Sample_Group,SRX)][gene.counts]
-
-### do the same for the isoforms
-combined.isoform.reads <-
-    categorized.samples[,list(Sample_Group,SRX)][isoform.counts]
-
 .narm.mean <- function(x){
-    mean(x,na.rm=TRUE)
+    if(length(x)==0) {return(0)} else {return(mean(x,na.rm=TRUE))}
 }
 
 c.gene.reads.wide <-
-    dcast(combined.gene.reads,
+    dcast(interesting.gene.reads,
           gene_short_name~Sample_Group,
           fun.aggregate=.narm.mean,
           value.var="FPKM"
           )
 
 c.isoform.reads.wide <-
-    dcast(combined.isoform.reads,
+    dcast(interesting.isoform.reads,
           tracking_id~Sample_Group,
           fun.aggregate=.narm.mean,
           value.var="FPKM")
 
 ## this is the Tissue Specificity Index (eq 1) from Yanai et al.
 tissue.specificity.index <- function(expression){
+    expression <- expression[!is.na(expression)]
+    expression <- expression[is.finite(expression)]
+    if (length(expression)==0) {
+        return(0)
+    }
+    if (max(expression)==0) {
+        return(0)
+    }
     return(sum(1-expression/max(expression))/
                (length(expression)-1))
 }
+
 ##' Calculate genes or isoforms which are specific to a tissue
 ##'
 ##' Identifies genes or isoforms which are specific to certain tissue.
@@ -85,16 +81,7 @@ specific.genes.isoforms <- function(data,min.specificity=0.98,min.max.expression
[…]

### 2015-08-28 poster/poster.Rnw
start loading specific genes into poster

@@
 library("proto")
+library("xtable")
+@
+
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum("../tissue_specific_expression/tissue_specific_markers")>>=
+load("../tissue_specific_expression/tissue_specific_markers")
+tissue.specific.genes <- data.table(tissue.specific.genes)
 @
@@
 
+  \block{Uterus Specific Genes}{
+
+<<uterus.specific.markers.table,results="asis",echo=FALSE>>=
+temp <- tissue.specific.genes[tissue=="uterus",list(gene,specificity)]
+setnames(temp,c("Gene","Tissue Specificity"))
+print(xtable(temp,
+             caption="Uterus specific genes",label="tab:uterus_specific_genes"),
+      floating=FALSE,
+      include.rownames=FALSE,
+      )
+rm(temp)
+@ 
+}
+
+\column{0.3}
+  \block{Placenta Specific Genes}{
+
+<<placenta.specific.markers.table,results="asis",echo=FALSE>>=
+temp <- tissue.specific.genes[tissue=="placenta" & gi.max > 100,list(gene,specificity)]
+setnames(temp,c("Gene","Tissue Specificity"))
+print(xtable(temp,
+             caption="Placenta specific genes",label="tab:placenta_specific_genes"),
+      floating=FALSE,
+      include.rownames=FALSE,
+      )
+rm(temp)
+@ 
+}
+
 \end{columns}

### 2015-08-31 tissue_specific_expression/R/subset_reads.R
ignore genes/isoforms with low variance

@@ -66,6 +66,13 @@ gene.max <-
           value.var="FPKM")
 colnames(gene.max)[2] <- "max"
 
+gene.variance <-
+    dcast(combined.gene.reads,
+          gene_short_name~.,
+          fun.aggregate=function(x){var(x,na.rm=TRUE)},
+          value.var="FPKM")
+colnames(gene.variance)[2] <- "var"
+
 isoform.entropy <-
     dcast(combined.isoform.reads,
           tracking_id~.,
@@ -79,6 +86,13 @@ isoform.max <-
           fun.aggregate=function(x){max(x,na.rm=TRUE)},
           value.var="FPKM")
 colnames(isoform.max)[2] <- "max"
+
+isoform.var <-
+    dcast(combined.isoform.reads,
+          tracking_id~.,
+          fun.aggregate=function(x){var(x,na.rm=TRUE)},
+          value.var="FPKM")
+colnames(isoform.max)[2] <- "var"
 
 
 
@@ -99,6 +113,7 @@ tissue.specificity.index <- function(expression){
 min.entropy <- 1.75
 min.tissue.specificity <- 0.98
 min.expression <- 3
+min.var <- 0.01
 
 gene.tissue.specificity <-
     data.frame(gene_short_name=c.gene.reads.wide[,1],
@@ -113,13 +128,15 @@ isoform.tissue.specificity <-
 interesting.genes <-
     gene.entropy[(gene.entropy[,2] >= min.entropy |
                       gene.tissue.specificity[,2] >= min.tissue.specificity)
-                 & gene.max[,2] >= min.expression,
+                 & gene.max[,2] >= min.expression
+                 & gene.var[,2] >= min.var,
                  1]
 
 interesting.isoforms <-
     isoform.entropy[(isoform.entropy[,2] >= min.entropy |
                       isoform.tissue.specificity[,2] >= min.tissue.specificity)
-                 & isoform.max[,2] >= min.expression,
+                 & isoform.max[,2] >= min.expression
+                 & isoform.var[,2] >= min.var,
                  1]
 setkey(combined.gene.reads,"gene_short_name")
 interesting.gene.reads <-

### 2015-08-31 tissue_specific_expression/R/subset_reads.R
fix gene.variance/gene.var issue (should be gene.var)

@@ -66,7 +66,7 @@ gene.max <-
           value.var="FPKM")
 colnames(gene.max)[2] <- "max"
 
-gene.variance <-
+gene.var <-
     dcast(combined.gene.reads,
           gene_short_name~.,
           fun.aggregate=function(x){var(x,na.rm=TRUE)},

### 2015-08-31 tissue_specific_expression/R/classification_svm.R
set seeds, use repeats in the SVM call
+ Document how to eventually use ROC for multiple-classification problems

@@ -29,6 +29,7 @@ gene.fpkm.wide$ut.pla.none <-
 ## set NA to zero
 gene.fpkm.wide[is.na(gene.fpkm.wide)] <- 0
 
+### seed generated with echo $RANDOM
 set.seed(10609)
 in_training <-
     createDataPartition(gene.fpkm.wide$ut.pla.none,
@@ -41,6 +42,8 @@ genes.training <-
 genes.testing <-
     gene.fpkm.wide[-in_training,]
 
+### seed generated with echo $RANDOM;
+set.seed(30224)
 knn.train <-
     train(form=ut.pla.none~.,
           data=genes.training[,-(1:2)],
@@ -49,12 +52,17 @@ knn.train <-
           tuneLength=5
           )
 
-
+### seed generated with echo $RANDOM;
+set.seed(29224)
+### multi-class ROC https://www.kaggle.com/c/forest-cover-type-prediction/forums/t/10573/r-caret-using-roc-instead-of-accuracy-in-model-training
 svm.train <-
     train(form=ut.pla.none~.,
           data=genes.training[,-(1:2)],
+          preProc=c("center","scale"),
           method="svmRadial",
-          trControl=trainControl(method="cv",number=10)
+          trControl=trainControl(method="cv",number=10,repeats=10,classProbs=TRUE),
+          tuneLength=8,
+          metric="ROC"
           )
 
 

### 2015-08-31 tissue_specific_expression/R/subset_reads.R
fix the other location of gene.variance

@@ -71,7 +71,7 @@ gene.var <-
           gene_short_name~.,
           fun.aggregate=function(x){var(x,na.rm=TRUE)},
           value.var="FPKM")
-colnames(gene.variance)[2] <- "var"
+colnames(gene.var)[2] <- "var"
 
 isoform.entropy <-
     dcast(combined.isoform.reads,

### 2015-08-31 tissue_specific_expression/R/svm_testing.R
add code to test the results of the svm and knn predictions

@@ -1,0 +1,52 @@
+library(data.table)
+library(caret)
+library(reshape2)
+library(doMC)
+registerDoMC(cores=12)
+
+args <- c("svm_classification",
+          "svm_test_results"
+          )
+
+args <- commandArgs(trailingOnly=TRUE)
+
+load(args[1])
+
+
+### test the predictions using SVM
+gene.testing.predictions.svm.prob <-
+    predict(svm.train,
+            newdata=genes.testing,
+            type="prob")
+gene.testing.predictions.svm.prob$prediction <-
+    colnames(gene.testing.predictions.svm.prob)[apply(gene.testing.predictions.svm.prob,1,which.max)]
+
+gene.training.predictions.svm.prob <-
+    predict(svm.train,
+            newdata=genes.training,
+            type="prob")
+gene.training.predictions.svm.prob$prediction <-
+    colnames(gene.training.predictions.svm.prob)[apply(gene.training.predictions.svm.prob,1,which.max)]
+
+### test the predictions using the KNN
+gene.testing.predictions.knn.prob <-
+    predict(knn.train,
+            newdata=genes.testing,
+            type="prob")
+gene.testing.predictions.knn.prob$prediction <-
+    colnames(gene.testing.predictions.knn.prob)[apply(gene.testing.predictions.knn.prob,1,which.max)]
+
+gene.training.predictions.knn.prob <-
+    predict(knn.train,
+            newdata=genes.training,
+            type="prob")
+gene.training.predictions.knn.prob$prediction <-
+    colnames(gene.training.predictions.knn.prob)[apply(gene.training.predictions.knn.prob,1,which.max)]
+
+
+save(gene.testing.predictions.svm.prob,
+     gene.training.predictions.svm.prob,
+     gene.testing.predictions.knn.prob,
+     gene.training.predictions.knn.prob,
+     file=args[length(args)])
+

### 2015-08-31 tissue_specific_expression/R/classification_svm.R
also save the training and testing gene tests

@@ -68,5 +68,7 @@ svm.train <-
 
 save(knn.train,
      svm.train,
+     genes.training,
+     genes.testing,
      file=args[length(args)])
 

### 2015-09-08 tissue_specific_expression/R/classification_svm.R
add code to do multiple tries of svm and knn testing

@@ -4,13 +4,16 @@ library(reshape2)
 library(doMC)
 registerDoMC(cores=12)
 
-args <- c("interesting_gene_reads",
+args <- c("R/svm_training.R",
+          "interesting_gene_reads",
           "svm_classification"
           )
 
 args <- commandArgs(trailingOnly=TRUE)
 
-load(args[1])
+### this file provides the caret.run object
+source(args[1])
+load(args[2])
 
 
 gene.fpkm.wide <-
@@ -19,12 +22,23 @@ gene.fpkm.wide <-
           value.var="FPKM",
           fun.aggregate=function(x){x[1]})
 
+gene.names <- colnames(gene.fpkm.wide)[-(1:2)]
+
 ### for the time being, we're interested in uterus, placenta, or neither.
 gene.fpkm.wide$ut.pla.none <- as.character(gene.fpkm.wide$Sample_Group)
 gene.fpkm.wide$ut.pla.none[!(gene.fpkm.wide$ut.pla.none %in% c("uterus","placenta"))] <-
     "neither"
 gene.fpkm.wide$ut.pla.none <-
     factor(gene.fpkm.wide$ut.pla.none)
+
+for (group in c("uterus","placenta")) {
+    gene.fpkm.wide[[group]] <-
+        as.character(gene.fpkm.wide$Sample_Group)
+    gene.fpkm.wide[[group]][!(gene.fpkm.wide[[group]] %in% group)] <-
+        paste0("not ",group)
+    gene.fpkm.wide[[group]] <-
+        as.factor(gene.fpkm.wide[[group]])
+}
 
 ## set NA to zero
 gene.fpkm.wide[is.na(gene.fpkm.wide)] <- 0
@@ -35,40 +49,34 @@ in_training <-
     createDataPartition(gene.fpkm.wide$ut.pla.none,
                         p = .75,
                         list = FALSE)
-
 genes.training <-
     gene.fpkm.wide[in_training,]
 
 genes.testing <-
     gene.fpkm.wide[-in_training,]
 
-### seed generated with echo $RANDOM;
-set.seed(30224)
-knn.train <-
-    train(form=ut.pla.none~.,
-          data=genes.training[,-(1:2)],
-          method="knn",
-          trControl=trainControl(method="cv"),
-          tuneLength=5
-          )
-
-### seed generated with echo $RANDOM;
-set.seed(29224)
+trained.caret <- list()
[…]

### 2015-09-08 tissue_specific_expression/R/knn_training.R

@@ -1,0 +1,7 @@
+caret.run <-
+    list(run_name="knn",
+         run_method="knn",
+         run_control=trainControl(method="repeatedcv",number=10,repeats=10,classProbs=TRUE),
+         tune_length=8,
+         object_name="trained.caret.knn"
+         )

### 2015-09-08 tissue_specific_expression/R/svm_testing.R

@@ -43,10 +43,14 @@ gene.training.predictions.knn.prob <-
 gene.training.predictions.knn.prob$prediction <-
     colnames(gene.training.predictions.knn.prob)[apply(gene.training.predictions.knn.prob,1,which.max)]
 
+genes.testing.actual <- genes.testing[,c("Sample_Group","ut.pla.none")]
+genes.training.actual <- genes.training[,c("Sample_Group","ut.pla.none")]
 
 save(gene.testing.predictions.svm.prob,
      gene.training.predictions.svm.prob,
      gene.testing.predictions.knn.prob,
      gene.training.predictions.knn.prob,
+     genes.testing.actual,
+     genes.training.actual,
      file=args[length(args)])
 

### 2015-09-08 tissue_specific_expression/R/svm_training.R

@@ -1,0 +1,8 @@
+
+caret.run <-
+    list(run_name="svm",
+         run_method="svmRadial",
+         run_control=trainControl(method="repeatedcv",number=10,repeats=10,classProbs=TRUE),
+         tune_length=8,
+         object_name="trained.caret.knn"
+         )

### 2015-09-08 tissue_specific_expression/R/wide_reads.R

@@ -1,0 +1,47 @@
+library(data.table)
+library(reshape2)
+
+args <- c("interesting_gene_reads",
+          "interesting_gene_reads_wide"
+          )
+
+args <- commandArgs(trailingOnly=TRUE)
+
+load(args[1])
+
+is.genes <- TRUE
+if (any(grepl("isoform",args[length(args)]))) {
+    is.genes <- FALSE
+}
+
+if(is.genes) {
+    combined.reads <- interesting.gene.reads
+} else {
+    combined.reads <- interesting.isoform.reads
+}
+    
+grouping.by.sample.formula <-
+    as.formula(gene_short_name+tracking_id~Sample_Group)
+grouping.formula <-
+    as.formula(tracking_id~.)
+
+.narm.mean <- function(x){
+    if(length(x)==0) {return(0)} else {return(mean(x,na.rm=TRUE))}
+}
+
+c.reads.wide <-
+    dcast(combined.reads,
+          grouping.by.sample.formula,
+          fun.aggregate=.narm.mean,
+          value.var="FPKM"
+          )
+
+if (is.genes) {
+    interesting.gene.reads.wide <- data.table(c.reads.wide)
+    save(interesting.gene.reads.wide,
+         file=args[length(args)])
+} else {
+    interesting.isoform.reads.wide <- data.table(c.reads.wide)
+    save(interesting.isoform.reads.wide,
+         file=args[length(args)])
+}

### 2015-09-08 tissue_specific_expression/R/subset_reads.R
do the interesting reads wide; combine star and isoform logs

@@ -3,45 +3,59 @@ library(reshape2)
 
 args <- c("categorized_samples",
           "combined_read_counts",
-          "interesting_gene_reads",
-          "interesting_isoform_reads"
+          "additional_interesting_genes.txt",
+          "interesting_gene_reads"
           )
 
 args <- commandArgs(trailingOnly=TRUE)
 
 load(args[1])
 load(args[2])
+additional.genes <- fread(args[3])
+
+is.genes <- TRUE
+if (any(grepl("isoform",args[length(args)]))) {
+    is.genes <- FALSE
+}
 
 setkey(categorized.samples,"SRX")
-setnames(gene.counts,"srx","SRX")
-setkey(gene.counts,"SRX")
-setnames(isoform.counts,"srx","SRX")
-setkey(isoform.counts,"SRX")
+
+
+## we're using tracking_id for genes instead of gene_short_name
+## because SNORD5 and some other genes have multiple gene ids but also
+## share a common name
+if (is.genes) {
+    read.counts <- gene.counts
+} else {
+    read.counts <- isoform.counts
+}
+grouping.by.sample.formula <-
+    as.formula(tracking_id~Sample_Group)
+grouping.formula <-
+    as.formula(tracking_id~.)
+
+additional.tracking.ids <-
+    read.counts[!duplicated(tracking_id),
+                list(tracking_id,gene_short_name)][gene_short_name %in%
+                                                       additional.genes[[1]]][[1]]
+
+setnames(read.counts,"srx","SRX")
+setkey(read.counts,"SRX")
 
 ### combine sample name and srx into the gene counts file
-combined.gene.reads <-
-    categorized.samples[,list(Sample_Group,SRX)][gene.counts]
-
-### do the same for the isoforms
-combined.isoform.reads <-
-    categorized.samples[,list(Sample_Group,SRX)][isoform.counts]
+combined.reads <-
+    categorized.samples[,list(Sample_Group,SRX)][read.counts]
 
 .narm.mean <- function(x){
     if(length(x)==0) {return(0)} else {return(mean(x,na.rm=TRUE))}
 }
 
-c.gene.reads.wide <-
-    dcast(combined.gene.reads,
-          gene_short_name~Sample_Group,
+c.reads.wide <-
+    dcast(combined.reads,
+          grouping.by.sample.formula,
           fun.aggregate=.narm.mean,
           value.var="FPKM"
           )
-
-c.isoform.reads.wide <-
[…]

### 2015-09-08 manuscript/manuscript.Rnw
add manuscript

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{txfonts}
\usepackage{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=numeric-comp]{biblatex}
\addbibresource{references.bib}
%\usepackage[noblocks]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\pagestyle{fancy}
\author{AUTHOR}
\title{TITLE}
\date{\ }
\onehalfspacing

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
#opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("xtable")
@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum("../tissue_specific_expression/tissue_specific_markers")>>=
load("../tissue_specific_expression/tissue_specific_markers")
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")
@

\section{Tissue Specific Markers}

\subsection{Uterus Specific}
<<uterus.specific.markers.table,results="asis",echo=FALSE>>=
temp <- tissue.specific.genes[tissue=="uterus",
                              list(gene,tissue,specificity,gi.max)]
setorder(temp,-gi.max)
setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
print(xtable(temp,
             caption="Uterus specific genes",label="tab:uterus_specific_genes"),
      include.rownames=FALSE,
      )
rm(temp)
@

FTX_1 is very interseting

PCP4 is troubling, because this is a purkingie cell protein

Some of the markers which we had previously seen (and in fact mention
in our abstract have gone away).

We originally identified MIR1914, MIR126, SNORD12, and SNORD5 as
uterus specific. MIR1914 is actually placenta specific, as is MIR126.

This occured because the placenta FPKMs we based the original list on
did not properly align against miRNA or snRNA
\fref{tab:missing_uterus_specific_genes}. The uterus calls also appear
to be wrong or cooked; for example, SNORD5 has counts of 7000ish which
are not inside of its confidence interval. This may be indicative of a
off-by-one error or something else that is screwed up in their file.

<<missing.uterus.specific.markers.table,results="asis",echo=FALSE>>=
temp <- tissue.specific.genes[c("MIR1914","MIR126","SNORD12","SNORD5"),
                              list(gene,tissue,specificity,gi.max)]
setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
print(xtable(temp,
             caption="Missing previously uterus specific genes",
             label="tab:missing_uterus_specific_genes"),
      include.rownames=FALSE,
      )
rm(temp)
@

\subsection{Placenta Specific Genes}

<<placenta.specific.markers.table,results="asis",echo=FALSE>>=
temp <- tissue.specific.genes[tissue=="placenta" & gi.max > 100,
                              list(gene,tissue,specificity,gi.max)]
setorder(temp,-gi.max)
setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
print(xtable(temp,
             caption="Placenta specific genes",label="tab:placenta_specific_genes"),
      include.rownames=FALSE,
      floating=FALSE,
      tabular.environment="longtable"
      )
rm(temp)
@

\end{document}

### 2015-09-08 poster/poster.Rnw
start working on the introduction section by adapting the abstract

@@
 
-  \block{Abstract}{ Extracellular vesicles (EVs) are small (40-1000nm)
-    compartments that can be found in blood, saliva, and urine. EVs
-    originate from the plasma membrane or multivesicular endosome of
-    many cell types and contain proteins, lipids, mRNA, miRNA, and
-    many other non-coding RNAs. Because EVs contain mRNA from the cell
-    of origin and can be obtained without invasive procedures, they
-    are a promising remote sensor of the transcriptome of tissues
-    which would otherwise be inaccessible. One requirement for
-    remotely sensing the transcriptome is identifying the source
-    tissue of a specific EV. To identify the source tissue of a
-    specific EV, we have identified tissue-specific miRNA and mRNA
-    markers for multiple organs and tissues from existing
-    publicly-available RNAseq transcriptomes by calculating a tissue
-    specificity index. Using machine learning techniques we trained a
-    multiclass support vector machine to classify the tissue of
-    origin. Our approach identified MIR1914, MIR126, SNORD12, and
-    SNORD5 as uterus specific. CGB, LGALS13, and members of the PSG
-    family were identified as placenta specific. The accuracy of the
-    classifier was verified using RNAseq transcriptome data not in the
-    training set. We show the applicability of the classifier to
-    simulated EVs by down sampling reads to $10^5$ reads and biasing
-    the reads to account for preferential association of RNA motifs
-    with EV packaging machinery. When coupled with single-EV
-    sequencing, our classifier enables clinicians to monitor
-    disease-specific markers over the course of treatment without
-    biopsies, increasing the ability of clinicians to adapt treatment
-    modalities to the individual patient. }
+  \block{Introduction}{\acp{EV} are small (40-1000nm) compartments
+    that can be found in blood, saliva, and urine. \acp{EV} originate
+    from the plasma membrane or multivesicular endosome of many cell
+    types and contain proteins, lipids, mRNA, miRNA, and many other
+    non-coding RNAs. Because \acp{EV} contain mRNA from the cell of
+    origin and can be obtained without invasive procedures, they are a
+    promising remote sensor of the transcriptome of tissues which
+    would otherwise be inaccessible.
+
+    One requirement for remotely sensing the transcriptome is
+    identifying the source tissue of a specific \ac{EV}. To identify
+    the source tissue of a specific \ac{EV}, we have identified
+    tissue-specific miRNA and mRNA markers for multiple organs and
+    tissues from existing publicly-available RNAseq transcriptomes by
+    calculating a tissue specificity index. Using machine learning
+    techniques we trained a multiclass support vector machine to
+    classify the tissue of origin.
+
+%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
+%     uterus specific. CGB, LGALS13, and members of the PSG family were
+%     identified as placenta specific. The accuracy of the classifier
+%     was verified using RNAseq transcriptome data not in the training
+%     set. We show the applicability of the classifier to simulated EVs
+%     by down sampling reads to $10^5$ reads and biasing the reads to
+%     account for preferential association of RNA motifs with EV
+%     packaging machinery. When coupled with single-EV sequencing, our
+%     classifier enables clinicians to monitor disease-specific markers
+%     over the course of treatment without biopsies, increasing the
+%     ability of clinicians to adapt treatment modalities to the
+%     individual patient. 
+  }
 

### 2015-09-08 poster/poster.Rnw
start making UofI color scheme
add uterus and placenta specific genes as heatmaps

@@
 \author{Don L. Armstrong, Monica Uddin, Derek Wildman}
-\usetheme{Basic}
+\usetheme{Default}
+
+% use this
+% http://www.identitystandards.illinois.edu/graphicstandardsmanual/othermedia/posterpresentation.html
+\definecolorstyle{UofI}{
+  \definecolor{boldBlue}{HTML}{002058}
+  \definecolor{boldOrange}{HTML}{E87722}
+  \definecolor{professionalBlue}{HTML}{606EB2}
+  \definecolor{professionalOrange}{HTML}{D45D00}
+  }{
+    \colorlet{backgroundcolor}{professionalBlue}
+    \colorlet{framecolor}{black}
+    \colorlet{titlefgcolor}{white}
+    \colorlet{titlebgcolor}{professionalOrange}
+    \colorlet{blocktitlebgcolor}{professionalOrange}
+    \colorlet{blocktitlefgcolor}{white}
+    }
+\usecolorstyle{UofI}
 
@@
 <<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
-opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
+opts_chunk$set(dev="cairo_pdf",out.width="0.20\\columnwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
 #opts_chunk$set(cache=TRUE, autodep=TRUE)
@@
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum("../tissue_specific_expression/tissue_specific_markers")>>=
-load("../tissue_specific_expression/tissue_specific_markers")
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide")))>>=
+for (a in paste0("../tissue_specific_expression/",
+                 c("tissue_specific_markers",
+                   "interesting_gene_reads_wide"))) {
+
+    load(a)
+}
 tissue.specific.genes <- data.table(tissue.specific.genes)
@@
 
+<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
+specific.markers.plot <- function(tissue.name,genes) {
+    setkey(interesting.gene.reads.wide,"gene_short_name")
+    if (missing(genes))
+        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
+    interesting.tissues <- 
+        colnames(interesting.gene.reads.wide)[
+           which(apply(apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
+                             function(x){order(-x)==2}),1,any))+2]
+     interesting.tissues <- 
+        c("gene_short_name",tissue.name,interesting.tissues)
+
+    tissue.specific.gene.reads <- 
+        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
+                        id="gene_short_name",
+                        variable.name="tissue",
+                        value.name="FPKM"
+                        ))
+    tissue.specific.gene.reads[,log2.FPKM:=log2(FPKM+1)]
+    return(ggplot(tissue.specific.gene.reads,
+                  aes(y=gene_short_name,x=tissue))
+           + geom_tile(aes(fill=log2.FPKM))
+           + scale_fill_gradient(low="white",high="#002058")
+           + scale_x_discrete("", expand = c(0, 0))
+           + scale_y_discrete("", expand = c(0, 0))
+           + theme_grey(base_size=14)
+           + theme(legend.position = "none",
+                   axis.ticks = element_blank(), 
+                   axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
+           )
+}
+@
+
+\acro{EV}{Extracellular Vesicles}
+
 \begin{columns}
 
-  \column{0.3}
+  \column{0.25}
 
@@
 
+\column{0.25}
   \block{Uterus Specific Genes}{
 
-<<uterus.specific.markers.table,results="asis",echo=FALSE>>=
-temp <- tissue.specific.genes[tissue=="uterus",list(gene,specificity)]
-setnames(temp,c("Gene","Tissue Specificity"))
-print(xtable(temp,
-             caption="Uterus specific genes",label="tab:uterus_specific_genes"),
-      floating=FALSE,
-      include.rownames=FALSE,
-      )
-rm(temp)
+{\centering
+% <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
+% temp <- tissue.specific.genes[tissue=="uterus",list(gene,specificity,gi.max)]
+% setorder(temp,-gi.max)
+% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
+% print(xtable(temp,
+%              caption="Uterus specific genes",label="tab:uterus_specific_genes"),
+%       floating=FALSE,
+%       include.rownames=FALSE,
+%       )
+% rm(temp)
+% @ 
+% }
+
+<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+print(specific.markers.plot("uterus")
+      + xlab("Uterus Specific Genes")
+      + ylab("Tissue"))
 @
 }
+\captionof{figure}{Uterus specific genes. $\log_2 \mathrm{FPKM}$ of
+  genes with tissue specific indices $\ge 0.98$ which are specific to
+  uterus tissue in uterus and tissues with second highest expression
+  for those genes. With the exception of FTX\_1, which is expressed in
+  multiple fetal tissues, all genes are highly specific to uterus.}
 
-\column{0.3}
+}
+
   \block{Placenta Specific Genes}{
+{\centering
+% <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
+% temp <- tissue.specific.genes[tissue=="placenta" & gi.max >= 1000,
+%                               list(gene,specificity,gi.max)]
+% setorder(temp,-gi.max)
+% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
+% print(xtable(temp,
+%              caption="Placenta specific genes",label="tab:placenta_specific_genes"),
+%       floating=FALSE,
+%       include.rownames=FALSE,
+%       )
+% rm(temp)
+% @
 
-<<placenta.specific.markers.table,results="asis",echo=FALSE>>=
-temp <- tissue.specific.genes[tissue=="placenta" & gi.max > 100,list(gene,specificity)]
-setnames(temp,c("Gene","Tissue Specificity"))
-print(xtable(temp,
-             caption="Placenta specific genes",label="tab:placenta_specific_genes"),
-      floating=FALSE,
-      include.rownames=FALSE,
-      )
-rm(temp)
+<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+print(specific.markers.plot("placenta",
+                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
+                                                                         gi.max >= 1024
+                               ,gene]))
+      + xlab("Uterus Specific Genes")
+      + ylab("Tissue"))
 @
+
+}
+\captionof{figure}{Placenta specific genes. $\log_2 \mathrm{FPKM}$ of
+  genes with tissue specific indices $\ge 0.98$ and
+  $\mathrm{FPKM} \ge 1024$ which are specific to uterus tissue in
+  placenta and tissues with second highest expression for those
+  genes.}
+}
+
+\column{0.25}
+
+\block{SVM to distinguish tissues}{
+  % Describe SVM approach
+
+  % Figure showing the samples
+
+  % Ability to identify placenta
+
+  % Ability to identify uterus
+
+  % ROC curve for uterus/not uterus
+}
+
+\column{0.25}
+
+\block{Identification of tissue in reduced reads}{
+  % Describe reduced read approach
+
+  % 
+}
+
+\block{Conclusions}{
+}
+
+\block{References and Acknowledgements}{ The code and datasets for
+  this poster can be found at
+  \url{http://github.com/uiuc-cgm/exosome_markers_per_med_sep_2015.git}
 }

### 2015-09-09 tissue_expression/R/combine_star_logs.R
remove trailing only line

@@ -28,7 +28,7 @@ for (file in star.log.files) {
     star.log <- read.table(file,sep="|",fill=TRUE,stringsAsFactors=FALSE)
     colnames(star.log) <- c("field","value")
     star.log$value <- gsub("\\t","",star.log$value)
-    star.log$field <- gsub("^\\s+","",star.log$field)
+    star.log$field <- gsub("(^\\s+|\\s+$)","",star.log$field)
     star.log$srx <- srx.accession
     star.logs[[srx.accession]] <- data.table(star.log)[!grepl(":",field),]
     i <- i + 1

### 2015-09-09 tissue_expression/R/combine_fpkms.R
add missing 57 epigenomes file reads

@@ -21,6 +21,10 @@ star.log.files <- grep("_star/Log.final.out",args,value=TRUE)
 
 gtex.file <- grep("GTEx_Analysis",args,value=TRUE)
 
+epigenomes.files <- grep("57epigenomes.RPKM",args,value=TRUE)
+epigenomes.samples.file <- grep("EG.name.txt",args,value=TRUE)
+epigenomes.annotation.file <- grep("Ensembl_v65.Gencode_v10.ENSG.gene_info",args,value=TRUE)
+
 .get.srx <- function(x){
     gsub("(SRX\\d+)/.+","\\1",x)
 }
@@ -51,6 +55,37 @@ gtex.counts <-
          value.name="FPKM",
          variable.name="srx")
 gene.counts[["gtex"]] <- gtex.counts
+
+### deal with 57epigenomes data
+epigenomes.names <- fread(epigenomes.samples.file,header=FALSE)
+setnames(epigenomes.names,c("GEO_Accession","Sample_Name"))
+
+epigenomes.annotation <- fread(epigenomes.annotation.file,header=FALSE)
+setnames(epigenomes.annotation,
+         c("gene_id","chr","start","stop",
+           "strand","type",
+           "gene_short_name","gene_full_name"))
+setkey(epigenomes.annotation,"gene_id")
+
+epigenomes.counts <- list()
+for (file in epigenomes.files) {
+    epigenomes.counts[[file]] <-
+        fread(paste0("zcat ",file),skip=0)
+}
+epigenomes.counts <- rbindlist(epigenomes.counts)
+setnames(epigenomes.counts,epigenomes.names[,GEO_Accession],epigenomes.names[,Sample_Name])
+
+setkey(epigenomes.counts,"gene_id")
+epigenomes.counts <-
+    epigenomes.annotation[,list(gene_id,gene_short_name)][epigenomes.counts]
+
+epigenomes.counts <-
+    melt(epigenomes.counts,
+         id.vars=c("gene_id","gene_short_name"),
+         value.name="FPKM",
+         variable.name="srx")
+gene.counts[["57epigenomes"]] <- epigenomes.counts
+
 
 gene.counts <- rbindlist(gene.counts,fill=TRUE)
 

### 2015-09-09 tissue_expression/R/combine_fpkms.R
do not set the sample name to the srx for 57 epigenomes files

@@ -73,7 +73,6 @@ for (file in epigenomes.files) {
         fread(paste0("zcat ",file),skip=0)
 }
 epigenomes.counts <- rbindlist(epigenomes.counts)
-setnames(epigenomes.counts,epigenomes.names[,GEO_Accession],epigenomes.names[,Sample_Name])
 
 setkey(epigenomes.counts,"gene_id")
 epigenomes.counts <-

### 2015-09-10 tissue_expression/R/combine_fpkms.R
set tracking_id to gene_id for 57 epigenomes datasets

@@ -83,6 +83,7 @@ epigenomes.counts <-
          id.vars=c("gene_id","gene_short_name"),
          value.name="FPKM",
          variable.name="srx")
+epigenomes.counts[,tracking_id:=gene_id]
 gene.counts[["57epigenomes"]] <- epigenomes.counts
 
 

### 2015-09-10 tissue_expression/R/combine_fpkms.R
also set tracking_id for gtex counts

@@ -54,6 +54,7 @@ gtex.counts <-
          id.vars=c("gene_id","gene_short_name"),
          value.name="FPKM",
          variable.name="srx")
+gtex.counts[,tracking_id:=gene_id]
 gene.counts[["gtex"]] <- gtex.counts
 
 ### deal with 57epigenomes data

### 2015-09-10 tissue_specific_expression/R/subset_reads.R
properly calculate entropy; raise minimum entropy to 1.9

@@ -1,5 +1,6 @@
 library(data.table)
 library(reshape2)
+library(entropy)
 
 args <- c("categorized_samples",
           "combined_read_counts",
@@ -72,6 +73,9 @@ grouping.entropy <-
           fun.aggregate=calculate.entropy,
           value.var="FPKM")
 colnames(grouping.entropy)[2] <- "entropy"
+grouping.entropy <- data.table(grouping.entropy)
+setkey(grouping.entropy,"tracking_id")
+    
 
 grouping.max <-
     dcast(combined.reads,
@@ -79,6 +83,8 @@ grouping.max <-
           fun.aggregate=function(x){max(x,na.rm=TRUE)},
           value.var="FPKM")
 colnames(grouping.max)[2] <- "max"
+grouping.max <- data.table(grouping.max)
+setkey(grouping.max,"tracking_id")
 
 grouping.var <-
     dcast(combined.reads,
@@ -86,6 +92,8 @@ grouping.var <-
           fun.aggregate=function(x){var(x,na.rm=TRUE)},
           value.var="FPKM")
 colnames(grouping.var)[2] <- "var"
+grouping.var <- data.table(grouping.var)
+setkey(grouping.var,"tracking_id")
 
 ## this is the Tissue Specificity Index (eq 1) from Yanai et al.
 tissue.specificity.index <- function(expression){
@@ -101,23 +109,27 @@ tissue.specificity.index <- function(expression){
                (length(expression)-1))
 }
 
-min.entropy <- 1.75
+min.entropy <- 1.9
 min.tissue.specificity <- 0.98
-min.expression <- 3
+min.expression <- 10
 min.var <- 0.01
 
 tissue.specificity <-
-    data.frame(grouping=c.reads.wide[,1],
-               tissue.specificity.index=apply(c.reads.wide[,-1],1,
-                   tissue.specificity.index))
+    data.table(data.frame(tracking_id=c.reads.wide[,1],
+                          tissue.specificity.index=apply(c.reads.wide[,-1],1,
+                              tissue.specificity.index)))
+setkey(tissue.specificity,"tracking_id")
+
+grouping.variables <-
+    grouping.entropy[grouping.var][grouping.max][tissue.specificity]
 
 interesting.groups <-
-    grouping.entropy[((grouping.entropy[,2] >= min.entropy |
-                          tissue.specificity[,2] >= min.tissue.specificity)
-                      & grouping.max[,2] >= min.expression
[…]

### 2015-09-11 tissue_specific_expression/R/categorize_samples.R
include EG samples abstract out wide reads and subset reads

@@ -2,20 +2,27 @@ library(data.table)
 
 args <- c("chosen_samples",
           "GTEx_Data_V4_Annotations_SampleAttributesDS.txt",
+          "EG.name.txt",
           "categorized_samples")
 
 args <- commandArgs(trailingOnly=TRUE)
 
 load(args[1])
 gtex.samples <- fread(args[2])
+eg.samples <- fread(args[3],header=FALSE)
 
 categorized.samples <- copy(chosen.samples[,list(Sample_Name,SRX)])
 
 gtex.samples <- gtex.samples[,list(SMTSD,SAMPID)]
 setnames(gtex.samples,c("SAMPID","SMTSD"),c("SRX","Sample_Name"))
+setnames(eg.samples,c("SRX","Sample_Name"))
 categorized.samples <-
     rbind(categorized.samples,
-          gtex.samples)
+          gtex.samples,
+          eg.samples
+          )
+
+
 
 categorized.samples[,Sample_Group:=Sample_Name]
 categories <-
@@ -54,7 +61,13 @@ categories <-
                  replace="uterus"),
         placenta=
             list(match="^[AESX]\\d+_[abc]\\d?(?:\\.\\d+)?$",
-                 replace="placenta")
+                 replace="placenta"),
+        remove.primary.cells.skin=
+            list(match="(?i)_primary_cells_skin\\d+$",
+                 replace=""),
+        remove.understore=
+            list(match="[_ ]+",
+                 replace=" ")
         )
 for (category in names(categories)) {
     categorized.samples[grepl(categories[[category]]$match,Sample_Group),

### 2015-09-11 tissue_specific_expression/R/subset_reads.R

@@ -108,7 +108,7 @@ tissue.specificity.index <- function(expression){
     return(sum(1-expression/max(expression))/
                (length(expression)-1))
 }
-
+R
 min.entropy <- 1.9
 min.tissue.specificity <- 0.98
 min.expression <- 10

### 2015-09-11 tissue_specific_expression/R/tissue_specific_expression.R

@@ -1,8 +1,8 @@
 library(data.table)
 library(reshape2)
 
-args <- c("interesting_gene_reads",
-          "interesting_isoform_reads",
+args <- c("interesting_gene_reads_wide",
+          "interesting_isoform_reads_wide",
           "tissue_specific_markers")
 
 args <- commandArgs(trailingOnly=TRUE)
@@ -14,20 +14,6 @@ load(args[2])
 .narm.mean <- function(x){
     if(length(x)==0) {return(0)} else {return(mean(x,na.rm=TRUE))}
 }
-
-c.gene.reads.wide <-
-    dcast(interesting.gene.reads,
-          gene_short_name~Sample_Group,
-          fun.aggregate=.narm.mean,
-          value.var="FPKM"
-          )
-
-c.isoform.reads.wide <-
-    dcast(interesting.isoform.reads,
-          tracking_id~Sample_Group,
-          fun.aggregate=.narm.mean,
-          value.var="FPKM")
-
 ## this is the Tissue Specificity Index (eq 1) from Yanai et al.
 tissue.specificity.index <- function(expression){
     expression <- expression[!is.na(expression)]
@@ -78,8 +64,8 @@ specific.genes.isoforms <- function(data,min.specificity=0.98,min.max.expression
     return(tissue.specific.gi)
 }    
 
-tissue.specific.genes <- specific.genes.isoforms(c.gene.reads.wide)
-tissue.specific.isoforms <- specific.genes.isoforms(c.isoform.reads.wide)
+tissue.specific.genes <- specific.genes.isoforms(interesting.gene.reads.wide)
+tissue.specific.isoforms <- specific.genes.isoforms(interesting.isoform.reads.wide)
 
 save(tissue.specific.genes,
      tissue.specific.isoforms,

### 2015-09-11 poster/poster.Rnw
add methods section about reads include references for methods

@@
 \usepackage{booktabs}
-\usepackage{multirow}
+% \usepackage{multirow}
 \usepackage{acronym}
-\usepackage[backend=biber,natbib=true,hyperref=true,style=numeric-comp]{biblatex}
+\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
 \addbibresource{references.bib}
-\usepackage[nomargin,inline,draft]{fixme}
-\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
+% \usepackage[nomargin,inline,draft]{fixme}
+% \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
 \usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
@@
 \begin{document}
+\acrodef{EV}[EV]{Extracellular Vesicle}
+\acrodef{SVM}{Support Vector Machine}
 
@@
 })
+options(digits=2)
 library("data.table")
@@
 library("proto")
+library("grid")
 library("xtable")
@@
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide")))>>=
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs")))>>=
 for (a in paste0("../tissue_specific_expression/",
                  c("tissue_specific_markers",
-                   "interesting_gene_reads_wide"))) {
+                   "interesting_gene_reads_wide",
+                   "combined_star_logs"))) {
 
@@
                         ))
-    tissue.specific.gene.reads[,log2.FPKM:=log2(FPKM+1)]
+    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
     return(ggplot(tissue.specific.gene.reads,
                   aes(y=gene_short_name,x=tissue))
-           + geom_tile(aes(fill=log2.FPKM))
-           + scale_fill_gradient(low="white",high="#002058")
+           + geom_tile(aes(fill=`log_2(FPKM)`))
+           + scale_fill_gradient(limits=c(0,17),
+                                 low="white",
+                                 high="#002058",
+                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
            + scale_x_discrete("", expand = c(0, 0))
@@
            + theme_grey(base_size=14)
-           + theme(legend.position = "none",
+           + theme(legend.position = "top",
                axis.ticks = element_blank(), 
+               plot.margin=unit(c(1,1,1,1),"mm"),
                axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
@@
 @
-
-\acro{EV}{Extracellular Vesicles}
 
@@
     tissues from existing publicly-available RNAseq transcriptomes by
-    calculating a tissue specificity index. Using machine learning
-    techniques we trained a multiclass support vector machine to
-    classify the tissue of origin.
+    calculating a tissue specificity index.
+
+    The tissue specificity index for gene $g$ ($\tau_g$) is
+
+    \begin{equation}
+      \tau_g = \frac{\sum_{i=1}^{i=N} 1-\frac{X_{i,g}}{\mathrm{max}(X_g)}}{N-1}
+    \end{equation}
+    \label{eq:tissue_specificity_index}
+
+    where $N$ is the number of tissues, $X_{i,g}$ is the mean
+    expression of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is
+    the maximal expression of gene $g$ across all tissues.
+
+    Using machine learning techniques we trained a multiclass \ac{SVM} to classify the tissue of origin.
 
@@
 
-  \block{Methods}{ Nothing here yet. }
+  \block{Methods}{ %
+    \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
+    %
+    RNAseq reads from %
+    % 
+    \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
+    % 
+    samples corresponding to %
+    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
+    different tissues obtained from the Roadmap Epigenomics
+    \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx
+    \cite{Gtexconsortium.2015:Thegenotypetissueexpression} projects,
+    as well as uterine (GSE50599,
+    \cite{Chan.ea2014:Assessmentofmyometrialtranscriptome}) and
+    placenta (GSE66622,
+    \cite{Hughes.ea2015:Evaluatingintra-andinter-individual}) samples
+    from the SRA.
+
+    Reads were aligned to \textit{Homo sapiens} reference genome
+    GRCh38 with Ensembl annotation release 80 using STAR v2.4.2a
+    \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq} and called using
+    Tophat 2 v2.0.10
+    \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}. The
+    average percentage of uniquely mapped reads was %
+    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}
+    and the average total mapping percentage was %
+    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}.
+    % 
+    For samples where the reads were not publicly available (GTEx and
+    some Roadmap Epigenomics samples), the RPKM were taken from
+    publicly available result files (see code for details).
+
+    The \ac{SVM} was generated using caret and svmRadial with 10
+    cycles of 10 repeats and a tune length of 8.
+
+}
 
@@
 
-\block{References and Acknowledgements}{ The code and datasets for
-  this poster can be found at
+\block{References}{
+
+  \begin{center}
+    \mbox{}\vspace{-\baselineskip}
+  \printbibliography[heading=none]
+\end{center}
+}
+
+\block{Code and Datasets}{
+The code and datasets for this poster can be found at
 \url{http://github.com/uiuc-cgm/exosome_markers_per_med_sep_2015.git}
+
+% 
 }

### 2015-09-11 tissue_specific_expression/R/classification_svm.R
be more strict on read subsets; use fewer cores

@@ -2,7 +2,7 @@ library(data.table)
 library(caret)
 library(reshape2)
 library(doMC)
-registerDoMC(cores=12)
+registerDoMC(cores=8)
 
 args <- c("R/svm_training.R",
           "interesting_gene_reads",

### 2015-09-11 tissue_specific_expression/R/subset_reads.R

@@ -108,11 +108,11 @@ tissue.specificity.index <- function(expression){
     return(sum(1-expression/max(expression))/
                (length(expression)-1))
 }
-R
-min.entropy <- 1.9
-min.tissue.specificity <- 0.98
+
+min.entropy <- 2
+min.tissue.specificity <- 0.985
 min.expression <- 10
-min.var <- 0.01
+min.var <- 1
 
 tissue.specificity <-
     data.table(data.frame(tracking_id=c.reads.wide[,1],

### 2015-09-14 tissue_specific_expression/R/knn_training.R
add predictions object fix name of trained object for svm

@@ -3,5 +3,8 @@ caret.run <-
          run_method="knn",
          run_control=trainControl(method="repeatedcv",number=10,repeats=10,classProbs=TRUE),
          tune_length=8,
-         object_name="trained.caret.knn"
+         object_name="trained.caret.knn",
+         predictions="predictions.caret.knn"
          )
+
+registerDoMC(cores=2)

### 2015-09-14 tissue_specific_expression/R/svm_training.R

@@ -4,5 +4,6 @@ caret.run <-
          run_method="svmRadial",
          run_control=trainControl(method="repeatedcv",number=10,repeats=10,classProbs=TRUE),
          tune_length=8,
-         object_name="trained.caret.knn"
+         object_name="trained.caret.svm",
+         predictions="predictions.caret.svm"
          )

### 2015-09-14 tissue_specific_expression/R/classification_svm.R
add output to show the current progress of the classification routine

@@ -69,6 +69,7 @@ for (group in c("ut.pla.none","uterus","placenta")) {
               tuneLength=caret.run[["tune_length"]],
               metric="ROC"
               )
+    message(paste0("finished ",group))
 }
 save.env <- new.env()
 save.env[[caret.run[["object_name"]]]] <- trained.caret

### 2015-09-14 tissue_specific_expression/R/svm_testing.R
add rule to predict testing results for each of the svm and knn runs

@@ -5,52 +5,49 @@ library(doMC)
 registerDoMC(cores=12)
 
 args <- c("svm_classification",
+          "R/svm_training.R",
           "svm_test_results"
           )
 
 args <- commandArgs(trailingOnly=TRUE)
 
 load(args[1])
+source(args[2])
 
+trained.object <- NULL
+eval(parse(text=paste0("trained.object <- ",caret.run[["object_name"]])))
 
+predictions.caret <- list()
 ### test the predictions using SVM
-gene.testing.predictions.svm.prob <-
-    predict(svm.train,
-            newdata=genes.testing,
-            type="prob")
-gene.testing.predictions.svm.prob$prediction <-
-    colnames(gene.testing.predictions.svm.prob)[apply(gene.testing.predictions.svm.prob,1,which.max)]
+for (group in names(trained.object)) {
+    predictions.caret[[group]] <-
+        list()
+    predictions.caret[[group]][["training"]] <- 
+        predict(trained.object[[group]],
+                newdata=genes.training,
+                type="prob")
+    predictions.caret[[group]][["training"]]$prediction <-
+        colnames(predictions.caret[[group]])[apply(predictions.caret[[group]],
+                                                   1,
+                                                   which.max)]
+    predictions.caret[[group]][["training"]]$actual <-
+        genes.training[,group]
+    predictions.caret[[group]][["testing"]] <-
+        predict(trained.object[[group]],
+                newdata=genes.training,
+                type="prob")
+    predictions.caret[[group]][["testing"]]$prediction <-
+        colnames(predictions.caret[[group]])[apply(predictions.caret[[group]],
+                                                   1,
+                                                   which.max)]
+    predictions.caret[[group]][["testing"]]$actual <-
+        genes.testing[,group]
+}
 
-gene.training.predictions.svm.prob <-
-    predict(svm.train,
-            newdata=genes.training,
-            type="prob")
-gene.training.predictions.svm.prob$prediction <-
-    colnames(gene.training.predictions.svm.prob)[apply(gene.training.predictions.svm.prob,1,which.max)]
+save.env <- new.env()
+save.env[[caret.run[["predictions"]]]] <- predictions.caret
 
-### test the predictions using the KNN
-gene.testing.predictions.knn.prob <-
-    predict(knn.train,
-            newdata=genes.testing,
-            type="prob")
[…]

### 2015-09-14 tissue_specific_expression/R/svm_testing.R
add rules to make svm_test_results

@@ -4,15 +4,76 @@ library(reshape2)
 library(doMC)
 registerDoMC(cores=12)
 
-args <- c("svm_classification",
-          "R/svm_training.R",
+args <- c("R/svm_training.R",
+          "svm_classification",
           "svm_test_results"
           )
 
+## this is a version of predict.train from
+## https://github.com/dondelelcaro/caret/tree/fix_obslevels_in_predict
+my.predict.train <- function(object, newdata = NULL, type = "raw", na.action = na.omit, ...)
+{
+  if(all(names(object) != "modelInfo")) {
+    object <- update(object, param = NULL)
+  }
+  if(!is.null(object$modelInfo$library))
+    for(i in object$modelInfo$library) 
+      do.call("require", list(package = i))
+  if(!(type %in% c("raw", "prob"))) stop("type must be either \"raw\" or \"prob\"")
+  if(type == "prob")
+  {
+    if (is.null(object$modelInfo$prob))
+      stop("only classification models that produce probabilities are allowed")
+  }
+  
+  if(!is.null(newdata))
+  {
+    if (inherits(object, "train.formula"))
+    {
+      newdata <- as.data.frame(newdata)
+      rn <- row.names(newdata)
+      Terms <- delete.response(object$terms)
+      m <- model.frame(Terms, newdata, na.action = na.action, xlev = object$xlevels)
+      if (!is.null(cl <- attr(Terms, "dataClasses"))) 
+        .checkMFClasses(cl, m)
+      keep <- match(row.names(m), rn)
+      newdata <- model.matrix(Terms, m, contrasts = object$contrasts)
+      xint <- match("(Intercept)", colnames(newdata), nomatch = 0)
+      if (xint > 0) 
+        newdata <- newdata[, -xint, drop = FALSE]   
+    }
+  }
+  else {
+    if(!is.null(object$trainingData))
+    {            
+      newdata <- if(object$method == "pam") object$finalModel$xData else object$trainingData
+      ##newdata$.outcome <-NULL
+    } else stop("please specify data via newdata")
+  }
+  
+  if(type == "prob")
+  {
+    out <- extractProb(list(object),
+                       unkX = newdata,
+                       unkOnly = TRUE,
+                       ...)
+    obsLevels <- levels(object)
+    out <- out[, make.names(obsLevels), drop = FALSE]
+  } else {
+    out <- extractPrediction(list(object),
+                             unkX = newdata,
[…]

### 2015-09-14 tissue_specific_expression/R/knn_training.R
only use one core for knn

@@ -7,4 +7,4 @@ caret.run <-
          predictions="predictions.caret.knn"
          )
 
-registerDoMC(cores=2)
+registerDoMC(cores=1)

### 2015-09-14 tissue_specific_expression/R/tissue_specific_expression.R
fix up tissue specific expression to use data.table properly

@@ -41,31 +41,38 @@ tissue.specificity.index <- function(expression){
 ##' @author Don Armstrong
 specific.genes.isoforms <- function(data,min.specificity=0.98,min.max.expression=10) {
     ## calculate the specificity index
+    genes <- data[,gene_short_name]
+    tracking.id=data[,tracking_id]
+    data <- data[,-c(1:2),with=FALSE]
+    if (any(colnames(data)=="universal human reference")) {
+        data <- data[,colnames(data)!="universal human reference",with=FALSE]
+    }
     specificity <-
-        apply(data[,-1],1,tissue.specificity.index)
+        apply(data,1,tissue.specificity.index)
     per.gi.max <-
-        apply(data[,-1],1,max)
+        apply(data,1,max)
     per.gi.which.max <-
-        apply(data[,-1],1,which.max)
+        apply(data,1,which.max)
     possibly.specific <-
         (per.gi.max >= min.max.expression &
              specificity >= min.specificity)
     possibly.specific[is.na(possibly.specific)] <-
         FALSE
     tissue.specific.gi <-
-        data.frame(gene=data[,1],
+        data.frame(gene=genes,
+                   tracking_id=tracking.id,
                    tissue=NA,
                    gi.max=per.gi.max,
                    specificity=specificity,
                    possibly.specific=possibly.specific,
                    gi.which.max=per.gi.which.max)
     tissue.specific.gi$tissue[possibly.specific] <- 
-        colnames(data)[per.gi.which.max+1][possibly.specific]
+        colnames(data)[per.gi.which.max][possibly.specific]
     return(tissue.specific.gi)
 }    
 
-tissue.specific.genes <- specific.genes.isoforms(interesting.gene.reads.wide)
-tissue.specific.isoforms <- specific.genes.isoforms(interesting.isoform.reads.wide)
+tissue.specific.genes <- data.table(specific.genes.isoforms(interesting.gene.reads.wide))
+tissue.specific.isoforms <- data.table(specific.genes.isoforms(interesting.isoform.reads.wide))
 
 save(tissue.specific.genes,
      tissue.specific.isoforms,

### 2015-09-14 poster/poster.Rnw
add roc curves to poster

@@
 \author{Don L. Armstrong, Monica Uddin, Derek Wildman}
+\institute{Institute for Genomic Biology, Computing Genomes for Reproductive Health, University of Illinois, Urbana-Champaign}
 \usetheme{Default}
@@
     \colorlet{framecolor}{black}
-    \colorlet{titlefgcolor}{white}
-    \colorlet{titlebgcolor}{professionalOrange}
+    \colorlet{titlefgcolor}{boldBlue}
+    \colorlet{titlebgcolor}{white}
     \colorlet{blocktitlebgcolor}{professionalOrange}
@@
 \usecolorstyle{UofI}
+\definebackgroundstyle{UofIVertical}{
+    \draw[line width=0pt, bottom color=backgroundcolor, 
+    top color=white] (bottomleft) rectangle (topright);
+}
+\usebackgroundstyle{UofIVertical}
+\usetitlestyle{Empty}
+
+\makeatletter
+\renewcommand\maketitle[1][]{  % #1 keys
+    \normalsize
+    \setkeys{title}{#1}
+    % Title dummy to get title height
+    \node[transparent,inner sep=\TP@titleinnersep, line width=\TP@titlelinewidth, anchor=north, minimum width=\TP@visibletextwidth-2\TP@titleinnersep]
+        (TP@title) at ($(0, 0.5\textheight-\TP@titletotopverticalspace)$) {\parbox{\TP@titlewidth-2\TP@titleinnersep}{\TP@maketitle}};
+    \draw let \p1 = ($(TP@title.north)-(TP@title.south)$) in node {
+        \setlength{\TP@titleheight}{\y1}
+        \setlength{\titleheight}{\y1}
+        \global\TP@titleheight=\TP@titleheight
+        \global\titleheight=\titleheight
+    };
+
+    % Compute title position
+    \setlength{\titleposleft}{-0.5\titlewidth}
+    \setlength{\titleposright}{\titleposleft+\titlewidth}
+    \setlength{\titlepostop}{0.5\textheight-\TP@titletotopverticalspace}
+    \setlength{\titleposbottom}{\titlepostop-\titleheight}
+
+    % Title style (background)
+    \TP@titlestyle
+
+    % Title node
+    \node[inner sep=\TP@titleinnersep, line width=\TP@titlelinewidth, anchor=north, minimum width=\TP@visibletextwidth-2\TP@titleinnersep]
+        at (0,0.5\textheight-\TP@titletotopverticalspace) {\parbox{\TP@titlewidth-2\TP@titleinnersep}{\TP@maketitle}};
+
+    % Settings for blocks
+    \normalsize
+    \setlength{\TP@blocktop}{\titleposbottom-\TP@titletoblockverticalspace}
+}
+\makeatother
 
@@
 \acrodef{SVM}{Support Vector Machine}
+\acrodef{ROC}{Receiver Operating Characteristic}
+\acrodef{KNN}{K Nearest Neighbors}
 
@@
 library("xtable")
+library("caret")
+library("pROC")
 @
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs")))>>=
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results")))>>=
 for (a in paste0("../tissue_specific_expression/",
@@
                    "interesting_gene_reads_wide",
-                   "combined_star_logs"))) {
-
+                   "combined_star_logs",
+                   "svm_test_results"
+                   ))) {
     load(a)
@@
     \begin{equation}
-      \tau_g = \frac{\sum_{i=1}^{i=N} 1-\frac{X_{i,g}}{\mathrm{max}(X_g)}}{N-1}
+      \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
     \end{equation}
@@
 
-    The \ac{SVM} was generated using caret and svmRadial with 10
-    cycles of 10 repeats and a tune length of 8.
+    The \ac{SVM} was generated using caret
+    \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
+    \cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a
+    tune length of 8 with a training set representing $\approx 75\%$ of the
+    samples ($\approx 2450$ samples total, including $\approx 35$
+    uterine or $\approx 60$ placenta) in each of the groups chosen at
+    random, with the remaining samples making up a testing set
+    ($\approx 820$ samples). For comparison purposes, we also
+    generated a \ac{KNN} based predictor using the same training and
+    testing sets of genes.
 
@@
   for those genes. With the exception of FTX\_1, which is expressed in
-  multiple fetal tissues, all genes are highly specific to uterus.}
+  multiple fetal tissues and MYH11, a major component of smooth
+  muscle, all genes are highly specific to uterus.}
 
@@
 
+<<placenta.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
+pl.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"placenta"]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#1c61b6"
+             )
+pl.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"placenta"]),
+              percent=TRUE,
+              col="#008600")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",
+       legend=c("Placenta Testing Set", "Placenta Training Set"),
+       col=c("#1c61b6", "#008600"),
+       lwd=2)
+@ 
+\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM}. The
+  placenta \ac{SVM} is fairly effective at distinguishing between
+  samples of placental origin and non-placental origin.}
+
   % ROC curve for uterus/not uterus
+
+<<uterus.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
+ut.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"uterus"]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#1c61b6"
+             )
+ut.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"uterus"]),
+              percent=TRUE,
+              col="#008600")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",
+       legend=c("Uterus Testing Set", "Uterus Training Set"),
+       col=c("#1c61b6", "#008600"),
+       lwd=2)
+@ 
+\captionof{figure}{\ac{ROC} curve of the uterus \ac{SVM}. The uterus \ac{SVM} is
+  highly effective at distinguishing between samples of uterus origin and
+  non-uterus origin}
+
 }

### 2015-09-14 tissue_specific_expression/R/knn_training.R
switch knn back to 12 cores

@@ -7,4 +7,4 @@ caret.run <-
          predictions="predictions.caret.knn"
          )
 
-registerDoMC(cores=1)
+registerDoMC(cores=12)

### 2015-09-14 poster/poster.Rnw
follow the UofI style more closely

@@
     \colorlet{titlebgcolor}{white}
+    \colorlet{blockbodybgcolor}{white}
     \colorlet{blocktitlebgcolor}{professionalOrange}
-    \colorlet{blocktitlefgcolor}{white}
+    \colorlet{blocktitlefgcolor}{boldBlue}
     }
 \usecolorstyle{UofI}
+\defineblockstyle{UofIMinimal}{
+    titlewidthscale=1, bodywidthscale=1, titleleft,
+    titleoffsetx=0pt, titleoffsety=0pt, bodyoffsetx=0pt, bodyoffsety=0pt,
+    bodyverticalshift=0pt, roundedcorners=0, linewidth=0.2cm,
+    titleinnersep=1cm, bodyinnersep=1cm
+}{
+    \begin{scope}[line width=\blocklinewidth, rounded corners=\blockroundedcorners]
+       \ifBlockHasTitle %
+           \draw[draw=none,fill=blockbodybgcolor]
+               (blockbody.south west) rectangle (blocktitle.north east);
+           \draw[color=blocktitlebgcolor, loosely dashed]
+               (blocktitle.south west) -- (blocktitle.south east);%
+       \else
+             \draw[draw=none,fill=blockbodybgcolor]
+                 (blockbody.south west) rectangle (blockbody.north east);
+        \fi
+    \end{scope}
+}
+
+\useblockstyle{UofIMinimal}
 \definebackgroundstyle{UofIVertical}{
@@
 \usebackgroundstyle{UofIVertical}
-\usetitlestyle{Empty}
+\definetitlestyle{UofITitle}{
+    width=\paperwidth, roundedcorners=0, linewidth=0pt, innersep=12mm,
+    titletotopverticalspace=5mm, titletoblockverticalspace=10mm,
+    titlegraphictotitledistance=10pt
+}{}
+\usetitlestyle{UofITitle}
 
 \makeatletter
-\renewcommand\maketitle[1][]{  % #1 keys
-    \normalsize
-    \setkeys{title}{#1}
-    % Title dummy to get title height
-    \node[transparent,inner sep=\TP@titleinnersep, line width=\TP@titlelinewidth, anchor=north, minimum width=\TP@visibletextwidth-2\TP@titleinnersep]
-        (TP@title) at ($(0, 0.5\textheight-\TP@titletotopverticalspace)$) {\parbox{\TP@titlewidth-2\TP@titleinnersep}{\TP@maketitle}};
-    \draw let \p1 = ($(TP@title.north)-(TP@title.south)$) in node {
-        \setlength{\TP@titleheight}{\y1}
-        \setlength{\titleheight}{\y1}
-        \global\TP@titleheight=\TP@titleheight
-        \global\titleheight=\titleheight
-    };
-
-    % Compute title position
-    \setlength{\titleposleft}{-0.5\titlewidth}
-    \setlength{\titleposright}{\titleposleft+\titlewidth}
-    \setlength{\titlepostop}{0.5\textheight-\TP@titletotopverticalspace}
-    \setlength{\titleposbottom}{\titlepostop-\titleheight}
-
-    % Title style (background)
-    \TP@titlestyle
-
-    % Title node
-    \node[inner sep=\TP@titleinnersep, line width=\TP@titlelinewidth, anchor=north, minimum width=\TP@visibletextwidth-2\TP@titleinnersep]
-        at (0,0.5\textheight-\TP@titletotopverticalspace) {\parbox{\TP@titlewidth-2\TP@titleinnersep}{\TP@maketitle}};
-
-    % Settings for blocks
-    \normalsize
-    \setlength{\TP@blocktop}{\titleposbottom-\TP@titletoblockverticalspace}
+\renewcommand\TP@maketitle{%
+   \begin{minipage}{0.8\textwidth}
+        \color{titlefgcolor}
+        {\bfseries \Huge \sc \@title \par}
+        \vspace*{1em}
+        {\huge \@author \par}
+        \vspace*{1em}
+        {\LARGE \@institute}
+    \end{minipage}%
+    \hfill
+    \begin{minipage}{0.2\linewidth}
+      \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/igb_wordmark}
+      \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
+    \end{minipage}
 }
@@
              ## main="Training vs Testing set for Uterus",
-             col="#1c61b6"
+             col="#002058"
              )
@@
               percent=TRUE,
-              col="#008600")
+              col="#E87722")
 # testobj <- roc.test(rocobj1, rocobj2)
@@
        legend=c("Placenta Testing Set", "Placenta Training Set"),
-       col=c("#1c61b6", "#008600"),
+       col=c("#002058", "#E87722"),
        lwd=2)
@@
        legend=c("Uterus Testing Set", "Uterus Training Set"),
-       col=c("#1c61b6", "#008600"),
+       col=c("#002058", "#E87722"),
        lwd=2)

### 2015-09-15 poster/poster.Rnw
move UofI logo to the end
start to add conclusions

@@
       \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/igb_wordmark}
-      \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
     \end{minipage}
@@
 \block{Conclusions}{
+  \begin{itemize}
+  \item Pregnancy-specific glycoproteins (PSG1-8) are specific to
+    placental tissues,and PAEP Progestagen-Associated Endometrial
+    Protein is specific to uterus, as previously described.
+  \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
+  \item \ac{SVM} are able to accurately identify the tissue of origin
+    for uterus and placenta.
+%   \item Non-maternal variants may also be able to distinguish between
+%     placental reads and maternal reads in addition to the approach
+%     used here.
+  \end{itemize}
 }
@@
 }
+\block{}{
+  \includegraphics[width=\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
+}
 

### 2015-09-15 poster/poster.Rnw
add roc of reduced reads

@@
 \usepackage{zref-xr,zref-user}
+% \renewcommand*{\bibfont}{\small}
 
@@
 library("pROC")
+
+capwords <- function(s, strict = FALSE) {
+    cap <- function(s) paste(toupper(substring(s, 1, 1)),
+                             {s <- substring(s, 2); if(strict) tolower(s) else s},
+                             sep = "", collapse = " " )
+    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
+}
 @
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results")))>>=
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
 for (a in paste0("../tissue_specific_expression/",
@@
                    "combined_star_logs",
-                   "svm_test_results"
+                   "svm_test_results",
+                   "svm_test_results_reduced"
                    ))) {
@@
 
-<<placenta.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
+<<placenta.roc.curve,fig.width=8,fig.height=5,echo=FALSE>>=
+layout(matrix(1:2,1,2))
 pl.roc.obj1 <- 
@@
        lwd=2)
-@ 
-\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM}. The
-  placenta \ac{SVM} is fairly effective at distinguishing between
-  samples of placental origin and non-placental origin.}
 
-  % ROC curve for uterus/not uterus
-
-<<uterus.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
 ut.roc.obj1 <- 
@@
 @ 
-\captionof{figure}{\ac{ROC} curve of the uterus \ac{SVM}. The uterus \ac{SVM} is
-  highly effective at distinguishing between samples of uterus origin and
-  non-uterus origin}
+\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM} (left) and
+  uterus \ac{SVM} right) . Both \ac{SVM} are highly effective at
+  distinguishing between samples of uterus origin and non-uterus
+  origin.}
+\label{fig:plac_ut_full_roc}
+}
 
+\block{Identification of tissue in reduced reads}{
+  % Describe reduced read approach
+  % 
+<<placenta.roc.reduced.curve,fig.width=8,fig.height=10,echo=FALSE>>=
+layout(matrix(1:2,2,1))
+for (a in c("placenta","uterus")) {
+pl.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#002058"
+             )
+pl.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
+              percent=TRUE,
+              col="#E87722")
+pl.roc.obj3 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
+              percent=TRUE,
+              col="green")
+pl.roc.obj4 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
+              percent=TRUE,
+              col="brown")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
+       col=c("brown","green","#002058", "#E87722"),
+       lwd=2)
+}
+@ 
+\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
+  Uterus samples with the indicated number of total reads. Reads were
+  reduced by sampling reads for each sample weighted by the FPKM
+  measured for each sample, and run using the same trained SVM
+  depicted in Figure \ref{fig:plac_ut_full_roc}. %
+  % 
+  The placenta \ac{SVM} is fairly accurate at distinguishing between
+  samples of placental origin and non-placental origin even when the
+  number of reads is very low (10). The uterus \ac{SVM} requires many
+  more reads in order to achieve similar accuracy. This is likely due
+  to the many fewer uterus specific genes.}
 }
@@
 \column{0.25}
-
-\block{Identification of tissue in reduced reads}{
-  % Describe reduced read approach
-
-  % 
-}
 
@@
     for uterus and placenta.
-%   \item Non-maternal variants may also be able to distinguish between
-%     placental reads and maternal reads in addition to the approach
-%     used here.
+  \item Very few sequenced reads are required to accurately identify
+    tissues of Placental origin. We are able to detect 100\% of the
+    placenta tissues with 80\% specificity with just 10 reads.
+  \item Non-maternal variants may also be able to distinguish between
+    placental reads and maternal reads in addition to the approach
+    used here.
   \end{itemize}
@@
 \block{Code and Datasets}{
-The code and datasets for this poster can be found at
-\url{http://github.com/uiuc-cgm/exosome_markers_per_med_sep_2015.git}
+The code and underlying data for this poster can be found at
+\url{http://dla2.us/p/em2015}
 

### 2015-09-15 tissue_specific_expression/R/classification_svm.R
reduce the reads into the classifier and scale them

@@ -1,7 +1,7 @@
-library(data.table)
-library(caret)
-library(reshape2)
-library(doMC)
+library("data.table")
+library("caret")
+library("reshape2")
+library("doMC")
 registerDoMC(cores=8)
 
 args <- c("R/svm_training.R",
@@ -18,10 +18,15 @@ load(args[2])
 
 gene.fpkm.wide <-
     dcast(interesting.gene.reads,
-          Sample_Group+SRX~gene_short_name,
+          Sample_Group+SRX~tracking_id,
           value.var="FPKM",
           fun.aggregate=function(x){x[1]})
+## set NA to zero
+gene.fpkm.wide[is.na(gene.fpkm.wide)] <- 0
 
+## ignore genes with maximum expression less than 100 FPKM.
+gene.max <- apply(gene.fpkm.wide,2,function(x){max(x,na.rm=TRUE)})
+gene.fpkm.wide <- gene.fpkm.wide[,c(1,2,which(as.numeric(gene.max)>100))]
 gene.names <- colnames(gene.fpkm.wide)[-(1:2)]
 
 ### for the time being, we're interested in uterus, placenta, or neither.
@@ -40,9 +45,6 @@ for (group in c("uterus","placenta")) {
         as.factor(gene.fpkm.wide[[group]])
 }
 
-## set NA to zero
-gene.fpkm.wide[is.na(gene.fpkm.wide)] <- 0
-
 ### seed generated with echo $RANDOM
 set.seed(10609)
 in_training <-
@@ -57,13 +59,16 @@ genes.testing <-
 
 trained.caret <- list()
 ### multi-class ROC https://www.kaggle.com/c/forest-cover-type-prediction/forums/t/10573/r-caret-using-roc-instead-of-accuracy-in-model-training
-for (group in c("ut.pla.none","uterus","placenta")) {
+for (group in c("uterus","placenta")) {
     ## seed generated with echo $RANDOM;
     set.seed(29224)
+    data.scaled <- scale(genes.training[,(gene.names)])
+    data.scaled[!is.finite(data.scaled)] <- 0
+
     trained.caret[[group]] <-
         train(form=as.formula(paste0(group,"~.")),
-              data=genes.training[,c(group,gene.names)],
-              preProc=c("center","scale"),
+              data=data.frame(genes.training[,group,drop=FALSE],
+                  data.scaled),
               method=caret.run[["run_method"]],
               trControl=caret.run[["run_control"]],
               tuneLength=caret.run[["tune_length"]],

### 2015-09-15 tissue_specific_expression/R/knn_training.R

@@ -1,10 +1,10 @@
 caret.run <-
     list(run_name="knn",
          run_method="knn",
-         run_control=trainControl(method="repeatedcv",number=10,repeats=10,classProbs=TRUE),
-         tune_length=8,
+         run_control=trainControl(method="repeatedcv",number=5,repeats=3,classProbs=TRUE),
+         tune_length=3,
          object_name="trained.caret.knn",
          predictions="predictions.caret.knn"
          )
 
-registerDoMC(cores=12)
+registerDoMC(cores=2)

### 2015-09-15 tissue_specific_expression/R/svm_testing.R

@@ -78,31 +78,40 @@ source(args[1])
 trained.object <- NULL
 eval(parse(text=paste0("trained.object <- ",caret.run[["object_name"]])))
 
+scale.and.zero <- function(x){
+    x <- scale(x)
+    x[!is.finite(x)] <- 0
+    x
+}
+
 predictions.caret <- list()
 ### test the predictions using SVM
 for (group in names(trained.object)) {
     predictions.caret[[group]] <-
         list()
-    predictions.caret[[group]][["training"]] <- 
+    result <- 
         my.predict.train(trained.object[[group]],
-                        newdata=genes.training[,c(gene.names)],
-                        type="prob")
-    predictions.caret[[group]][["training"]]$prediction <-
-        colnames(predictions.caret[[group]][["training"]])[apply(predictions.caret[[group]][["training"]],
-                                                   1,
-                                                   which.max)]
-    predictions.caret[[group]][["training"]]$actual <-
+                         newdata=scale.and.zero(genes.training[,c(gene.names)]),
+                         type="prob")
+    result$prediction <-
+        colnames(result)[apply(result,1,which.max)]
+    result$actual <-
         genes.training[,group]
-    predictions.caret[[group]][["testing"]] <-
+    
+    predictions.caret[[group]][["training"]] <- 
+        result
+
+    result <- 
         my.predict.train(trained.object[[group]],
-                newdata=genes.testing,
-                type="prob")
-    predictions.caret[[group]][["testing"]]$prediction <-
-        colnames(predictions.caret[[group]][["testing"]])[apply(predictions.caret[[group]][["testing"]],
-                                                   1,
-                                                   which.max)]
-    predictions.caret[[group]][["testing"]]$actual <-
+                         newdata=scale.and.zero(genes.testing[,c(gene.names)]),
+                         type="prob")
+    result$prediction <-
+        colnames(result)[apply(result,1,which.max)]
+    result$actual <-
         genes.testing[,group]
+    
+    predictions.caret[[group]][["testing"]] <- 
+        result
 }
 
 save.env <- new.env()

### 2015-09-15 tissue_specific_expression/R/svm_testing_reduced.R
add reduced svm testing

@@ -1,0 +1,129 @@
+library("data.table")
+library("caret")
+library("reshape2")
+library("doMC")
+library("sampling")
+registerDoMC(cores=12)
+
+args <- c("R/svm_training.R",
+          "svm_classification",
+          "svm_test_results_reduced"
+          )
+
+## this is a version of predict.train from
+## https://github.com/dondelelcaro/caret/tree/fix_obslevels_in_predict
+my.predict.train <- function(object, newdata = NULL, type = "raw", na.action = na.omit, ...)
+{
+  if(all(names(object) != "modelInfo")) {
+    object <- update(object, param = NULL)
+  }
+  if(!is.null(object$modelInfo$library))
+    for(i in object$modelInfo$library) 
+      do.call("require", list(package = i))
+  if(!(type %in% c("raw", "prob"))) stop("type must be either \"raw\" or \"prob\"")
+  if(type == "prob")
+  {
+    if (is.null(object$modelInfo$prob))
+      stop("only classification models that produce probabilities are allowed")
+  }
+  
+  if(!is.null(newdata))
+  {
+    if (inherits(object, "train.formula"))
+    {
+      newdata <- as.data.frame(newdata)
+      rn <- row.names(newdata)
+      Terms <- delete.response(object$terms)
+      m <- model.frame(Terms, newdata, na.action = na.action, xlev = object$xlevels)
+      if (!is.null(cl <- attr(Terms, "dataClasses"))) 
+        .checkMFClasses(cl, m)
+      keep <- match(row.names(m), rn)
+      newdata <- model.matrix(Terms, m, contrasts = object$contrasts)
+      xint <- match("(Intercept)", colnames(newdata), nomatch = 0)
+      if (xint > 0) 
+        newdata <- newdata[, -xint, drop = FALSE]   
+    }
+  }
+  else {
+    if(!is.null(object$trainingData))
+    {            
+      newdata <- if(object$method == "pam") object$finalModel$xData else object$trainingData
+      ##newdata$.outcome <-NULL
+    } else stop("please specify data via newdata")
+  }
+  
+  if(type == "prob")
+  {
+    out <- extractProb(list(object),
+                       unkX = newdata,
+                       unkOnly = TRUE,
+                       ...)
+    obsLevels <- levels(object)
+    out <- out[, make.names(obsLevels), drop = FALSE]
+  } else {
+    out <- extractPrediction(list(object),
+                             unkX = newdata,
[…]

### 2015-09-16 poster/poster.Rnw
add roc of reduced reads

@@
     origin and can be obtained without invasive procedures, they are a
-    promising remote sensor of the transcriptome of tissues which
+    promising remote sensor of the transcriptome of tissues that
     would otherwise be inaccessible.
@@
 
-    Using machine learning techniques we trained a multiclass \ac{SVM} to classify the tissue of origin.
+    we we trained a multiclass \ac{SVM} to accurately identify the
+    tissue of origin, and present the results of the identification of
+    reads from uterine and placental origin here. Tables showing
+    tissue specific genes from the other %
+    %
+    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
+    % 
+    tissues can be found on \url{http://dla2.us/p/em2015}.
 
@@
     % 
-    \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
+    \Sexpr{nrow(genes.training)+nrow(genes.testing)} %
     % 

### 2015-09-16 poster/poster.Rnw
clarify figure legends and make other changes Monica suggested

@@
 \acrodef{KNN}{K Nearest Neighbors}
+\acrodef{SRA}{Sequence Read Archive}
 
@@
     reads from uterine and placental origin here. Tables showing
-    tissue specific genes from the other %
+    tissue specific genes from all %
     %
@@
     % 
-    \Sexpr{nrow(genes.training)+nrow(genes.testing)} %
+    \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
     % 
-    samples corresponding to %
+    samples (supplemented with FPKMs from
+    \Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
+      nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
+      length(star.logs[grepl("Number of input reads",field),value])} %
+    %
+    additional samples without raw data in the \ac{SRA})
+    % 
+    corresponding to %
     \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
-    different tissues obtained from the Roadmap Epigenomics
+    different tissues were obtained from the Roadmap Epigenomics
     \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx
@@
     \cite{Hughes.ea2015:Evaluatingintra-andinter-individual}) samples
-    from the SRA.
+    from the \ac{SRA}.
 
@@
     random, with the remaining samples making up a testing set
-    ($\approx 820$ samples). For comparison purposes, we also
-    generated a \ac{KNN} based predictor using the same training and
-    testing sets of genes.
+    ($\approx 820$ samples).
 
@@
 }
-\captionof{figure}{Uterus specific genes. $\log_2 \mathrm{FPKM}$ of
-  genes with tissue specific indices $\ge 0.98$ which are specific to
-  uterus tissue in uterus and tissues with second highest expression
-  for those genes. With the exception of FTX\_1, which is expressed in
-  multiple fetal tissues and MYH11, a major component of smooth
-  muscle, all genes are highly specific to uterus.}
+\captionof{figure}{Expression of uterus specific genes in uterus and
+  tissues with the second highest expression for those genes.
+  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
+  if they have a tissue specific expression $\tau_g \ge 0.98$. The
+  known uterus specific gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp)
+  and AL356310.1 (94bp) are highly uterus specific.}
 
@@
 }
-\captionof{figure}{Placenta specific genes. $\log_2 \mathrm{FPKM}$ of
-  genes with tissue specific indices $\ge 0.98$ and
-  $\mathrm{FPKM} \ge 1024$ which are specific to uterus tissue in
-  placenta and tissues with second highest expression for those
-  genes.}
+\captionof{figure}{Expression of placenta specific genes in placenta
+  and tissues with the second highest expression for those genes.
+  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
+  if they have a tissue specific expression $\tau_g \ge 0.98$ and
+  $\mathrm{FPKM} \ge 1024$ in placenta. The known placenta specific
+  PSG genes and CSH genes are all highly placenta specific, and short
+  RNAs mir-126, mir-622, and PAPPA-AS1 are fairly placenta specific.}
 }
@@
 
-<<placenta.roc.curve,fig.width=8,fig.height=5,echo=FALSE>>=
+<<placenta.roc.curve,fig.width=8,fig.height=4,echo=FALSE>>=
 layout(matrix(1:2,1,2))
@@
   \item Pregnancy-specific glycoproteins (PSG1-8) are specific to
-    placental tissues,and PAEP Progestagen-Associated Endometrial
-    Protein is specific to uterus, as previously described.
-  \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
+    placental tissues, and PAEP is specific to uterus, as previously
+    described.
+%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
   \item \ac{SVM} are able to accurately identify the tissue of origin
@@
     placenta tissues with 80\% specificity with just 10 reads.
+  \item We can easily identify exosomes originating from the placenta,
+    as even the smallest exosomes can contain 70 RNA molecules
+    \cite{Li.ea2014:AnalysisofRNAcontent}.
   \item Non-maternal variants may also be able to distinguish between
-    placental reads and maternal reads in addition to the approach
-    used here.
+    placental reads and maternal reads.
   \end{itemize}

### 2015-09-17 poster/poster.Rnw
fix captions and gene levels in poster figures

@@
         colnames(interesting.gene.reads.wide)[
-           which(apply(apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
-                             function(x){order(-x)==2}),1,any))+2]
+         apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
+                             function(x){order(-x)[2]+2})]
+
     interesting.tissues <- 
-        c("gene_short_name",tissue.name,interesting.tissues)
+        c("gene_short_name",tissue.name,unique(interesting.tissues))
 
@@
     tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
+    gene.levels <- 
+        unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
+    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
     return(ggplot(tissue.specific.gene.reads,
-                  aes(y=gene_short_name,x=tissue))
+                  aes(y=gene_short_name,
+                          x=tissue))
            + geom_tile(aes(fill=`log_2(FPKM)`))
@@
                    axis.ticks = element_blank(),
+                   panel.grid=element_blank(),
+                   panel.border=element_blank(),
+                   panel.background=element_blank(),
+                   plot.background=element_blank(),
                    plot.margin=unit(c(1,1,1,1),"mm"),
@@
 }
-\captionof{figure}{Expression of uterus specific genes in uterus and
-  tissues with the second highest expression for those genes.
-  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
-  if they have a tissue specific expression $\tau_g \ge 0.98$. The
-  known uterus specific gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp)
-  and AL356310.1 (94bp) are highly uterus specific.}
+\captionof{figure}{Expression of uterus specific genes
+  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
+  highest expression for any of the uterus specific genes ordered by
+  their specificity to uterus. The known uterus specific gene PAEP and
+  short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly
+  uterus specific.}
 
@@
 }
-\captionof{figure}{Expression of placenta specific genes in placenta
-  and tissues with the second highest expression for those genes.
-  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
-  if they have a tissue specific expression $\tau_g \ge 0.98$ and
-  $\mathrm{FPKM} \ge 1024$ in placenta. The known placenta specific
-  PSG genes and CSH genes are all highly placenta specific, and short
-  RNAs mir-126, mir-622, and PAPPA-AS1 are fairly placenta specific.}
+\captionof{figure}{Expression of placenta specific genes
+  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
+  in placenta and all tissues with the second highest expression for
+  any placenta-specific gene ordered by their specificity to placenta.
+  The previously known placenta-specific PSG genes and CSH genes are
+  highly placenta specific and the short RNAs mir-126, mir-622, and
+  PAPPA-AS1 are fairly placenta specific.}
 }
@@
   more reads in order to achieve similar accuracy. This is likely due
-  to the many fewer uterus specific genes.}
+  to the much smaller number of uterus specific genes in comparison to
+  placenta.}
 }

### 2015-09-17 poster/poster.Rnw
add qr code; fix a typo; make bibliography smaller

@@
 \usepackage{acronym}
+\usepackage{qrcode}
 \usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
@@
 \usepackage{zref-xr,zref-user}
-% \renewcommand*{\bibfont}{\small}
+\renewcommand*{\bibfont}{\small}
 
@@
 
-    we we trained a multiclass \ac{SVM} to accurately identify the
+    We also trained a multiclass \ac{SVM} to accurately identify the
     tissue of origin, and present the results of the identification of
@@
     average percentage of uniquely mapped reads was %
-    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}
+    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}\%
     and the average total mapping percentage was %
-    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}.
+    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}\%.
     % 
@@
   \begin{itemize}
-  \item Pregnancy-specific glycoproteins (PSG1-8) are specific to
-    placental tissues, and PAEP is specific to uterus, as previously
-    described.
+  \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
+    specific to placental tissues, and PAEP is specific to uterus, as
+    previously described.
 %   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
@@
 The code and underlying data for this poster can be found at
+\begin{center}
+\qrcode[padding]{http://dla2.us/p/em2015}
 \url{http://dla2.us/p/em2015}
-
+\end{center}
 % 

### 2015-09-17 poster/poster.Rnw
switch to diagnostic

@@
     identifying the source tissue of a specific \ac{EV}. To identify
-    the source tissue of a specific \ac{EV}, we have identified
-    tissue-specific miRNA and mRNA markers for multiple organs and
+    the source tissue of a specific \ac{EV}, we have identified miRNA
+    and mRNA markers which are diagnostic of multiple organs and
     tissues from existing publicly-available RNAseq transcriptomes by
@@
 \column{0.25}
-  \block{Uterus Specific Genes}{
+  \block{Genes Diagnostic of Uterus}{
 
@@
 print(specific.markers.plot("uterus")
-      + xlab("Uterus Specific Genes")
+      + xlab("Genes Diagnostic of Uterus ")
       + ylab("Tissue"))
@@
 }
-\captionof{figure}{Expression of uterus specific genes
+\captionof{figure}{Expression of genes diagnostic of uterus
   ($\tau_g \ge 0.98$) in uterus and all tissues with the second
-  highest expression for any of the uterus specific genes ordered by
-  their specificity to uterus. The known uterus specific gene PAEP and
-  short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly
-  uterus specific.}
+  highest expression for any of the uterus-diagnostic genes
+  ordered by their tissue specificity. The known uterus-diagnostic
+  gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp)
+  are highly diagnostic of uterus.}
 
@@
 
-  \block{Placenta Specific Genes}{
+  \block{Genes Diagnostic of Placenta}{
 {\centering
@@
                                ,gene]))
-      + xlab("Uterus Specific Genes")
+      + xlab("Genes Diagnostic of Placenta")
       + ylab("Tissue"))
@@
 }
-\captionof{figure}{Expression of placenta specific genes
+\captionof{figure}{Expression of genes diagnostic of placenta
   ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
   in placenta and all tissues with the second highest expression for
-  any placenta-specific gene ordered by their specificity to placenta.
-  The previously known placenta-specific PSG genes and CSH genes are
-  highly placenta specific and the short RNAs mir-126, mir-622, and
-  PAPPA-AS1 are fairly placenta specific.}
+  any placenta-diagnostic gene ordered by their specificity to
+  placenta. The previously known placenta-diagnostic PSG genes and
+  CSH genes are nearly unique to placenta and the short RNAs mir-126,
+  mir-622, and PAPPA-AS1 are less so.}
 }
@@
   \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
-    specific to placental tissues, and PAEP is specific to uterus, as
-    previously described.
+    diagnostic of placental tissues, and PAEP is diagnostic of
+    uterus, as previously described.
 %   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta

### 2015-09-17 poster/poster.Rnw
switch back to characterizing

--- 2015-09-17 poster/poster.Rnw
@@
     identifying the source tissue of a specific \ac{EV}. To identify
-    the source tissue of a specific \ac{EV}, we have identified
-    tissue-specific miRNA and mRNA markers for multiple organs and
+    the source tissue of a specific \ac{EV}, we have identified miRNA
+    and mRNA markers which are characteristic of multiple organs and
     tissues from existing publicly-available RNAseq transcriptomes by
@@
 \column{0.25}
-  \block{Uterus Specific Genes}{
+  \block{Genes Characteristic of Uterus}{
 
@@
 print(specific.markers.plot("uterus")
-      + xlab("Uterus Specific Genes")
+      + xlab("Genes Characteristic of Uterus ")
       + ylab("Tissue"))
@@
 }
-\captionof{figure}{Expression of uterus specific genes
+\captionof{figure}{Expression of genes characteristic of uterus
   ($\tau_g \ge 0.98$) in uterus and all tissues with the second
-  highest expression for any of the uterus specific genes ordered by
-  their specificity to uterus. The known uterus specific gene PAEP and
-  short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly
-  uterus specific.}
+  highest expression for any of the uterus-characterizing genes
+  ordered by their tissue specificity. The known uterus-characterizing
+  gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp)
+  are highly characteristic of uterus.}
 
@@
 
-  \block{Placenta Specific Genes}{
+  \block{Genes Characteristic of Placenta}{
 {\centering
@@
                                ,gene]))
-      + xlab("Uterus Specific Genes")
+      + xlab("Genes Characteristic of Placenta")
       + ylab("Tissue"))
@@
 }
-\captionof{figure}{Expression of placenta specific genes
+\captionof{figure}{Expression of genes characteristic of placenta
   ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
   in placenta and all tissues with the second highest expression for
-  any placenta-specific gene ordered by their specificity to placenta.
-  The previously known placenta-specific PSG genes and CSH genes are
-  highly placenta specific and the short RNAs mir-126, mir-622, and
-  PAPPA-AS1 are fairly placenta specific.}
+  any placenta-characterizing gene ordered by their specificity to
+  placenta. The previously known placenta-characterizing PSG genes and
+  CSH genes are nearly unique to placenta and the short RNAs mir-126,
+  mir-622, and PAPPA-AS1 are less so.}
 }
@@
   \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
-    specific to placental tissues, and PAEP is specific to uterus, as
-    previously described.
+    characteristic of placental tissues, and PAEP is characteristic of
+    uterus, as previously described.
 %   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta

### 2015-09-25 slides/slides.Rnw
add start of slides for presentation based on the poster

--- 2015-09-17 poster/poster.Rnw
@@
-\documentclass[a0paper,landscape]{tikzposter}
+\documentclass[ignorenonframetext]{beamer}
 \usepackage{fontspec}
@@
 % \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
-\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
+% \usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
 \usepackage{texshade}
@@
 \usepackage{zref-xr,zref-user}
-\renewcommand*{\bibfont}{\small}
+\renewcommand*{\bibfont}{\tiny}
 
@@
 
-\title{Identifying the Tissue of Origin of Extracellular Vesicles Using RNA Expression Signatures}
-\author{Don L. Armstrong, Monica Uddin, Derek Wildman}
-\institute{Institute for Genomic Biology, Computing Genomes for Reproductive Health, University of Illinois, Urbana-Champaign}
-\usetheme{Default}
+\mode<presentation>{ 
+  \usetheme{CambridgeUS}
+  % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
+  \definecolor{ilboldblue}{HTML}{002058}
+  \definecolor{ilboldorange}{HTML}{E87722}
+  \definecolor{ilblue}{HTML}{606EB2}
+  \definecolor{ilorange}{HTML}{D45D00}
+  \setbeamercolor{alerted text}{fg=ilboldorange}
+  \setbeamercolor*{palette primary}{fg=ilblue,bg=ilorange}
+  \setbeamercolor*{palette secondary}{fg=ilblue!20!white,bg=ilorange}
+  \setbeamercolor*{palette tertiary}{bg=ilblue,fg=ilorange}
+  \setbeamercolor*{palette quaternary}{fg=ilblue,bg=ilorange}
 
-% use this
-% http://www.identitystandards.illinois.edu/graphicstandardsmanual/othermedia/posterpresentation.html
-\definecolorstyle{UofI}{
-  \definecolor{boldBlue}{HTML}{002058}
-  \definecolor{boldOrange}{HTML}{E87722}
-  \definecolor{professionalBlue}{HTML}{606EB2}
-  \definecolor{professionalOrange}{HTML}{D45D00}
-  }{
-    \colorlet{backgroundcolor}{professionalBlue}
-    \colorlet{framecolor}{black}
-    \colorlet{titlefgcolor}{boldBlue}
-    \colorlet{titlebgcolor}{white}
-    \colorlet{blockbodybgcolor}{white}
-    \colorlet{blocktitlebgcolor}{professionalOrange}
-    \colorlet{blocktitlefgcolor}{boldBlue}
-    }
-\usecolorstyle{UofI}
-\defineblockstyle{UofIMinimal}{
-    titlewidthscale=1, bodywidthscale=1, titleleft,
-    titleoffsetx=0pt, titleoffsety=0pt, bodyoffsetx=0pt, bodyoffsety=0pt,
-    bodyverticalshift=0pt, roundedcorners=0, linewidth=0.2cm,
-    titleinnersep=1cm, bodyinnersep=1cm
-}{
-    \begin{scope}[line width=\blocklinewidth, rounded corners=\blockroundedcorners]
-       \ifBlockHasTitle %
-           \draw[draw=none,fill=blockbodybgcolor]
-               (blockbody.south west) rectangle (blocktitle.north east);
-           \draw[color=blocktitlebgcolor, loosely dashed]
-               (blocktitle.south west) -- (blocktitle.south east);%
-       \else
-             \draw[draw=none,fill=blockbodybgcolor]
-                 (blockbody.south west) rectangle (blockbody.north east);
-        \fi
-    \end{scope}
+  \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}
+
+  \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
+  \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
+  \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
+  \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}
+
+  % \setbeamercolor*{titlelike}{parent=palette primary}
+  \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
+  \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!80!white}
+  \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}
+
+  \setbeamercolor*{separation line}{}
+  \setbeamercolor*{fine separation line}{}
+  \setbeamercovered{transparent}  
+  \logo{\begin{tikzpicture}% Pale figure
+      {\node[opacity=0.8]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
+        };}%
+    \end{tikzpicture}}
 }
 
-\useblockstyle{UofIMinimal}
-\definebackgroundstyle{UofIVertical}{
-    \draw[line width=0pt, bottom color=backgroundcolor, 
-    top color=white] (bottomleft) rectangle (topright);
-}
-\usebackgroundstyle{UofIVertical}
-\definetitlestyle{UofITitle}{
-    width=\paperwidth, roundedcorners=0, linewidth=0pt, innersep=12mm,
-    titletotopverticalspace=5mm, titletoblockverticalspace=10mm,
-    titlegraphictotitledistance=10pt
-}{}
-\usetitlestyle{UofITitle}
-
-\makeatletter
-\renewcommand\TP@maketitle{%
-   \begin{minipage}{0.8\textwidth}
-        \color{titlefgcolor}
-        {\bfseries \Huge \sc \@title \par}
-        \vspace*{1em}
-        {\huge \@author \par}
-        \vspace*{1em}
-        {\LARGE \@institute}
-    \end{minipage}%
-    \hfill
-    \begin{minipage}{0.2\linewidth}
-      \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/igb_wordmark}
-    \end{minipage}
-}
-\makeatother
+\title{Identifying Extracellular Vesicles of Placenta and Uterine 
+  Origin Using RNA Expression Signatures}
+\author{Don L. Armstrong, Monica Uddin, Derek Wildman}
+\institute{Institute for Genomic Biology, Computing Genomes 
+  for Reproductive Health, University of Illinois, Urbana-Champaign}
+\subject{RNA expression}
 
 \begin{document}
-\acrodef{EV}[EV]{Extracellular Vesicle}
-\acrodef{SVM}{Support Vector Machine}
-\acrodef{ROC}{Receiver Operating Characteristic}
-\acrodef{KNN}{K Nearest Neighbors}
-\acrodef{SRA}{Sequence Read Archive}
 
 <<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
-opts_chunk$set(dev="cairo_pdf",out.width="0.20\\columnwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
+opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.8\\textheight",out.extra="keepaspectratio")
 #opts_chunk$set(cache=TRUE, autodep=TRUE)
@@
 
-\maketitle
+\IfFileExists{figures/relevant_xkcd.png}{\frame[plain]{\centering \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{figures/relevant_xkcd.png}
+
+    \url{https://xkcd.com/1217/}}}
+
+\frame[plain]{\titlepage}
 
@@
 
-\begin{columns}
+\section{Introduction}
 
-  \column{0.25}
+\begin{frame}{Introduction: \acsp{EV}}
+  \acp{EV} are
+  \begin{itemize}
+  \item small (40-1000nm) compartments
+  \item found in blood, saliva, and urine. 
+  \item originate from the plasma membrane or multivesicular endosome
+  \item from many cell types
+  \item contain proteins, lipids, mRNA, miRNA, and other ncRNAs.
+  \item \alert{remote sensor of the transcriptome of inaccessible tissues}
+  \end{itemize}
+\end{frame}
 
-  \block{Introduction}{\acp{EV} are small (40-1000nm) compartments
-    that can be found in blood, saliva, and urine. \acp{EV} originate
-    from the plasma membrane or multivesicular endosome of many cell
-    types and contain proteins, lipids, mRNA, miRNA, and many other
-    non-coding RNAs. Because \acp{EV} contain mRNA from the cell of
-    origin and can be obtained without invasive procedures, they are a
-    promising remote sensor of the transcriptome of tissues that
-    would otherwise be inaccessible.
+\subsection{Identifying source tissue of \acp{EV}}
+
+\begin{frame}{Identifying source tissue of \acp{EV}}
 
   One requirement for remotely sensing the transcriptome is
-    identifying the source tissue of a specific \ac{EV}. To identify
-    the source tissue of a specific \ac{EV}, we have identified miRNA
-    and mRNA markers which are characteristic of multiple organs and
-    tissues from existing publicly-available RNAseq transcriptomes by
+  identifying the source tissue of a specific \ac{EV}. To identify the
+  source tissue of a specific \ac{EV}, we have identified miRNA and
+  mRNA markers which are characteristic of multiple organs and tissues
+  from existing publicly-available RNAseq transcriptomes by
   calculating a tissue specificity index.
@@
 
-    where $N$ is the number of tissues, $X_{i,g}$ is the mean
-    expression of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is
-    the maximal expression of gene $g$ across all tissues.
+  where $N$ is the number of tissues, $X_{i,g}$ is the mean expression
+  of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
+  expression of gene $g$ across all tissues.
 
+ \end{frame}
+
+ \begin{frame}{\ac{SVM}}
     We also trained a multiclass \ac{SVM} to accurately identify the
@@
 %     individual patient. 
-  }
+\end{frame}
 
-  \block{Methods}{ %
-    \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
+\section{Methods}
+\begin{frame}{Methods}
+  \begin{itemize}
   % 
-    RNAseq reads from %
+  \item \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
     %
-    \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
+    RNAseq reads (\Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]/length(star.logs[grepl("Number of input reads",field),value])} average per sample)%
+    % 
+  \item  \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
     % 
@@
     %
-    corresponding to %
-    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
-    different tissues were obtained from the Roadmap Epigenomics
-    \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx
-    \cite{Gtexconsortium.2015:Thegenotypetissueexpression} projects,
-    as well as uterine (GSE50599,
+
+  \item  \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
+    different tissues
+
+  \item Samples from Roadmap Epigenomics
+    \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping}, GTEx
+    \cite{Gtexconsortium.2015:Thegenotypetissueexpression},
+    uterine (GSE50599,
     \cite{Chan.ea2014:Assessmentofmyometrialtranscriptome}) and
     placenta (GSE66622,
-    \cite{Hughes.ea2015:Evaluatingintra-andinter-individual}) samples
-    from the \ac{SRA}.
+    \cite{Hughes.ea2015:Evaluatingintra-andinter-individual})
+  \item aligned to GRCh38 using STAR
+    \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq}
+  \item called using Cufflinks 2 v2.0.10
+    \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}.
 
-    Reads were aligned to \textit{Homo sapiens} reference genome
-    GRCh38 with Ensembl annotation release 80 using STAR v2.4.2a
-    \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq} and called using
-    Tophat 2 v2.0.10
-    \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}. The
-    average percentage of uniquely mapped reads was %
+  \item Mean uniquely mapped reads: %
     \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}\%
-    and the average total mapping percentage was %
+  \item Mean total mapping percentage: %
       \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}\%.
     % 
-    For samples where the reads were not publicly available (GTEx and
-    some Roadmap Epigenomics samples), the RPKM were taken from
-    publicly available result files (see code for details).
+      \item RPKMs from publicly available result files were used for
+        samples where reads were not present in SRA.
+   \end{itemize}
 
-    The \ac{SVM} was generated using caret
-    \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
-    \cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a
-    tune length of 8 with a training set representing $\approx 75\%$ of the
-    samples ($\approx 2450$ samples total, including $\approx 35$
-    uterine or $\approx 60$ placenta) in each of the groups chosen at
-    random, with the remaining samples making up a testing set
-    ($\approx 820$ samples).
+\end{frame}
 
-}
+\section{Genes Characteristic of Uterus}
 
-\column{0.25}
-  \block{Genes Characteristic of Uterus}{
-
-{\centering
 % <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
@@
 
+\begin{frame}{Genes Characteristic of Uterus}
+\begin{columns}
+\column{0.6\textwidth}
+
 <<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
@@
 @
-}
+\column{0.4\textwidth}
+
 \captionof{figure}{Expression of genes characteristic of uterus
@@
   are highly characteristic of uterus.}
+\end{columns}
+\end{frame}
 
-}
+\section{Genes Characteristic of Placenta}
 
-  \block{Genes Characteristic of Placenta}{
-{\centering
+\begin{frame}{Genes Characteristic of Placenta}
 % <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
@@
 
+\begin{columns}
+\column{0.6\textwidth}
 <<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
@@
 
-}
+\column{0.4\textwidth}
 \captionof{figure}{Expression of genes characteristic of placenta
@@
   mir-622, and PAPPA-AS1 are less so.}
-}
+\end{columns}
+\end{frame}
 
-\column{0.25}
+\section{SVM to distinguish tissues}
 
-\block{SVM to distinguish tissues}{
   % Describe SVM approach
@@
 
-<<placenta.roc.curve,fig.width=8,fig.height=4,echo=FALSE>>=
+\begin{frame}{Generating SVM}
+  The \ac{SVM} was generated using caret
+    \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
+    \cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a
+    tune length of 8 with a training set representing $\approx 75\%$ of the
+    samples ($\approx 2450$ samples total, including $\approx 35$
+    uterine or $\approx 60$ placenta) in each of the groups chosen at
+    random, with the remaining samples making up a testing set
+    ($\approx 820$ samples).
+  \end{frame}
+
+\begin{frame}{Placenta and Uterus SVM}
+  \centering
+<<placenta.roc.curve,fig.width=8,fig.height=4,out.height="0.6\\textheight",echo=FALSE>>=
 layout(matrix(1:2,1,2))
@@
 \label{fig:plac_ut_full_roc}
-}
+\end{frame}
 
-\block{Identification of tissue in reduced reads}{
-  % Describe reduced read approach
-  % 
-<<placenta.roc.reduced.curve,fig.width=8,fig.height=10,echo=FALSE>>=
-layout(matrix(1:2,2,1))
+\subsection{Identification of tissue in reduced reads}
+
+\begin{frame}{Reducing Reads}
+\end{frame}
+
+\begin{frame}{Identification of tissue in reduced reads}
+  \centering
+<<placenta.roc.reduced.curve,fig.width=8,fig.height=4,out.height="0.6\\textheight",echo=FALSE>>=
+layout(matrix(1:2,1,2))
 for (a in c("placenta","uterus")) {
@@
   placenta.}
-}
 
-\column{0.25}
+\end{frame}
 
-\block{Conclusions}{
+\section{Conclusions}
+
+\begin{frame}{Conclusions}
   \begin{itemize}
@@
   \end{itemize}
-}
+\end{frame}
 
-\block{References}{
+\section*{Acknowledgements}
 
-  \begin{center}
-    \mbox{}\vspace{-\baselineskip}
-    \printbibliography[heading=none]
-  \end{center}
-}
-
-\block{Code and Datasets}{
+\begin{frame}
   The code and underlying data for this poster can be found at
@@
 \end{center}
-% 
-}
-\block{}{
-  \includegraphics[width=\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
-}
+\end{frame}
 
-\end{columns}
+\section*{References}
+
+\begin{frame}[plain]{References}
+  \begin{center}
+    \mbox{}\vspace{-\baselineskip}
+    \printbibliography[heading=none]
+  \end{center}
+\end{frame}
+
+\begin{frame}{Acronyms}
+  \acrodef{EV}[EV]{Extracellular Vesicle}
+\acrodef{SVM}{Support Vector Machine}
+\acrodef{ROC}{Receiver Operating Characteristic}
+\acrodef{KNN}{K Nearest Neighbors}
+\acrodef{SRA}{Sequence Read Archive}
+
+\end{frame}
 

### 2015-09-25 slides/slides.Rnw
finalize slides

@@
 
-\title{Identifying Extracellular Vesicles of Placenta and Uterine 
+\title[EVs of Placenta and Uterus]{Identifying Extracellular Vesicles of Placenta and Uterine 
   Origin Using RNA Expression Signatures}
-\author{Don L. Armstrong, Monica Uddin, Derek Wildman}
-\institute{Institute for Genomic Biology, Computing Genomes 
+\author[Don Armstrong et. al.]{Don L. Armstrong, Monica Uddin, Derek Wildman}
+\institute[IGB]{Institute for Genomic Biology, Computing Genomes 
   for Reproductive Health, University of Illinois, Urbana-Champaign}
@@
 \begin{frame}{Identifying source tissue of \acp{EV}}
+  \begin{enumerate}
+  \item We want to remotely sense the transcriptome of tissues
+  \item For each \ac{EV}, identify the source tissue
 
-  One requirement for remotely sensing the transcriptome is
-  identifying the source tissue of a specific \ac{EV}. To identify the
-  source tissue of a specific \ac{EV}, we have identified miRNA and
-  mRNA markers which are characteristic of multiple organs and tissues
-  from existing publicly-available RNAseq transcriptomes by
-  calculating a tissue specificity index.
+  \item So, identify miRNA and mRNA markers which are characteristic of
+    multiple organs and tissues
+    \begin{itemize}    
+    \item Use publicly-available RNAseq transcriptomes
 
-  The tissue specificity index for gene $g$ ($\tau_g$) is
+    \item Identify characteristic markers by calculating a tissue
+      specificity index for each gene $g$ ($\tau_g$):
 
@@
       \label{eq:tissue_specificity_index}
-
-  where $N$ is the number of tissues, $X_{i,g}$ is the mean expression
+ \item $N$ is the number of tissues, $X_{i,g}$ is the mean expression
    of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
-  expression of gene $g$ across all tissues.
+   expression of gene $g$
+    \end{itemize}
+  \end{enumerate}
 
@@
 
- \begin{frame}{\ac{SVM}}
-    We also trained a multiclass \ac{SVM} to accurately identify the
-    tissue of origin, and present the results of the identification of
-    reads from uterine and placental origin here. Tables showing
-    tissue specific genes from all %
-    %
-    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
-    % 
-    tissues can be found on \url{http://dla2.us/p/em2015}.
-
-%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
-%     uterus specific. CGB, LGALS13, and members of the PSG family were
-%     identified as placenta specific. The accuracy of the classifier
-%     was verified using RNAseq transcriptome data not in the training
-%     set. We show the applicability of the classifier to simulated EVs
-%     by down sampling reads to $10^5$ reads and biasing the reads to
-%     account for preferential association of RNA motifs with EV
-%     packaging machinery. When coupled with single-EV sequencing, our
-%     classifier enables clinicians to monitor disease-specific markers
-%     over the course of treatment without biopsies, increasing the
-%     ability of clinicians to adapt treatment modalities to the
-%     individual patient. 
+\begin{frame}{Source of \acp{EV} --- Machine Learning}
+  \begin{itemize}
+  \item High tissue specificity might not be the best discriminator
+  \item Use machine learning techniques!
+  \item Train a \ac{SVM}
+  \item Use genes which are highly expressed and have either
+    \begin{itemize}
+    \item High Tissue Specific Index ($\tau_g$)
+    \item High Entropy
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
     \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq}
-  \item called using Cufflinks 2 v2.0.10
+  \item quantified using Cufflinks 2 v2.0.10
     \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}.
@@
 \begin{frame}{Generating SVM}
-  The \ac{SVM} was generated using caret
+ \begin{itemize}
+ \item \ac{SVM} was generated using caret
     \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
-    \cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a
-    tune length of 8 with a training set representing $\approx 75\%$ of the
-    samples ($\approx 2450$ samples total, including $\approx 35$
-    uterine or $\approx 60$ placenta) in each of the groups chosen at
-    random, with the remaining samples making up a testing set
-    ($\approx 820$ samples).
+    \cite{Karatzoglou2004:kernlab}
+  \item 10 cycles of 10 repeats and a
+    tune length of 8
+
+  \item training set representing $\approx 75\%$ of the samples
+    ($\approx 2450$ samples total, including $\approx 35$ uterine or
+    $\approx 60$ placenta) in each of the groups chosen at random,
+  \item remaining samples making up a testing set ($\approx 820$
+    samples)
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Reducing Reads}
+  \begin{itemize}
+  \item Reads sub-sampled from each sample 
+  \item Weighted by the FPKM measured
+  \item Run using the same trained SVM shown previously
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Identification of tissue in reduced reads}
-  \centering
-<<placenta.roc.reduced.curve,fig.width=8,fig.height=4,out.height="0.6\\textheight",echo=FALSE>>=
-layout(matrix(1:2,1,2))
-for (a in c("placenta","uterus")) {
+
+\begin{columns}
+\column{0.6\textwidth}
+<<placenta.roc.reduced.curve,fig.width=6,fig.height=6,out.height="0.8\\textheight",echo=FALSE>>=
+#layout(matrix(1:2,1,2))
+for (a in c("placenta")) {
 pl.roc.obj1 <- 
@@
 @ 
-\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
-  Uterus samples with the indicated number of total reads. Reads were
-  reduced by sampling reads for each sample weighted by the FPKM
-  measured for each sample, and run using the same trained SVM
-  depicted in Figure \ref{fig:plac_ut_full_roc}. %
+\column{0.4\textwidth}
+\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta with the indicated number of total reads.  %
   % 
@@
   samples of placental origin and non-placental origin even when the
-  number of reads is very low (10). The uterus \ac{SVM} requires many
-  more reads in order to achieve similar accuracy. This is likely due
-  to the much smaller number of uterus specific genes in comparison to
-  placenta.}
+  number of reads is very low (10).}
+\end{columns}
+\end{frame}
 
+\begin{frame}{Reduced Reads: Uterus}
+\begin{columns}
+\column{0.6\textwidth}
+<<uterus.roc.reduced.curve,fig.width=6,fig.height=6,out.height="0.8\\textheight",echo=FALSE>>=
+# layout(matrix(1:2,1,2))
+for (a in c("uterus")) {
+pl.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#002058"
+             )
+pl.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
+              percent=TRUE,
+              col="#E87722")
+pl.roc.obj3 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
+              percent=TRUE,
+              col="green")
+pl.roc.obj4 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
+              percent=TRUE,
+              col="brown")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
+       col=c("brown","green","#002058", "#E87722"),
+       lwd=2)
+}
+@ 
+\column{0.4\textwidth}
+\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
+  Uterus samples with the indicated number of total reads.  %
+  % 
+  The uterus \ac{SVM} requires many more reads in order to achieve
+  similar accuracy. This is likely due to the much smaller number of
+  uterus specific genes in comparison to placenta.}
+\end{columns}
 \end{frame}
@@
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 
-\begin{frame}
-  The code and underlying data for this poster can be found at
+\begin{frame}{Acknowledgments}
+  \begin{itemize}
+  \item Derek Wildman
+  \item Monica Uddin
+  \end{itemize}
+
+  The code and underlying data for this talk can be found at
   \begin{center}
@@
 
-\begin{frame}{Acronyms}
+\begin{frame}[plain]
   \acrodef{EV}[EV]{Extracellular Vesicle}

### 2015-10-12 slides/slides.Rnw
add better slides for ISRS presentation; fix date

@@
 \usepackage{qrcode}
+\usepackage[absolute,overlay]{textpos}
 \usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
@@
 \usepackage{tikz}
+\usetikzlibrary{positioning}
+\usetikzlibrary{arrows.meta}
 \usepackage{nameref}
@@
 \renewcommand*{\bibfont}{\tiny}
+\usepackage{mathtools,cancel}
+
+\renewcommand{\CancelColor}{\color{red}} %change cancel color to red
+
+\makeatletter
+\let\my@cancelto\cancelto %copy over the original cancelto command
+\newcommand<>{\cancelto}[2]{\alt#3{\my@cancelto{#1}{#2}}{\mathrlap{#2}\phantom{\my@cancelto{#1}{#2}}}}
+% redefine the cancelto command, using \phantom to assure that the
+% result doesn't wiggle up and down with and without the arrow
+\makeatother
 
@@
   \usetheme{CambridgeUS}
-  % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
+  \usecolortheme{crane}
+%   % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
   \definecolor{ilboldblue}{HTML}{002058}
@@
   \definecolor{ilorange}{HTML}{D45D00}
-  \setbeamercolor{alerted text}{fg=ilboldorange}
-  \setbeamercolor*{palette primary}{fg=ilblue,bg=ilorange}
-  \setbeamercolor*{palette secondary}{fg=ilblue!20!white,bg=ilorange}
-  \setbeamercolor*{palette tertiary}{bg=ilblue,fg=ilorange}
-  \setbeamercolor*{palette quaternary}{fg=ilblue,bg=ilorange}
-
-  \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}
-
-  \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
-  \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
-  \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
-  \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}
-
-  % \setbeamercolor*{titlelike}{parent=palette primary}
-  \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
-  \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!80!white}
-  \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}
-
-  \setbeamercolor*{separation line}{}
-  \setbeamercolor*{fine separation line}{}
-  \setbeamercovered{transparent}  
+%   \setbeamercolor{alerted text}{fg=ilboldorange}
+%   \setbeamercolor*{palette primary}{fg=ilboldblue,bg=ilorange!70!white}
+%   \setbeamercolor*{palette secondary}{fg=ilboldblue,bg=ilorange!60!white}
+%   \setbeamercolor*{palette tertiary}{bg=ilblue!60!white,fg=ilboldorange}
+%   \setbeamercolor*{palette quaternary}{fg=ilboldblue,bg=ilorange!80!white}
+%   
+%   \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}
+%   
+%   \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
+%   \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
+%   \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
+%   \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}
+%   
+%   % \setbeamercolor*{titlelike}{parent=palette primary}
+%   \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
+%   \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!20!white}
+%   \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}
+%   
+%   \setbeamercolor*{separation line}{}
+%   \setbeamercolor*{fine separation line}{}
+%   \setbeamercovered{transparent}  
   \logo{\begin{tikzpicture}% Pale figure
-      {\node[opacity=0.8]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
+      {\node[opacity=0.734]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
         };}%
@@
 \subject{RNA expression}
-
+\date{October 12, 2015}
 \begin{document}
@@
 library("pROC")
+library("sampling")
 
@@
 \end{enumerate}
-
+\only<2-4>{
+\begin{textblock*}{64mm}(32mm,0.25\textheight)
+\begin{block}{Tissue Specificity}
+  \begin{itemize}
+    \item Highly specific $\tau_g = \frac{\cancelto<3->{N-1}{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}}{N-1} = 1$
+    \item Not specific $\tau_g = \frac{\cancelto<4->{0}{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}}{N-1} = 0$
+    \end{itemize}
+\end{block}
+\end{textblock*}
+}
 \end{frame}
@@
 \begin{frame}{Methods}
+  \begin{center}
+  \begin{tikzpicture}
+    [rawreads/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilblue!40!white,inner sep=2pt,minimum size=8mm,thick},
+    processing/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilboldblue!40!white,inner sep=2pt,minimum size=8mm,thick},
+    arrow/.style={->,shorten >=1pt,>=stealth,thick},
+    tissues/.style={rectangle,rounded corners=2mm,draw=ilorange!90!white,fill=ilorange!40!white,inner sep=2pt,minimum size=8mm,thick}]
+    \node[tissues] (tissues) {\Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} tissues};
+    \node[processing] (star) [above left=of tissues] {STAR+Cufflinks};
+    \node[processing] (fpkm) [above right=of tissues] {FPKMs published};
+    \node[rawreads] (samples) [above=of star] {\Sexpr{length(star.logs[grepl("Number of input reads",field),value])} samples in SRA};
+    \node[rawreads] (samples noraw) [above=of fpkm] {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
+        nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
+        length(star.logs[grepl("Number of input reads",field),value])} samples not in SRA};
+    \node[processing] (matrix) [below=of tissues] {FPKMs of 63967 genes $\times$ \Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
+        nrow(predictions.caret.svm[["uterus"]][["testing"]])} samples};
+    \draw[arrow] (samples)--(star);
+    \draw[arrow] (star)--(tissues);
+    \draw[arrow] (samples noraw)--(fpkm);
+    \draw[arrow] (fpkm)--(tissues);
+    \draw[arrow] (tissues)--(matrix);
+  \end{tikzpicture}
+\end{center}
+\end{frame}
+
+\begin{frame}<beamer:0>{Methods}
   \begin{itemize}
@@
 \begin{columns}
-\column{0.6\textwidth}
+\column{0.7\textwidth}
 
@@
 @
-\column{0.4\textwidth}
+\column{0.3\textwidth}
 
 \captionof{figure}{Expression of genes characteristic of uterus
-  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
-  highest expression for any of the uterus-characterizing genes
-  ordered by their tissue specificity. The known uterus-characterizing
-  gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp)
-  are highly characteristic of uterus.}
+  ($\tau_g \ge 0.98$) in uterus and tissues with second-highest
+  expression. The known uterus-characterizing gene
+  progestagen-associated endometrial protein (PAEP) and short RNAs
+  ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly characteristic
+  of uterus.}
 \end{columns}
@@
   ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
-  in placenta and all tissues with the second highest expression for
-  any placenta-characterizing gene ordered by their specificity to
-  placenta. The previously known placenta-characterizing PSG genes and
-  CSH genes are nearly unique to placenta and the short RNAs mir-126,
-  mir-622, and PAPPA-AS1 are less so.}
+  in placenta and tissues with the second highest expression. The
+  previously known placenta-characterizing PSG genes and CSH genes are
+  nearly unique to placenta and the short RNAs mir-126, mir-622, and
+  PAPPA-AS1 are less so.}
 \end{columns}
@@
 
+\subsection{SVM Methodology}
+
 \begin{frame}{Generating SVM}
  \begin{itemize}
+ \item \ac{SVM} is a machine learning technique which finds the
+   hyperplane with the maximal margins which separates two groups
  \item \ac{SVM} was generated using caret
@@
     tune length of 8
-
-  \item training set representing $\approx 75\%$ of the samples
-    ($\approx 2450$ samples total, including $\approx 35$ uterine or
-    $\approx 60$ placenta) in each of the groups chosen at random,
-  \item remaining samples making up a testing set ($\approx 820$
-    samples)
   \end{itemize}
@@
 
-\begin{frame}{Placenta and Uterus SVM}
-  \centering
-<<placenta.roc.curve,fig.width=8,fig.height=4,out.height="0.6\\textheight",echo=FALSE>>=
-layout(matrix(1:2,1,2))
+\begin{frame}{Testing and Training Sets}
+  \begin{center}
+  \begin{tikzpicture}
+    [training/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilblue!40!white,inner sep=2pt,minimum size=8mm,thick},
+    testing/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilboldblue!40!white,inner sep=2pt,minimum size=8mm,thick},
+    arrow/.style={->,shorten >=1pt,>=stealth,thick},
+    sample/.style={rectangle,rounded corners=2mm,draw=ilorange!90!white,fill=ilorange!40!white,inner sep=2pt,minimum size=8mm,thick}]
+    \node[sample] (samples) {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
+        nrow(predictions.caret.svm[["uterus"]][["testing"]])} samples};
+\onslide<2->{    \node[training] (training) [below left=of samples] {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])} training samples};
+    \node[training] (training uterus) [below=of training] {\Sexpr{sum(predictions.caret.svm[["uterus"]][["training"]][,"actual"]=="uterus")} uterus,
+      \Sexpr{sum(predictions.caret.svm[["placenta"]][["training"]][,"actual"]=="placenta")} placenta};
+
+    \draw[arrow] (samples) -- node[above left=8pt]{75\%} (training);
+    \draw[arrow] (training) -- (training uterus);
+    }
+\onslide<3->{    \node[testing] (testing) [below right=of samples] {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["testing"]])} testing samples};
+    \node[testing] (testing uterus) [below=of testing] {\Sexpr{sum(predictions.caret.svm[["uterus"]][["testing"]][,"actual"]=="uterus")} uterus,
+      \Sexpr{sum(predictions.caret.svm[["placenta"]][["testing"]][,"actual"]=="placenta")} placenta};
+    \draw[arrow] (samples) -- node[above right=8pt]{25\%} (testing);
+    \draw[arrow] (testing) -- (testing uterus);}
+  \end{tikzpicture}
+\end{center}
+\end{frame}
+
+\begin{frame}{Placenta SVM: Training and Testing}
+  \begin{columns}
+    \column{0.6\textwidth}
+<<placenta.roc.curve,fig.width=4,fig.height=4,out.height="0.7\\textheight",echo=FALSE>>=
+# layout(matrix(1:2,1,2))
+par(mar=c(0,0,0,0))
 pl.roc.obj1 <- 
@@
        lwd=2)
+@ 
+\column{0.4\textwidth}
+\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM}. The
+  \ac{SVM} is highly effective at distinguishing between samples of
+  placental origin and non-placental origin.}
+\label{fig:plac_full_roc}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Uterus SVM: Training and Testing Sets}
+  \begin{columns}
+    \column{0.6\textwidth}
+<<uterus.roc.curve,fig.width=4,fig.height=4,out.height="0.7\\textheight",echo=FALSE>>=
 
@@
 @ 
-\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM} (left) and
-  uterus \ac{SVM} right) . Both \ac{SVM} are highly effective at
-  distinguishing between samples of uterus origin and non-uterus
-  origin.}
-\label{fig:plac_ut_full_roc}
+\column{0.4\textwidth}
+\captionof{figure}{\ac{ROC} curve of the uterus \ac{SVM}. The \ac{SVM}
+  is highly effective at distinguishing between samples of uterus
+  origin and non-uterus origin.}
+\label{fig:ut_full_roc}
+\end{columns}
 \end{frame}
@@
   \end{itemize}
+  \only<2->{
+\begin{textblock*}{100mm}(20mm,0.25\textheight)
+\begin{block}{Sub sampling}
+<<calc.reduced,echo=FALSE,results="asis">>=
+reduced.reads <- function(fpkms,num.reads) {
+    non.zero <- fpkms > 0
+    ret.val <- rep.int(0,length(fpkms))
+    if (!any(non.zero)) {
+        return(fpkms)
+    }
+    inc.prob <- inclusionprobabilities(fpkms[non.zero],1)
+    ret.val[non.zero] <-
+        apply(sapply(1:(num.reads/1),function(y){UPrandomsystematic(inc.prob,0)}),1,sum)
+    names(ret.val) <- names(fpkms)
+    ret.val
+}
+set.seed(1021)
+reads <- c("A"=1000,
+           "B"=2000,
+           "C"=20,
+           "D"=5000)
+reads.to.table <- function(read,reads){
+    data.table("Total Reads"=read,
+               "Gene A"=reads["A"],
+               "Gene B"=reads["B"],
+               "Gene C"=reads["C"],
+               "Gene D"=reads["D"])
+}
+rr.table <- list(reads.to.table("Orig. FPKM",reads))
+i <- 0
+for (read in c(500,500,100,100,5,5,5)){
+    i <- i+1
+    reads <- reduced.reads(reads,read)
+    rr.table[[paste0("read.",read,'.',i)]] <- 
+        reads.to.table(read,reads)
+}
+print(xtable(rbindlist(rr.table),
+             digits=0
+             ),
+      include.rownames=FALSE)
+@
+
+\end{block}
+\end{textblock*}
+}
+
 \end{frame}
@@
   \end{center}
+  \begin{itemize}
+  \item University of Illinois and Institute for Genomic Biology
+    Startup Funds to Derek and Monica
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[plain]
+  \acrodef{EV}[EV]{Extracellular Vesicle}
+  \acrodef{SVM}{Support Vector Machine}
+  \acrodef{ROC}{Receiver Operating Characteristic}
+  \acrodef{KNN}{K Nearest Neighbors}
+  \acrodef{SRA}{Sequence Read Archive}
+
 \end{frame}
@@
 
-\begin{frame}[plain]
-  \acrodef{EV}[EV]{Extracellular Vesicle}
-\acrodef{SVM}{Support Vector Machine}
-\acrodef{ROC}{Receiver Operating Characteristic}
-\acrodef{KNN}{K Nearest Neighbors}
-\acrodef{SRA}{Sequence Read Archive}
-
-\end{frame}
-
 \end{document}

### 2016-01-14 manuscript/manuscript.Rnw
begin work on manuscript

@@
 \addbibresource{references.bib}
-%\usepackage[noblocks]{authblk}
+\usepackage[noblocks,auth-sc]{authblk}
 \usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
@@
 \usepackage{nameref}
+\usepackage{xspace}
+% this package will stretch the font slightly if necessary.
+\usepackage{microtype}
 \usepackage{zref-xr,zref-user}
@@
 \oddsidemargin 0.0in 
-\textwidth 6.5in
+\textwidth 6in
 \raggedbottom
@@
 \pagestyle{fancy}
-\author{AUTHOR}
-\title{TITLE}
-\date{\ }
-\onehalfspacing
 
 \begin{document}
+\author[1]{Don Armstrong}
+\author[1,2]{Monica Uddin}
+\author[1,3,*]{Derek E. Wildman}
+\title{Human tissue specific markers}
+\affil[1]{Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
+\affil[2]{Department of Psychology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
+\affil[3]{Department of Molecular and Integrative Physiology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
+\affil[*]{To whom correspondence should be addressed}
+\date{\today}
+
+\input{acronyms}
+
+\maketitle
+
+\doublespacing
 
@@
 })
+options(scipen = -1, digits = 2)
 library("data.table")
@@
 library("xtable")
+library("caret")
+library("pROC")
+library("grid")
+
+capwords <- function(s, strict = FALSE) {
+    cap <- function(s) paste(toupper(substring(s, 1, 1)),
+                             {s <- substring(s, 2); if(strict) tolower(s) else s},
+                             sep = "", collapse = " " )
+    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
+}
+
 @
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum("../tissue_specific_expression/tissue_specific_markers")>>=
-load("../tissue_specific_expression/tissue_specific_markers")
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
+for (a in paste0("../tissue_specific_expression/",
+                 c("tissue_specific_markers",
+                   "interesting_gene_reads_wide",
+                   "combined_star_logs",
+                   "svm_test_results",
+                   "svm_test_results_reduced"
+                   ))) {
+    load(a)
+}
 tissue.specific.genes <- data.table(tissue.specific.genes)
@@
 
-\section{Tissue Specific Markers}
+<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
+specific.markers.plot <- function(tissue.name,genes) {
+    setkey(interesting.gene.reads.wide,"gene_short_name")
+    if (missing(genes))
+        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
+    interesting.tissues <- 
+        colnames(interesting.gene.reads.wide)[
+         apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
+                             function(x){order(-x)[2]+2})]
+
+    interesting.tissues <- 
+        c("gene_short_name",tissue.name,unique(interesting.tissues))
+
+    tissue.specific.gene.reads <- 
+        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
+                        id="gene_short_name",
+                        variable.name="tissue",
+                        value.name="FPKM"
+                        ))
+    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
+    gene.levels <- 
+        unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
+    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
+    return(ggplot(tissue.specific.gene.reads,
+                  aes(y=gene_short_name,
+                          x=tissue))
+           + geom_tile(aes(fill=`log_2(FPKM)`))
+           + scale_fill_gradient(limits=c(0,17),
+                                 low="white",
+                                 high="#002058",
+                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
+           + scale_x_discrete("", expand = c(0, 0))
+           + scale_y_discrete("", expand = c(0, 0))
+           + theme_grey(base_size=14)
+           + theme(legend.position = "top",
+                   axis.ticks = element_blank(),
+                   panel.grid=element_blank(),
+                   panel.border=element_blank(),
+                   panel.background=element_blank(),
+                   plot.background=element_blank(),
+                   plot.margin=unit(c(1,1,1,1),"mm"),
+                   axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
+           )
+}
+@
+
+\section{Abstract}
+
+\section{Introduction}
+
+\acp{EV} are small (40-1000nm) compartments that can be found in
+blood, saliva, and urine. \acp{EV} originate from the plasma membrane
+or multivesicular endosome of many cell types and contain proteins,
+lipids, mRNA, miRNA, and many other non-coding RNAs. Because \acp{EV}
+contain mRNA from the cell of origin and can be obtained without
+invasive procedures, they are a promising remote sensor of the
+transcriptome of tissues that would otherwise be inaccessible.
+
+One requirement for remotely sensing the transcriptome is identifying
+the source tissue of a specific \ac{EV}. To identify the source tissue
+of a specific \ac{EV}, we have identified miRNA and mRNA markers which
+are characteristic of multiple organs and tissues from existing
+publicly-available RNAseq transcriptomes by calculating a tissue
+specificity index.
+
+The tissue specificity index for gene $g$ ($\tau_g$) is
+
+\begin{equation}
+  \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
+\end{equation}
+\label{eq:tissue_specificity_index}
+
+where $N$ is the number of tissues, $X_{i,g}$ is the mean expression
+of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
+expression of gene $g$ across all tissues.
+
+We also trained a multiclass \ac{SVM} to accurately identify the
+tissue of origin, and present the results of the identification of
+reads from uterine and placental origin here. Tables showing tissue
+specific genes from all
+\Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))}
+tissues can be found on \url{http://dla2.us/p/em2015}.
+
+\section{Methods}
+
+\Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
+%
+RNAseq reads from %
+%
+\Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
+% 
+samples (supplemented with FPKMs from
+\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
+  nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
+  length(star.logs[grepl("Number of input reads",field),value])} %
+%
+additional samples without raw data in the \ac{SRA}) corresponding
+to %
+\Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
+different tissues were obtained from the Roadmap Epigenomics
+\cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx
+\cite{Gtexconsortium.2015:Thegenotypetissueexpression} projects, as
+well as uterine (GSE50599,
+\cite{Chan.ea2014:Assessmentofmyometrialtranscriptome}) and placenta
+(GSE66622, \cite{Hughes.ea2015:Evaluatingintra-andinter-individual})
+samples from the \ac{SRA}.
+
+Reads were aligned to \textit{Homo sapiens} reference genome GRCh38
+with Ensembl annotation release 80 using STAR v2.4.2a
+\cite{Dobin.ea2013:STARultrafastuniversalRNA-seq} and called using
+Tophat 2 v2.0.10
+\cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}. The average
+percentage of uniquely mapped reads was %
+\Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}\%
+and the average total mapping percentage was %
+\Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}\%.
+% 
+For samples where the reads were not publicly available (GTEx and
+some Roadmap Epigenomics samples), the RPKM were taken from publicly
+available result files (see code for details).
+
+The \ac{SVM} was generated using caret
+\cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
+\cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a tune
+length of 8 with a training set representing $\approx 75\%$ of the
+samples ($\approx 2450$ samples total, including $\approx 35$ uterine
+or $\approx 60$ placenta) in each of the groups chosen at random, with
+the remaining samples making up a testing set ($\approx 820$ samples).
+
+\section{Results}
 
@@
 
-FTX_1 is very interseting
+\begin{figure}
+<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+print(specific.markers.plot("uterus")
+      + xlab("Genes Characteristic of Uterus ")
+      + ylab("Tissue"))
+@
+\caption{Expression of genes characteristic of uterus
+  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
+  highest expression for any of the uterus-characterizing genes
+  ordered by their tissue specificity. The known uterus-characterizing
+  gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp)
+  are highly characteristic of uterus.}
+\label{fig:uterus_specific_markers}
+\end{figure}
+
+FTX\_1 is very interseting
 
@@
 
+\begin{figure}
+<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+print(specific.markers.plot("placenta",
+                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
+                                                                         gi.max >= 1024
+                               ,gene]))
+      + xlab("Genes Characteristic of Placenta")
+      + ylab("Tissue"))
+@
+
+}
+\caption{Expression of genes characteristic of placenta
+  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
+  in placenta and all tissues with the second highest expression for
+  any placenta-characterizing gene ordered by their specificity to
+  placenta. The previously known placenta-characterizing PSG genes and
+  CSH genes are nearly unique to placenta and the short RNAs mir-126,
+  mir-622, and PAPPA-AS1 are less so.}
+\label{fig:placenta_specific_markers}
+\end{figure}
+
+\subsection{Identification of Tissue source of Exosomes}
+
+\begin{figure}
+<<placenta.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
+layout(matrix(c(1,3,2,4),2,2))
+pl.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"placenta"]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#002058"
+             )
+pl.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"placenta"]),
+              percent=TRUE,
+              col="#E87722")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",
+       legend=c("Placenta Testing Set", "Placenta Training Set"),
+       col=c("#002058", "#E87722"),
+       lwd=2)
+
+ut.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"uterus"]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#002058"
+             )
+ut.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"uterus"]),
+              percent=TRUE,
+              col="#E87722")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",
+       legend=c("Uterus Testing Set", "Uterus Training Set"),
+       col=c("#002058", "#E87722"),
+       lwd=2)
+
+for (a in c("placenta","uterus")) {
+pl.roc.obj1 <- 
+    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
+             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
+             percent=TRUE,
+             ## main="Training vs Testing set for Uterus",
+             col="#002058"
+             )
+pl.roc.obj2 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
+              percent=TRUE,
+              col="#E87722")
+pl.roc.obj3 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
+              percent=TRUE,
+              col="green")
+pl.roc.obj4 <- 
+    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
+              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
+              percent=TRUE,
+              col="brown")
+# testobj <- roc.test(rocobj1, rocobj2)
+# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))
+
+legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
+       col=c("brown","green","#002058", "#E87722"),
+       lwd=2)
+}
+grid.text(c("A","B","C","D"),
+          x=unit(c(0,0.5,0,0.5),"npc"),
+          y=unit(c(1,1,0.5,0.5),"npc"),
+          just=c("left","top"),
+          gp=gpar(fontsize=32)
+          )
+
+@
+
+\caption{\ac{ROC} curve of the placenta \ac{SVM} (A) and
+  uterus \ac{SVM} (B) . Both \ac{SVM} are highly effective at
+  distinguishing between samples of uterus origin and non-uterus
+  origin.
+  %
+  \ac{ROC} curve of reduced reads in Placenta and
+  Uterus samples with the indicated number of total reads. Reads were
+  reduced by sampling reads for each sample weighted by the FPKM
+  measured for each sample, and run using the same trained SVM
+  depicted in Figure \ref{fig:plac_ut_full_roc}. %
+  % 
+  The placenta \ac{SVM} is fairly accurate at distinguishing between
+  samples of placental origin and non-placental origin even when the
+  number of reads is very low (10). The uterus \ac{SVM} requires many
+  more reads in order to achieve similar accuracy. This is likely due
+  to the much smaller number of uterus specific genes in comparison to
+  placenta.}
+\label{fig:plac_ut_roc}
+\end{figure}
+
+\begin{figure}
+\label{fig:plac_ut_red_roc}
+\end{figure}
+
+\section{Discussion}
+
+\section{Conclusion}
+
+  \begin{itemize}
+  \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
+    characteristic of placental tissues, and PAEP is characteristic of
+    uterus, as previously described.
+%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
+  \item \ac{SVM} are able to accurately identify the tissue of origin
+    for uterus and placenta.
+  \item Very few sequenced reads are required to accurately identify
+    tissues of Placental origin. We are able to detect 100\% of the
+    placenta tissues with 80\% specificity with just 10 reads.
+  \item We can easily identify exosomes originating from the placenta,
+    as even the smallest exosomes can contain 70 RNA molecules
+    \cite{Li.ea2014:AnalysisofRNAcontent}.
+  \item Non-maternal variants may also be able to distinguish between
+    placental reads and maternal reads.
+  \end{itemize}
+
+\printbibliography
+
 \end{document}

### 2016-01-14 slides/slides.Rnw
add tissue specificity shape plot

--- 2015-09-25 slides/slides.Rnw
@@
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))>>=
 for (a in paste0("../tissue_specific_expression/",
@@
                    "svm_test_results",
-                   "svm_test_results_reduced"
+                   "svm_test_results_reduced",
+                   "housekeeping_genes_superset"
                    ))) {
@@
 \end{columns}
+\end{frame}
+
+\subsection{Tissue specificity shape}
+
+\begin{frame}
+<<tissue.specificity,shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+tissue.specific.genes[,type:="not specific"]
+tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"]
+tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
+                          possibly.specific==FALSE,type:="housekeeping"]
+tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
+                          possibly.specific==TRUE,type:="housekeeping and specific"]
+
+tissue.specific.genes[,type:=
+                          factor(type,
+                                 levels=c("not specific","tissue specific",
+                                     "housekeeping",
+                                     "housekeeping and specific"
+                                          ))]
+ggplot(tissue.specific.genes[gi.max>0],
+       aes(y=specificity,x=gi.max,color=type))+
+    geom_hex(bins=100)+scale_x_log10()+
+        xlab("Max Expression (FPKM)")+
+            ylab("Tissue Specificity")
+@ 
 \end{frame}

### 2016-01-14 tissue_specific_expression/R/housekeeping_genes.R
modify housekeeping genes to have a specific sd

@@ -48,9 +48,15 @@ housekeeping.gene.levels <- function(gene) {
         gene=gene,
         tracking_id=gene.reads.wide[gene,tracking_id][1],
         mean=mean(gene.expression),
+        sd=sd(gene.expression),
         percentage=mean(gene.expression>50),
+        log.sd=sd(log(gene.expression[gene.expression>0])),
+        log.mean=mean(log(gene.expression[gene.expression>0])),
         percentage.10=mean(gene.expression>10),
-        percentage.1=mean(gene.expression>1)))
+        sd.10=sd(gene.expression[gene.expression>10]),
+        percentage.1=mean(gene.expression>1),
+        sd.1=sd(gene.expression[gene.expression>1])
+        ))
 }
 
 housekeeping.genes <-
@@ -61,6 +67,7 @@ housekeeping.genes <-
 housekeeping.genes <-
     rbindlist(housekeeping.genes)
 
+invisible(housekeeping.genes[,scaled.sd:=log.sd/log.mean])
 
 ### identify house keeping genes. These are genes which are expressed
 ### in pretty much every tissue. [Note that we'll have to be careful
@@ -69,11 +76,11 @@ housekeeping.genes <-
 ### We're using 60% here primarily to make sure that we get MT genes
 ### which in some cases have been excluded from alignments.
 
-### genes expressed with FPKM >= 50 in >= 60% of tissues 
+### genes expressed with FPKM >= 50 in >= 80% of tissues and with a scaled SD of less than 0.12
 
 housekeeping.genes <-
-    housekeeping.genes[percentage >= 0.6 |
-                           (grepl("^MT-",gene) & mean > 100)
+    housekeeping.genes[(percentage >= 0.8 & scaled.sd < 0.12) |
+                       (grepl("^MT-",gene) & mean > 100)
                       ,]
 setnames(housekeeping.genes,"gene","gene_short_name")
 

### 2016-12-21 tissue_expression/R/combine_isoform_fpkms.R
include gtex counts

@@ -14,6 +14,8 @@ args <- args[-length(args)]
 ### these are the isoform fpkm files
 isoform.files <- grep("_isoforms.fpkm_tracking",args,value=TRUE)
 
+gtex.file <- grep("GTEx_Analysis",args,value=TRUE)
+
 .get.srx <- function(x){
     gsub("(SRX\\d+)/.+","\\1",x)
 }
@@ -25,13 +27,29 @@ isoform.counts <- list()
 for (file in isoform.files) {
     srx.accession <- .get.srx(file)
     isoform.counts[[srx.accession]] <-
-        fread(file)
+        fread(file)[,list(tracking_id,gene_id,gene_short_name,FPKM,FPKM_status)]
     isoform.counts[[srx.accession]][,srx:=srx.accession]
     i <- i + 1
     setTxtProgressBar(pb,i)
 }
+gc()
 
-isoform.counts <- rbindlist(isoform.counts)
+gtex.counts <- fread(paste0("zcat ",gtex.file),skip="TargetID")
+
+gtex.counts[,TargetID:=gsub("\\.\\d+","",TargetID)]
+gtex.counts[,Gene_Symbol:=gsub("\\.\\d+","",Gene_Symbol)]
+gtex.counts[,Chr:=NULL]
+gtex.counts[,Coord:=NULL]
+setnames(gtex.counts,c("TargetID","Gene_Symbol"),c("tracking_id","gene_id"))
+gtex.counts <-
+    melt(gtex.counts,
+         id.vars=c("tracking_id","gene_id"),
+         value.name="FPKM",
+         variable.name="srx")
+gc()
+isoform.counts[["gtex"]] <- gtex.counts
+
+isoform.counts <- rbindlist(isoform.counts,fill=TRUE)
 
 save(file=results.file,
      isoform.counts)

### 2016-12-21 manuscript/manuscript.Rnw
switch to acro from acronym
work on manuscript

@@
 \usepackage{fancyref}
-\usepackage{acronym}
+\usepackage{acro}
 \usepackage{array}
@@
 \usepackage{zref-xr,zref-user}
+\usepackage[capitalize]{cleveref}
 \newenvironment{narrow}[2]{%
@@
 \pagestyle{fancy}
+
+\newcommand{\EV}{\EV}
 
@@
 
-\acp{EV} are small (40-1000nm) compartments that can be found in
-blood, saliva, and urine. \acp{EV} originate from the plasma membrane
+\subsection{Housekeeping genes}
+
+\subsection{Cell free RNA and Exosomes in Diagnostics}
+
+\EV are small (40-1000nm) compartments that can be found in
+blood, saliva, and urine. \EV originate from the plasma membrane
 or multivesicular endosome of many cell types and contain proteins,
-lipids, mRNA, miRNA, and many other non-coding RNAs. Because \acp{EV}
+lipids, mRNA, miRNA, and many other non-coding RNAs. Because \EV
 contain mRNA from the cell of origin and can be obtained without
@@
 
-The tissue specificity index for gene $g$ ($\tau_g$) is
+\section{Methods}
 
-\begin{equation}
-  \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
-\end{equation}
-\label{eq:tissue_specificity_index}
-
-where $N$ is the number of tissues, $X_{i,g}$ is the mean expression
-of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
-expression of gene $g$ across all tissues.
-
-We also trained a multiclass \ac{SVM} to accurately identify the
-tissue of origin, and present the results of the identification of
-reads from uterine and placental origin here. Tables showing tissue
-specific genes from all
-\Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))}
-tissues can be found on \url{http://dla2.us/p/em2015}.
-
-\section{Methods}
+\subsection{Tissue RNA-seq}
 
@@
 
+\subsection{Tissue Specificity}
+
+The tissue specificity index for gene $g$ ($\tau_g$) is
+
+\begin{equation}
+  \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
+\end{equation}
+\label{eq:tissue_specificity_index}
+
+where $N$ is the number of tissues, $X_{i,g}$ is the mean expression
+of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
+expression of gene $g$ across all tissues.
+
+We also trained a multiclass \ac{SVM} to accurately identify the
+tissue of origin, and present the results of the identification of
+reads from uterine and placental origin here. Tables showing tissue
+specific genes from all
+\Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))}
+tissues can be found on \url{http://dla2.us/p/em2015}.
+
+\subsection{Read Biasing}
+
+The population of RNAs found in \EV are distinct from the cytosolic
+pool, and are likely selectively sorted into \EV~%
+\cite{Villarroya-Beltri.ea2014:Sortingitoutregulation,Hoen.ea2012:DeepsequencingofRNA,Janas.ea2015:MechanismsofRNAloading,Balkom.ea2015:Quantitativeandqualitativeanalysis}.
+Specifically, the motives
+GGAG~\cite{Villarroya-Beltri.ea2013:SumoylatedhnRNPA2B1controlssorting},
+ACCAGCCU, CAGUGAGC, and
+UAAUCCCA~\cite{Batagov.ea2013:Exosomessecretedbyhuman} are enriched in
+\EV. Owing to the limited number of tissue-specific \ac{EV} RNA-seq
+results, we have biased the reads from existing tissue-specific
+RNA-seq results to preferentially include RNAs which contain motifs
+found in exosomes and also match the lower number of unique RNAs which
+are found in individual exosomes.
+
+\subsection{Machine Learning}
+
 The \ac{SVM} was generated using caret
@@
 
-\subsection{Uterus Specific}
+\subsection{Tissue Specific Genes and Housekeeping genes}
+
+\begin{figure}
+<<tissue.specificity,shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+invisible(tissue.specific.genes[,type:="not specific"])
+invisible(tissue.specific.genes[,housekeeping:="not housekeeping"])
+invisible(tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"])
+invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
+                                 possibly.specific==FALSE,housekeeping:="housekeeping"])
+invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
+                          possibly.specific==TRUE,housekeeping:="housekeeping and specific"])
+
+invisible(tissue.specific.genes[,type:=
+                          factor(type,
+                                 levels=c("not specific","tissue specific",
+                                     "housekeeping",
+                                     "housekeeping and specific"
+                                          ))])
+print(ggplot(tissue.specific.genes[gi.max>0],
+             aes(y=specificity,x=gi.max,colour=type))+
+      ##geom_point()+
+      geom_hex(bins=50)+
+      ## stat_summary_2d(fun=function(x){length(x)},bins=100)+
+      geom_point(aes(colour=housekeeping),data=tissue.specific.genes[housekeeping!="not housekeeping",])+
+      scale_x_log10()+
+      xlab("Max Expression (FPKM)")+
+      ylab("Tissue Specificity")+
+      theme(legend.position="top")
+    )
+@ 
+\caption{Shape of tissue specific genes}
+\label{fig:tissue_specific_shape}
+\end{figure}
+
+\Cref{fig:tissue_specific_shape} shows the shape of the tissue
+specificity measure of all genes along with the maximum expression in
+a particular tissue. Almost all housekeeping genes identified by
+\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} have
+$\tau_g \le 0.93$ with the exception of
+%
+\Sexpr{tissue.specific.genes[housekeeping=="housekeeping and specific",gene]}
+%
+
+\subsection{Uterus and Placenta Specific Genes}
+
 <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
@@
 print(xtable(temp,
-             caption="Uterus specific genes",label="tab:uterus_specific_genes"),
+             caption="Uterus specific genes",
+             label="tab:uterus_specific_genes"),
       include.rownames=FALSE,
@@
 
-\subsection{Placenta Specific Genes}
-
 <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
@@
 
-\subsection{Identification of Tissue source of Exosomes}
+\subsection{Biasing reads towards exosomes}
+
+\subsection{Machine learning to identify tissue source}
 
@@
 
-\begin{figure}
-\label{fig:plac_ut_red_roc}
-\end{figure}
+<<calc.reduced,echo=FALSE,results="asis">>=
+reduced.reads <- function(fpkms,num.reads) {
+    non.zero <- fpkms > 0
+    ret.val <- rep.int(0,length(fpkms))
+    if (!any(non.zero)) {
+        return(fpkms)
+    }
+    inc.prob <- inclusionprobabilities(fpkms[non.zero],1)
+    ret.val[non.zero] <-
+        apply(sapply(1:(num.reads/1),function(y){UPrandomsystematic(inc.prob,0)}),1,sum)
+    names(ret.val) <- names(fpkms)
+    ret.val
+}
+set.seed(1021)
+reads <- c("A"=1000,
+           "B"=2000,
+           "C"=20,
+           "D"=5000)
+reads.to.table <- function(read,reads){
+    data.table("Total Reads"=read,
+               "Gene A"=reads["A"],
+               "Gene B"=reads["B"],
+               "Gene C"=reads["C"],
+               "Gene D"=reads["D"])
+}
+rr.table <- list(reads.to.table("Orig. FPKM",reads))
+i <- 0
+for (read in c(500,500,100,100,5,5,5)){
+    i <- i+1
+    reads <- reduced.reads(reads,read)
+    rr.table[[paste0("read.",read,'.',i)]] <- 
+        reads.to.table(read,reads)
+}
+print(xtable(rbindlist(rr.table),
+             digits=0
+             ),
+      include.rownames=FALSE)
+@
 

### 2017-01-03 exosome_reads/read_biaser.pl
use gzip instead of IO::Uncompress

@@ -161,7 +161,7 @@ for my $fn (@files) {
     if (not defined $fn) {
         $fh = \*STDIN;
     } else {
-        $fh = IO::Uncompress::AnyUncompress->new($fn,MultiStream => 1) or
+        open($fh,'-|','gzip','-dc',$fn) or
             die "Unable to open $fn for reading: $!";
     }
     my $read_pos = 9999;
@@ -236,7 +236,7 @@ for my $fn (@files) {
     if (not defined $fn) {
         $fh = \*STDIN;
     } else {
-        $fh = IO::Uncompress::AnyUncompress->new($fn,MultiStream => 1) or
+        open($fh,'-|','gzip','-dc',$fn) or
             die "Unable to open $fn for reading: $!";
     }
     my $read_pos = 999;

### 2017-01-03 manuscript/manuscript.Rnw
fix cache and data loading

@@
 \usepackage{pdflscape}
-\usepackage[backend=biber,natbib=true,hyperref=true,style=numeric-comp]{biblatex}
+\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
 \addbibresource{references.bib}
@@
 
-\newcommand{\EV}{\EV}
-
 \begin{document}
@@
 opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
-#opts_chunk$set(cache=TRUE, autodep=TRUE)
+opts_chunk$set(cache=TRUE, autodep=TRUE)
 options(device = function(file, width = 8, height = 7, ...) {
@@
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))>>=
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
 for (a in paste0("../tissue_specific_expression/",
@@
 
-\EV are small (40-1000nm) compartments that can be found in
-blood, saliva, and urine. \EV originate from the plasma membrane
+\Ac{EV} are small (40-1000nm) compartments that can be found in
+blood, saliva, and urine. \Ac{EV} originate from the plasma membrane
 or multivesicular endosome of many cell types and contain proteins,
-lipids, mRNA, miRNA, and many other non-coding RNAs. Because \EV
+lipids, mRNA, miRNA, and many other non-coding RNAs. Because \ac{EV}
 contain mRNA from the cell of origin and can be obtained without
@@
 \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq} and called using
-Tophat 2 v2.0.10
+Cufflinks v2.2.1
 \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}. The average
@@
 
-The population of RNAs found in \EV are distinct from the cytosolic
-pool, and are likely selectively sorted into \EV~%
+The population of RNAs found in \ac{EV} are distinct from the cytosolic
+pool, and are likely selectively sorted into \ac{EV}~%
 \cite{Villarroya-Beltri.ea2014:Sortingitoutregulation,Hoen.ea2012:DeepsequencingofRNA,Janas.ea2015:MechanismsofRNAloading,Balkom.ea2015:Quantitativeandqualitativeanalysis}.
@@
 UAAUCCCA~\cite{Batagov.ea2013:Exosomessecretedbyhuman} are enriched in
-\EV. Owing to the limited number of tissue-specific \ac{EV} RNA-seq
+\ac{EV}. Owing to the limited number of tissue-specific \ac{EV} RNA-seq
 results, we have biased the reads from existing tissue-specific
@@
 
-The \ac{SVM} was generated using caret
+The classification of samples into reads of placental origin vs
+non-placental origin or uterine origin vs non-uterine origin was
+performed using a \ac{SVM}. The \ac{SVM} was generated using caret
 \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
@@
 \begin{figure}
-<<tissue.specificity,shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+<<tissue.specificity.shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
 invisible(tissue.specific.genes[,type:="not specific"])
@@
 }
-grid.text(c("A","B","C","D"),
-          x=unit(c(0,0.5,0,0.5),"npc"),
-          y=unit(c(1,1,0.5,0.5),"npc"),
-          just=c("left","top"),
-          gp=gpar(fontsize=32)
-          )
+### grid.text(c("A","B","C","D"),
+###           x=unit(c(0,0.5,0,0.5),"npc"),
+###           y=unit(c(1,1,0.5,0.5),"npc"),
+###           just=c("left","top"),
+###           gp=gpar(fontsize=32)
+###           )
 

### 2017-01-04 exosome_reads/read_biaser.pl
use reservoir sampler to increase speed
 - handle paired reads properly too

@@ -24,6 +24,7 @@ read_biaser.pl [options]
    --reads number of reads to sample
    --output-prefix name of output file prefix
    --seed,--random-seed  Random seed
+   --paired Are reads paired?
    --samplings Number of times to sample reads
    --motif-output-abundance Abundance of motif in output (0.25)
    --debug, -d debugging level (Default 0)
@@ -78,6 +79,9 @@ read_biaser.pl
 use List::Util qw(sum);
 use IO::Uncompress::AnyUncompress;
 use vars qw($DEBUG);
+#use Math::Random;
+use List::Util qw(shuffle);
+use POSIX qw(floor);
 
 # In T cells, GGAG is required to target miRNA to T cell
 # motifs\cite{Villarroya-Beltri.ea2013:SumoylatedhnRNPA2B1controlssorting}.
@@ -92,6 +96,7 @@ my %options = (debug           => 0,
                seed            => rand,
                samplings       => 1,
                reads           => [],
+               paired          => 0,
                skip_reads      => 0,
                min_likelihood => 0.5,
                'output_prefix' => 'biased_reads',
@@ -103,6 +108,7 @@ my %options = (debug           => 0,
 GetOptions(\%options,
            'reads=i@',
            'seed|random-seed=s',
+           'paired!',
            'motif_output_abundance|motif-output-abundance=s',
            'motif=s@',
            'samplings=i',
@@ -117,154 +123,197 @@ $DEBUG = $options{debug};
 my @USAGE_ERRORS;
 if (not @{$options{reads}} or
     grep {$_ <= 1} @{$options{reads}}
-   ) {
+  ) {
     push @USAGE_ERRORS,"You must pass --reads and all --reads must be >= 1";
 }
 
 srand($options{seed});
-
-pod2usage(join("\n",@USAGE_ERRORS)) if @USAGE_ERRORS;
-
+# random_set_seed(random_seed_from_phrase($options{seed}));
+print STDERR "Seed: $options{seed}\n";
 
 my @files = @ARGV;
 if (not @ARGV) {
-    @files = undef;
+    push @USAGE_ERRORS,"You must provide at least one fasta file";
 }
+if ($options{paired} and (@files % 2 != 0)) {
+    push @USAGE_ERRORS,"If reads are paired, you must give an even number of fasta files";
+}
+
+pod2usage(join("\n",@USAGE_ERRORS)) if @USAGE_ERRORS;
 
[…]

### 2017-01-04 exosome_reads/read_biaser.pl
remove extra data structures to reduce memory

@@ -175,13 +175,13 @@ OUTER: for my $i ($options{paired}?(0..floor(@files/2)):(0..$#files)) {
         }
         my $read = read_a_read($fh);
         my $matched = $read->{read} =~ $motif_regex;
-        my @entry = $read;
+        my @entry = $read->{fastq};
         if ($options{paired}) {
             my $pair = read_a_read($fh2);
             if (not $matched) {
                 $matched = $pair->{read} =~ $motif_regex;
             }
-            push @entry,$pair;
+            push @entry,$pair->{fastq};
         }
         if ($matched) {
             $total_counts_matched++;
@@ -227,15 +227,13 @@ for my $i (0..$#output_files) {
             if (rand(1) <= $options{motif_output_abundance}) {
                 $entry = $matched_entry;
             }
-            print {$output_files[$i][$j][0]} $entry->[0]{fastq};
+            print {$output_files[$i][$j][0]} $entry->[0];
             if ($options{paired}) {
-                print {$output_files[$i][$j][1]} $entry->[1]{fastq};
+                print {$output_files[$i][$j][1]} $entry->[1];
             }
         }
     }
 }
-
-use Data::Printer;
 
 # Ideally we would use a weighted random sampler, but because we only
 # know the output abundance and not the input abundance, we cannot use

### 2017-01-05 exosome_reads/read_biaser.pl
we no longer need IO::Uncompress; load perl

@@ -77,7 +77,6 @@ read_biaser.pl
 =cut
 
 use List::Util qw(sum);
-use IO::Uncompress::AnyUncompress;
 use vars qw($DEBUG);
 #use Math::Random;
 use List::Util qw(shuffle);

### 2017-01-05 exosome_reads/read_biaser.pl
use /usr/bin/env perl instead of /usr/bin/perl

@@ -1,4 +1,4 @@
-#!/usr/bin/perl
+#!/usr/bin/env perl
 # read_biaser.pl biases reads towards motifs
 # and is released under the terms of the GNU GPL version 3, or any
 # later version, at your option. See the file README and COPYING for

### 2017-01-11 exosome_reads/R/make_srx_dir.R
use a subset of the SRX files to do the biasing

@@ -1,10 +1,19 @@
-library(tools)
-library(data.table)
+library("tools")
+library("data.table")
+library("yaml")
+args <- c("../tissue_expression/chosen_tissues",
+          "../tissue_specific_expression/categorized_samples",
+          "chosen_tissues.yaml")
 args <- commandArgs(trailingOnly=TRUE)
 
 load(args[1])
+load(args[2])
 
-for (srx in chosen.samples[,SRX]) {
+chosen.tissues <- data.table(tissues=yaml.load_file(args[3])$tissues)
+
+
+for (srx in categorized.samples[Sample_Group %in% chosen.tissues[,tissues] &
+                                grepl("^SRX",SRX),][,list(SRX=SRX[1]),by=Sample_Group][,SRX]) {
     if (!dir.exists(srx)) {
         dir.create(srx)
     }

### 2017-01-11 exosome_reads/R/write_srx.R

@@ -1,0 +1,29 @@
+library("tools")
+library("data.table")
+library("yaml")
+args <- c("../tissue_expression/chosen_tissues",
+          "../tissue_specific_expression/categorized_samples",
+          "chosen_tissues.yaml")
+args <- commandArgs(trailingOnly=TRUE)
+
+load(args[1])
+load(args[2])
+
+chosen.tissues <- data.table(tissues=yaml.load_file(args[3])$tissues)
+
+args <- commandArgs(trailingOnly=TRUE)
+
+### load chosen_samples
+load(args[1])
+
+srxes <-
+    categorized.samples[Sample_Group %in% chosen.tissues[,tissues] &
+                        grepl("^SRX",SRX),][,list(SRX=SRX[1]),by=Sample_Group][,SRX]
+mk.output <-
+    paste0("SRX_FILES=",paste(collapse=" ",srxes),"\n")
+for(srx in srxes) {
+    mk.output <- paste0(mk.output,
+                        srx,"_SRR=",paste(collapse=" ",chosen.samples[SRX==srx,SRR]),"\n")
+}
+cat(mk.output,
+    file=args[length(args)])

### 2017-01-12 exosome_reads/read_biaser_bam.pl
add read biaser and srx.mk to the makefile

@@ -1,0 +1,287 @@
+#!/usr/bin/env perl
+# read_biaser.pl biases reads towards motifs
+# and is released under the terms of the GNU GPL version 3, or any
+# later version, at your option. See the file README and COPYING for
+# more information.
+# Copyright 2016 by Don Armstrong <don@donarmstrong.com>.
+
+
+use warnings;
+use strict;
+
+use Getopt::Long;
+use Pod::Usage;
+
+=head1 NAME
+
+read_biaser.pl - biases reads towards motifs by sampling reads without replacement
+
+=head1 SYNOPSIS
+
+read_biaser.pl [options]
+
+ Options:
+   --reads number of reads to sample
+   --output-prefix name of output file prefix
+   --seed,--random-seed  Random seed
+   --paired Are reads paired?
+   --samplings Number of times to sample reads
+   --motif-output-abundance Abundance of motif in output (0.25)
+   --debug, -d debugging level (Default 0)
+   --help, -h display this help
+   --man, -m display manual
+
+=head1 OPTIONS
+
+=over
+
+=item B<--reads>
+
+Number of reads to sample. If repeated, each read option will be
+output C<samplings> times. Required.
+
+=item B<--samplings>
+
+Number of samplings to perform; defaults to 1. [Can be used to sample
+multiple sets of reads from a single file without re-reading the
+file.]
+
+=item B<--output-prefix>
+
+Output prefix of file. Defaults to
+biased_reads_rC<reads>_sC<sample>.fastq where C<reads> is the number
+of reads, and C<sample> is the nth sampling.
+
+=item B<--seed,--random-seed>
+
+Random seed to use; defaults to rand()
+
+=item B<--debug, -d>
+
+Debug verbosity. (Default 0)
+
+=item B<--help, -h>
+
+Display brief usage information.
+
+=item B<--man, -m>
+
+Display this manual.
+
+=back
+
+=head1 EXAMPLES
+
+read_biaser.pl
+
+=cut
+
+use List::Util qw(sum);
+use vars qw($DEBUG);
+#use Math::Random;
+use List::Util qw(shuffle);
+use POSIX qw(floor);
+
+# In T cells, GGAG is required to target miRNA to T cell
+# motifs\cite{Villarroya-Beltri.ea2013:SumoylatedhnRNPA2B1controlssorting}.
+
+# ACCAGCCU, CAGUGAGC, and UAAUCCCA motifs account for about 25% of
[…]

### 2017-01-12 go_analyses/go_tissue_analysis.R
add start of go analyses

@@ -1,0 +1,67 @@
+library("data.table")
+library(doMC)
+ncore = multicore:::detectCores()
+registerDoMC(cores = ncore)
+library("topGO")
+library("org.Hs.eg.db")
+
+args <- c("sample_reads_wide","tissue_specific_markers","go_tissue_analyses")
+
+args <- commandArgs(trailingOnly=TRUE)
+
+load(args[1])
+load(args[2])
+
+
+
+do.go.analysis <- function(tissue,go.ontology) {
+
+    all.genes <- c.reads.wide[eval(parse(text=paste0("`",tissue,"`>1"))),list(tracking_id)]
+    setkey(all.genes,"tracking_id")
+    all.genes[,tissue.specific:=0]
+    all.genes[as.character(tissue.specific.genes[tissue==tissue ,tracking_id]),
+              tissue.specific:=1]
+    all.genes[,egid:=as.vector(unlist(sapply(mget(all.genes[,tracking_id],
+                   org.Hs.egENSEMBL2EG,ifnotfound=NA),function(x){x[1]})))]
+    all.genes <- all.genes[!is.na(egid),]
+
+    gene.list <- factor(all.genes[,tissue.specific])
+    names(gene.list) <- all.genes[,egid]
+    
+    godata <- new("topGOdata",
+                  ontology=go.ontology,
+                  allGenes=gene.list,
+                  annot=annFUN.org,
+                  mapping="org.Hs.eg.db")
+    test.stat <- new("classicCount", testStatistic = GOFisherTest, name = "Fisher test")
+    result.fisher <- getSigGroups(godata,test.stat)
+    res.table <- 
+        data.table(GenTable(godata,
+                            classic=result.fisher,
+                            orderBy="classic",
+                            topNodes=length(result.fisher@score)))
+    res.table[,classic:=as.numeric(gsub("< ","",classic))]
+    setnames(res.table,"classic","p value")
+    res.table[,FDR:=p.adjust(method="BH",`p value`)]
+    res.table[,Ontology:=go.ontology]
+    return(res.table)
+}
+ 
+results <- 
+    c(lapply("placenta",#tissue.specific.genes[!is.na(tissue),unique(tissue)],
+             do.go.analysis,
+             go.ontology="MF"),
+      lapply("placenta",#tissue.specific.genes[!is.na(tissue),unique(tissue)],
+             do.go.analysis,
+             go.ontology="BP")
+      )
+
+print(results)
+
[…]

### 2017-01-12 go_analyses/tissue_expression.R

@@ -1,0 +1,36 @@
+library("data.table")
+library("reshape2")
+library("entropy")
+
+args <- c("categorized_samples",
+          "combined_read_counts",
+          "sample_reads_wide"
+          )
+
+
+load(args[1])
+load(args[2])
+
+setkey(categorized.samples,"SRX")
+read.counts <- gene.counts
+
+grouping.by.sample.formula <-
+    as.formula(tracking_id~Sample_Group)
+grouping.formula <-
+    as.formula(tracking_id~.)
+setnames(read.counts,"srx","SRX")
+setkey(read.counts,"SRX")
+combined.reads <-
+    categorized.samples[,list(Sample_Group,SRX)][read.counts]
+.narm.mean <- function(x){
+    if(length(x)==0) {return(0)} else {return(mean(x,na.rm=TRUE))}
+}
+c.reads.wide <-
+    data.table(dcast(combined.reads,
+                     grouping.by.sample.formula,
+                     fun.aggregate=.narm.mean,
+                     value.var="FPKM"
+                     ))
+
+save(c.reads.wide,
+     file=args[length(args)])

### 2017-01-13 exosome_reads/read_biaser_bam.pl
properly sort output bam file

@@ -229,7 +229,7 @@ sub output_sampled_reads {
     open($fh,'|-','samtools','view','-b','-o',$fn,'-') or
         die "Unable to open samtools view -b -o $fn - for writing: $!";
     print {$fh} $header;
-    for my $entry (sort { $a->[2] cmp $b->[2] || $a->[1] <=> $b->[2] }
+    for my $entry (sort { $a->[2] cmp $b->[2] || $a->[3] <=> $b->[3] }
                    @{$outputs}
                   ) {
         print {$fh} join("\t",@{$entry});

### 2017-02-01 manuscript/manuscript.Rnw
add the headers for the new version of the manuscript
do tissue specific genes plot

@@
 \setmonofont{FreeMono}
+\setmathrm{FreeSerif}
+\setmathsf{FreeSans}
+\setmathtt{FreeSerif}
 \usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
@@
 \usepackage{setspace}
-\usepackage{txfonts}
-\usepackage{fancyref}
+\usepackage{amsmath}
 \usepackage{acro}
@@
 \usepackage{xspace}
+\usepackage[markifdraft,raisemark=0.01\paperheight,draft]{gitinfo2}
 % this package will stretch the font slightly if necessary.
@@
 \widowpenalty = 10000
+\setlength{\emergencystretch}{20em}
+% make the afill smaller
+\renewcommand\Affilfont{\itshape\small}
+\renewcommand\cite[1]{\autocite{#1}}
 \pagestyle{fancy}
+\ExplSyntaxOn
+\NewAcroCommand \gene {
+  \group_begin:
+    \keys_set:nn {acro} {short-format = \itshape}
+    \acro_use:n{#1}
+  \group_end:
+}
+\NewAcroCommand \protein { \acro_use:n {#1} }
+\ExplSyntaxOff
 
@@
 <<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
-opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
+opts_chunk$set(dev="cairo_pdf",out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
 opts_chunk$set(cache=TRUE, autodep=TRUE)
@@
 library("grid")
+library("scales")
 
@@
 setkey(tissue.specific.genes,"gene")
+
+tissue.specific.genes[,type:="not specific"]
+tissue.specific.genes[,housekeeping:="not housekeeping"]
+tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"]
+tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
+                      possibly.specific==FALSE,housekeeping:="housekeeping"]
+tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
+                      possibly.specific==TRUE,housekeeping:="housekeeping and specific"]
+
+tissue.specific.genes[,type:=
+                          factor(type,
+                                 levels=c("not specific","tissue specific",
+                                     "housekeeping",
+                                     "housekeeping and specific"
+                                          ))]
 @
@@
 \begin{figure}
-<<tissue.specificity.shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
-invisible(tissue.specific.genes[,type:="not specific"])
-invisible(tissue.specific.genes[,housekeeping:="not housekeeping"])
-invisible(tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"])
-invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
-                                 possibly.specific==FALSE,housekeeping:="housekeeping"])
-invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
-                          possibly.specific==TRUE,housekeeping:="housekeeping and specific"])
-
-invisible(tissue.specific.genes[,type:=
-                          factor(type,
-                                 levels=c("not specific","tissue specific",
-                                     "housekeeping",
-                                     "housekeeping and specific"
-                                          ))])
+<<tissue.specificity.shape.figure,fig.width=4,fig.height=4,echo=FALSE>>=
 print(ggplot(tissue.specific.genes[gi.max>0],
-             aes(y=specificity,x=gi.max,colour=type))+
+             aes(y=specificity,x=gi.max))+
       ##geom_point()+
       geom_hex(bins=50)+
+      geom_line(aes(x=x,y=y),inherit.aes=FALSE,
+                color="orange",
+                data=data.frame(x=c(10,10,tissue.specific.genes[,max(gi.max)]),
+                                y=c(1,0.98,0.98)))+
       ## stat_summary_2d(fun=function(x){length(x)},bins=100)+
-      geom_point(aes(colour=housekeeping),data=tissue.specific.genes[housekeeping!="not housekeeping",])+
-      scale_x_log10()+
+      geom_point(aes(colour=gsub("And","\nAnd",
+                                 capwords(housekeeping))),
+                 size=2,
+                 data=tissue.specific.genes[housekeeping!="not housekeeping",])+
+      scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
+                    labels = trans_format("log10", math_format(10^.x)))+
+      scale_color_brewer("Housekeeping",palette="Set2")+
+      scale_fill_continuous("Count")+
       xlab("Max Expression (FPKM)")+
-      ylab("Tissue Specificity")+
-      theme(legend.position="top")
+      ylab("Tissue Specificity Index")+
+      theme(legend.position=c(1,0),
+            legend.justification=c(1,0),
+            legend.box.just="right"
             )
+    )
+
 @ 
-\caption{Shape of tissue specific genes}
+\caption{Tissue specificity index versus maximum expression in one
+  tissue. Genes above and to the right of the orange lines are tissue
+  specific. Genes which have been previously
+  identified~\cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
+  as housekeeping genes are generally not tissue specific, with the
+  exception of %
+  %
+  \Sexpr{tissue.specific.genes[housekeeping=="housekeeping and specific",gene]}.%
+  %
+}
 \label{fig:tissue_specific_shape}

### 2017-06-08 common_r/array_to_text.R
use fwrite in array_to_text

@@ -1,4 +1,5 @@
 ### $(R) $(ROPTS) -f array_to_text.R --args data_file_name array_name array_file.txt
+library("data.table")
 
 arguments <- commandArgs(trailingOnly=TRUE)
 data.file.name <- arguments[1]
@@ -6,7 +7,7 @@ data.array <- arguments[2]
 text.file.name <- arguments[3]
 
 load(data.file.name)
-eval(parse(text=c("write.table(",data.array,",",
+eval(parse(text=c("fwrite(",data.array,",",
              "file=text.file.name,",
              'sep="\\t")')
            )

### 2017-06-08 tissue_specific_expression/R/housekeeping_genes_superset.R
indicate housekeeping gene sources

@@ -14,10 +14,15 @@ housekeeping.genes.eisenberg <-
 setnames(housekeeping.genes.eisenberg,
          c("hgnc_symbol","ensembl_gene_id"),
          c("gene_short_name","tracking_id"))
+housekeeping.genes.eisenberg[,`housekeeping_source`:="eisenberg 2013"]
+housekeeping.genes[,`housekeeping_source`:="tissue specificity"]
 housekeeping.genes.superset <-
-    rbindlist(list(housekeeping.genes,housekeeping.genes.eisenberg),
+    rbindlist(list(housekeeping.genes.eisenberg,
+                   housekeeping.genes
+                   ),
               use.names=TRUE,fill=TRUE)
 housekeeping.genes.superset <-
-    housekeeping.genes.superset[!duplicated(tracking_id),list(gene_short_name,tracking_id)]
+    housekeeping.genes.superset[!duplicated(tracking_id),
+                                list(gene_short_name,tracking_id,housekeeping_source)]
 
 save(housekeeping.genes.superset,file=args[length(args)])

### 2017-06-08 manuscript/manuscript.Rnw
update the latex headers
add tissue specific marker barplot

@@
 \usepackage{url}
+\usepackage{siunitx}
 \usepackage{fancyhdr}
@@
 \usepackage{amsmath}
-\usepackage{acro}
+\usepackage[macros=false,single=true,single-form=long]{acro}
 \usepackage{array}
 \usepackage{dcolumn}
-\usepackage{adjustbox}
+\usepackage[export]{adjustbox}
 \usepackage{longtable}
@@
 \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
-\usepackage[x11names,svgnames]{xcolor}
+% \usepackage[x11names,svgnames]{xcolor}
 \usepackage{texshade}
@@
 % this package will stretch the font slightly if necessary.
-\usepackage{microtype}
+%\usepackage{microtype}
 \usepackage{zref-xr,zref-user}
 \usepackage[capitalize]{cleveref}
-\newenvironment{narrow}[2]{%
-  \begin{list}{}{%
-      \setlength{\topsep}{0pt}%
-      \setlength{\leftmargin}{#1}%
-      \setlength{\rightmargin}{#2}%
-      \setlength{\listparindent}{\parindent}%
-      \setlength{\itemindent}{\parindent}%
-      \setlength{\parsep}{\parskip}}%
-  \item[]}{\end{list}}
-\def\newblock{\hskip}
-\newenvironment{paperquote}{%
-  \begin{quote}%
-     \it
-  }%
-  {\end{quote}}
 \renewcommand{\textfraction}{0.15}
@@
 \renewcommand{\floatpagefraction}{0.60}
-%\renewcommand{\baselinestretch}{1.8}
-\newenvironment{enumerate*}%
-  {\begin{enumerate}%
-    \setlength{\itemsep}{0pt}%
-    \setlength{\parskip}{0pt}}%
-  {\end{enumerate}}
-\newenvironment{itemize*}%
-  {\begin{itemize}%
-    \setlength{\itemsep}{0pt}%
-    \setlength{\parskip}{0pt}}%
-  {\end{itemize}}
 \oddsidemargin 0.0in 
@@
 
+\input{acronyms}
+
 \begin{document}
@@
 \date{\today}
-
-\input{acronyms}
 
@@
                    axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
+           )
+}
+@
+
+<<specific_markers_barplot,echo=FALSE,results="hide",warning=FALSE>>=
+specific.markers.barplot <- function(tissue.name,genes) {
+    setkey(interesting.gene.reads.wide,"gene_short_name")
+    if (missing(genes))
+        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
+    tissue.specific.gene.reads <- 
+        data.table(melt(interesting.gene.reads.wide[genes,],
+                        id=c("gene_short_name","tracking_id"),
+                        variable.name="tissue",
+                        value.name="FPKM"
+                        ))
+    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
+    return(ggplot(tissue.specific.gene.reads,
+                  aes(x=gene_short_name,
+                      y=FPKM+1))
+           + geom_violin()
+           + scale_x_discrete("Gene")
+           + scale_color_discrete("Tissue")
+           + scale_shape_discrete("Tissue")
+           + scale_y_log10()
+           + geom_jitter(data=tissue.specific.gene.reads[order(-FPKM),
+                                                        list(tissue=tissue[1:3],
+                                                             max_fpkm=pmax(FPKM)[1:3]),
+                                                        by=gene_short_name],
+                        aes(y=max_fpkm+1,
+                            x=gene_short_name,
+                            color=capwords(as.character(tissue)),
+                            shape=capwords(as.character(tissue))
+                            ),
+                        size=2
+                        )
+           + coord_flip()
            )

### 2017-06-08 manuscript/manuscript.Rnw
add discussion of parts of paper

@@
 
+In this paper we want to show that given a population of \ac{EV}
+
+\textbf{Can we determine the proportions and identities of
+  the tissue(s) of \ac{EV} origin?}
+
+\textbf{Can assign differential expression to a particular tissue of
+  origin in \ac{EV}?}
+
+\textbf{Housekeeping genes are not tissue specific}
+
+\textbf{Additional genes which are likely to be housekeeping but were
+  not identified by previous papers}
+
 \section{Introduction}
 
-\subsection{Housekeeping genes}
+\subsection{\ac{EV} background}
 
-\subsection{Cell free RNA and Exosomes in Diagnostics}
+\paragraph{What are \acsp{EV}}
 
@@
 transcriptome of tissues that would otherwise be inaccessible.
+
+\paragraph{Where are \acsp{EV} found?}
+
+\paragraph{How are \acsp{EV} formed and what are their contents}
+
+\subsection{\acsp{EV} will enable routine liquid biopsies}
+
+\paragraph{Existing work showing promise of \acsp{EV}}
+
+\paragraph{Existing methods require tumor/disease specific marker which is unique to that tissue}
+
+\subsection{Identifying tissues of origin}
+
+\paragraph{Existing protein surface markers}
+
+\DLA{Is there a database of tissue/cell specific markers?}
+
+\paragraph{Issues of single cell transcriptome heterogeneity}
+
+\paragraph{Tissue specificity index}
+
+\subsection{Assigning differential expression}
+
+\subsection{Housekeeping genes}
+
+\subsection{Cell free RNA and Exosomes in Diagnostics}
 
@@
 
+\subsubsection{Samples}
 \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
@@
 
+\subsubsection{Alignment and Calling}
 Reads were aligned to \textit{Homo sapiens} reference genome GRCh38
@@
 
-\subsection{Tissue Specificity}
+\subsubsection{Normalization}
+
+\subsection{Tissue Specific Markers}
+
+\subsubsection{Tissue Specificity Index}
 

### 2017-06-08 manuscript/manuscript.Rnw
include unique interesting tissues and order by level

@@
          apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
-                             function(x){order(-x)[2]+2})]
-
+                             function(x){order(-x)[1:2]+2})]
     interesting.tissues <-
-        c("gene_short_name",tissue.name,unique(interesting.tissues))
+        c(tissue.name,
+          interesting.tissues)
+    interesting.tissues <- 
+        c("gene_short_name",unique(interesting.tissues))
 
@@
     tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
+    if (!is.null(tissue.name)) {
         gene.levels <- 
             unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
+    } else {
+        gene.levels <- 
+            unique(as.character(tissue.specific.genes[tissue %in% interesting.tissues
+                                                      ][order(specificity),gene]))
+    }
     tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]

### 2017-06-08 manuscript/manuscript.Rnw
include plot of housekeeping tissue specific markers

@@
 
+\subsubsection{Genes which are housekeeping and specific}
+
+\DLA{Why are these two genes housekeeping and specific?}
+
+\begin{figure}
+<<housekeeping_tissue_specific_markers_figure,fig.width=4,fig.height=4,echo=FALSE>>=
+print(specific.markers.barplot(genes=c("MT-TM","C19orf28"))
+      + xlab("Housekeeping Tissue\nSpecific Genes")
+      + theme(legend.position="Top")
+      + ylab("Tissue"))
+@
+\caption{Genes which are tissue specific but are also housekeeping genes}
+\label{fig:housekeeping_tissue_specific_genes}
+\end{figure}
+
 \subsection{Uterus and Placenta Specific Genes}

### 2017-06-08 manuscript/manuscript.Rnw
use cref instead of fref
make figures 4x4

@@
 \begin{figure}
-<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+<<uterus.specific.markers.figure,fig.width=4,fig.height=4,echo=FALSE>>=
 print(specific.markers.plot("uterus")
@@
 did not properly align against miRNA or snRNA
-\fref{tab:missing_uterus_specific_genes}. The uterus calls also appear
+\cref{tab:missing_uterus_specific_genes}. The uterus calls also appear
 to be wrong or cooked; for example, SNORD5 has counts of 7000ish which
@@
 \begin{figure}
-<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
+<<placenta.specific.markers.figure,fig.width=4,fig.height=4,echo=FALSE>>=
 print(specific.markers.plot("placenta",
@@
 \begin{figure}
-<<placenta.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
+<<placenta.roc.curve,fig.width=4,fig.height=4,echo=FALSE>>=
 layout(matrix(c(1,3,2,4),2,2))

### 2017-06-08 manuscript/manuscript.Rnw
show reduced read probability

@@
     }
-    inc.prob <- inclusionprobabilities(fpkms[non.zero],1)
+    inc.prob <- sampling::inclusionprobabilities(fpkms[non.zero],1)
     ret.val[non.zero] <-
-        apply(sapply(1:(num.reads/1),function(y){UPrandomsystematic(inc.prob,0)}),1,sum)
+        apply(sapply(1:(num.reads/1),function(y){sampling::UPrandomsystematic(inc.prob,0)}),1,sum)
     names(ret.val) <- names(fpkms)

### 2017-06-08 manuscript/manuscript.Rnw
switch to CairoPDF

@@
 <<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
-opts_chunk$set(dev="cairo_pdf",out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
+opts_chunk$set(dev="CairoPDF",out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
 opts_chunk$set(cache=TRUE, autodep=TRUE)
 options(device = function(file, width = 8, height = 7, ...) {
-  cairo_pdf(tempfile(), width = width, height = height, ...)
+  CairoPDF(tempfile(), width = width, height = height, ...)
 })

### 2017-06-08 manuscript/manuscript.Rnw
re-enable grid.text

@@
 }
-### grid.text(c("A","B","C","D"),
-###           x=unit(c(0,0.5,0,0.5),"npc"),
-###           y=unit(c(1,1,0.5,0.5),"npc"),
-###           just=c("left","top"),
-###           gp=gpar(fontsize=32)
-###           )
+grid.text(c("A","B","C","D"),
+          x=unit(c(0,0.5,0,0.5),"npc"),
+          y=unit(c(1,1,0.5,0.5),"npc"),
+          just=c("left","top"),
+          gp=gpar(fontsize=32)
+          )
 
@@
 
-\caption{\ac{ROC} curve of the placenta \ac{SVM} (A) and
-  uterus \ac{SVM} (B) . Both \ac{SVM} are highly effective at
+\caption{\ac{ROC} curve of the placenta \ac{SVM} (A,C) and
+  uterus \ac{SVM} (B,D) . Both \ac{SVM} are highly effective at
   distinguishing between samples of uterus origin and non-uterus
-  origin.
+  origin using the entirety of reads from a sample.
   %
-  \ac{ROC} curve of reduced reads in Placenta and
-  Uterus samples with the indicated number of total reads. Reads were
+  \ac{ROC} curve of reduced reads in Placenta (C) and
+  Uterus (D) samples with the indicated number of total reads. Reads were
   reduced by sampling reads for each sample weighted by the FPKM
   measured for each sample, and run using the same trained SVM
-  depicted in Figure \ref{fig:plac_ut_full_roc}. %
+  from the full runs.
   % 

### 2017-06-08 manuscript/manuscript.Rnw
add compact_tissue name and switch back to cairo_pdf

@@
 <<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
-opts_chunk$set(dev="CairoPDF",out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
-opts_chunk$set(cache=TRUE, autodep=TRUE)
-options(device = function(file, width = 8, height = 7, ...) {
-  CairoPDF(tempfile(), width = width, height = height, ...)
-})
+opts_chunk$set(out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
+opts_chunk$set(dev="cairo_pdf",cache=TRUE, autodep=TRUE)
 options(scipen = -1, digits = 2)
@@
                                           ))]
+@
+
+<<compact_tissue_name,echo=FALSE,results="hide">>=
+compact_tissue_name <- function(tissue) {
+    gsub("nhlf","human lung fibroblasts",
+         gsub("cultured cells","",
+         tissue))
+}
 @

### 2017-06-08 manuscript/manuscript.Rnw
use compact tissue names and white-black heatmaps

@@
                         ))
+    tissue.specific.gene.reads[,compact_tissue:=factor(compact_tissue_name(tissue),
+                                                       levels=compact_tissue_name(interesting.tissues))
+                               ]
     tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
@@
                   aes(y=gene_short_name,
-                          x=tissue))
+                      x=compact_tissue))
            + geom_tile(aes(fill=`log_2(FPKM)`))
@@
                                  low="white",
-                                 high="#002058",
+                                 high="black",
                                  guide=guide_colourbar(title=expression(log[2](FPKM))))
@@
                    plot.margin=unit(c(1,1,1,1),"mm"),
-                   axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
+                   axis.text.y = element_text(size=10,face="italic"),
+                   axis.text.x = element_text(size=9,angle = 30, hjust = 1, vjust=1))
            )
@@
 \begin{figure}
-<<placenta.specific.markers.figure,fig.width=4,fig.height=4,echo=FALSE>>=
+<<placenta.specific.markers.figure,fig.width=4,fig.height=6,echo=FALSE>>=
 print(specific.markers.plot("placenta",
@@
 \begin{figure}
-<<placenta.roc.curve,fig.width=4,fig.height=4,echo=FALSE>>=
+<<placenta.roc.curve,fig.width=8,fig.height=8,echo=FALSE>>=
 layout(matrix(c(1,3,2,4),2,2))

### 2017-06-09 manuscript/manuscript.Rnw
only plot FPKM > 1 in tissue specificity

@@
 <<tissue.specificity.shape.figure,fig.width=4,fig.height=4,echo=FALSE>>=
-print(ggplot(tissue.specific.genes[gi.max>0],
+print(ggplot(tissue.specific.genes[gi.max>=1],
              aes(y=specificity,x=gi.max))+

### 2017-06-09 manuscript/manuscript.Rnw
include all housekeeping and specific genes

@@
 <<housekeeping_tissue_specific_markers_figure,fig.width=4,fig.height=4,echo=FALSE>>=
-print(specific.markers.barplot(genes=c("MT-TM","C19orf28"))
+print(specific.markers.barplot(genes=tissue.specific.genes[housekeeping=="housekeeping and specific",gene])
       + xlab("Housekeeping Tissue\nSpecific Genes")

### 2017-06-09 tissue_specific_expression/R/additional_reads.R
add interesting gene reads after SVM/learning

@@ -1,11 +1,10 @@
 library("data.table")
 library("reshape2")
-library("entropy")
 
 args <- c("categorized_samples",
-          "combined_read_counts",
-          "gene_grouping_variables",
-          "interesting_gene_reads"
+          "interesting_gene_reads",
+          "additional_interesting_genes.txt",
+          "all_interesting_gene_reads"
           )
 
 args <- commandArgs(trailingOnly=TRUE)
@@ -15,8 +14,12 @@ args <- commandArgs(trailingOnly=TRUE)
 load(args[1])
 ## combined_read_counts
 load(args[2])
-## gene_grouping_variables
+## interesting_gene_reads
 load(args[3])
+## additional_interesting_genes.txt
+additional.genes <- fread(args[4])
+## gene_grouping_variables
+load(args[4])
 
 
 is.genes <- TRUE
@@ -32,20 +35,27 @@ setkey(categorized.samples,"SRX")
 ## share a common name
 if (is.genes) {
     read.counts <- gene.counts
+    interesting.reads <- interesting.gene.reads
     grouping.variables <- gene.grouping.variables
     rm(gene.counts)
 } else {
     read.counts <- isoform.counts
+    interesting.reads <- interesting.isoform.reads
     grouping.variables <- isoform.grouping.variables
     rm(isoform.counts)
 }
 
-setnames(read.counts,"srx","SRX")
-setkey(read.counts,"SRX")
+additional.tracking.ids <-
+    read.counts[!duplicated(tracking_id),
+                list(tracking_id,gene_short_name)][gene_short_name %in%
+                                                   additional.genes[[1]]][[1]]
 
-min.entropy <- 2
-min.tissue.specificity <- 0.985
-min.expression <- 10
+additional.tracking.ids <-
+    additional.tracking.ids[!(additional.tracking.ids %in%
+                              interesting.reads[,unique(tracking_id)])]
+setkey(read.counts,"tracking_id")
+read.counts <-
+    read.counts[additional.tracking.ids]
 
 ### combine sample name and srx into the gene counts file
 combined.reads <-
@@ -54,22 +64,17 @@ combined.reads <-
                                             gene_short_name,
                                             FPKM,
                                             SRX)]]
-interesting.groups <-
-    grouping.variables[((entropy >= min.entropy |
-                         tissue.specificity.index >= min.tissue.specificity) &
-                        max >= min.expression),
-                       tracking_id]
-
[…]

### 2017-06-09 tissue_specific_expression/R/classification_svm.R
avoid spaces in group assignment

@@ -40,7 +40,7 @@ for (group in c("uterus","placenta")) {
     gene.fpkm.wide[[group]] <-
         as.character(gene.fpkm.wide$Sample_Group)
     gene.fpkm.wide[[group]][!(gene.fpkm.wide[[group]] %in% group)] <-
-        paste0("not ",group)
+        paste0("not_",group)
     gene.fpkm.wide[[group]] <-
         as.factor(gene.fpkm.wide[[group]])
 }

### 2017-06-12 manuscript/manuscript.Rnw
abstract out tissue specific marker tables

@@
                                           ))]
+@
+
+<<genes.table.sanitize,echo=FALSE,results="hide">>=
+genes.table.sanitize <- function(str){
+    return(str)
+}
+link.gene.names <- function(gene){
+    return(paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",
+                  gene,"}{",
+                  "\\textit{",xtable.sanitize.text(gene),
+                  "}}"))
+}
+xtable.sanitize.text <- function(str){
+    result <- str
+    result <- gsub("\\\\", "SANITIZE.BACKSLASH", result)
+    result <- gsub("$", "\\$", result, fixed = TRUE)
+    result <- gsub(">", "$>$", result, fixed = TRUE)
+    result <- gsub("<", "$<$", result, fixed = TRUE)
+    result <- gsub("|", "$|$", result, fixed = TRUE)
+    result <- gsub("{", "\\{", result, fixed = TRUE)
+    result <- gsub("}", "\\}", result, fixed = TRUE)
+    result <- gsub("%", "\\%", result, fixed = TRUE)
+    result <- gsub("&", "\\&", result, fixed = TRUE)
+    result <- gsub("_", "\\_", result, fixed = TRUE)
+    result <- gsub("#", "\\#", result, fixed = TRUE)
+    result <- gsub("^", "\\verb|^|", result, fixed = TRUE)
+    result <- gsub("~", "\\~{}", result, fixed = TRUE)
+    result <- gsub("SANITIZE.BACKSLASH", "$\\backslash$",
+                   result, fixed = TRUE)
+    return(result)
+}
+source("/home/don/projects/xtable/pkg/R/print.xtable.R")
+
 @
@@
 
+<<tissue.specific.markers.table,results="hide",echo=FALSE>>=
+tissue.specific.markers.table <- function(tissue,legend,min.gi=0) {
+    this.tissue <- tissue
+    temp <-
+        tissue.specific.genes[tissue==this.tissue & gi.max >= min.gi,
+                              list(Gene=link.gene.names(gene),
+                                   "{Tissue specificity}"=specificity,
+                                   "{Mean Expression}"=gi.max)]
+    setorder(temp,-`{Mean Expression}`)
+    print(xtable(temp,
+                 align=c("r","c","S","S[table-format=1.0]"),
+                 caption=legend,
+                 label=paste0("tab:",tissue,"_specific_genes")),
+          sanitize.text.function=genes.table.sanitize,
+          include.rownames=FALSE,
+          booktabs=TRUE,
+          floating=TRUE,
+          )
+    rm(temp)
+}
+
+@
+
 \subsection{Uterus and Placenta Specific Genes}
@@
 <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
-temp <- tissue.specific.genes[tissue=="uterus",
-                              list(gene,tissue,specificity,gi.max)]
-setorder(temp,-gi.max)
-setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
-print(xtable(temp,
-             caption="Uterus specific genes",
-             label="tab:uterus_specific_genes"),
-      include.rownames=FALSE,
-      )
-rm(temp)
+tissue.specific.markers.table(tissue="uterus",
+                              legend="Uterus specific genes")
 @
@@
 <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
-temp <- tissue.specific.genes[tissue=="placenta" & gi.max > 100,
-                              list(gene,tissue,specificity,gi.max)]
-setorder(temp,-gi.max)
-setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
-print(xtable(temp,
-             caption="Placenta specific genes",label="tab:placenta_specific_genes"),
-      include.rownames=FALSE,
-      floating=FALSE,
-      tabular.environment="longtable"
+tissue.specific.markers.table(tissue="placenta",
+                              legend=
+                                  paste0("Placenta specific genes with expression ",
+                                         "$\\ge 100$ ordered by median expression in placenta."),
+                              min.gi=100
                               )
-rm(temp)
 @

### 2017-06-13 exosome_reads/R/make_srx_dir.R
SRR786752 is bad; ignore it

@@ -11,6 +11,8 @@ load(args[2])
 
 chosen.tissues <- data.table(tissues=yaml.load_file(args[3])$tissues)
 
+## SRR786752 claims to be paired, but only has one read
+chosen.samples[SRX=="SRX252261",SRR:=gsub(",SRR786752","",SRR)]
 
 for (srx in categorized.samples[Sample_Group %in% chosen.tissues[,tissues] &
                                 grepl("^SRX",SRX),][,list(SRX=SRX[1]),by=Sample_Group][,SRX]) {

### 2017-10-17 tissue_specific_expression/R/calculate_grouping_variables.R
calculate 2nd max and second max tissue

@@ -76,8 +76,9 @@ tissue.specificity.index <- function(expression){
                (length(expression)-1))
 }
 
-which.max.both <- function(expression,group) {
-    return(group[which.max(expression)])
+which.max.both <- function(expression,group,names=list("tissue_max"=1,"tissue_max_2nd"=2)) {
+    temp <- group[order(expression,decreasing=TRUE)]
+    lapply(names,function(x){temp[x]})
 }
 
 combined.reads.mean <-
@@ -89,19 +90,23 @@ rm(combined.reads)
 
 grouping.variables <-
     combined.reads.mean[Sample_Group!="universal human reference",
-                        list(gene_id=gene_id[1],
-                             gene_short_name=gene_short_name[1],
-                              entropy=calculate.entropy(mean_FPKM),
-                             entropy.log2=calculate.entropy(log2(mean_FPKM+1)),
-                             max=max(mean_FPKM,na.rm=TRUE),
-                             var=var(mean_FPKM,na.rm=TRUE),
-                             var.log2=var(log2(mean_FPKM+1),na.rm=TRUE),
-                             mean=mean(mean_FPKM,na.rm=TRUE),
-                             mean.log2=mean(log2(mean_FPKM+1),na.rm=TRUE),
-                             tissue_max=which.max.both(mean_FPKM,Sample_Group),
-                             tissue.specificity.index=tissue.specificity.index(mean_FPKM),
-                             tissue.specificity.index.log2=tissue.specificity.index(log2(mean_FPKM+1))
-                             ),
+                        c(list(gene_id=gene_id[1],
+                               gene_short_name=gene_short_name[1],
+                               entropy=calculate.entropy(mean_FPKM),
+                               entropy.log2=calculate.entropy(log2(mean_FPKM+1)),
+                               max=max(mean_FPKM,na.rm=TRUE),
+                               max_2nd=-sort.int(-mean_FPKM,partial=2)[2]
+                               var=var(mean_FPKM,na.rm=TRUE),
+                               var.log2=var(log2(mean_FPKM+1),na.rm=TRUE),
+                               mean=mean(mean_FPKM,na.rm=TRUE),
+                               mean.log2=mean(log2(mean_FPKM+1),na.rm=TRUE)),
+                          which.max.both(mean_FPKM,
+                                         Sample_Group,
+                                         names=list(tissue_max=1,
+                                                    tissue_max_2nd=2)),
[…]

### 2017-10-17 tissue_specific_expression/R/tissue_specific_expression.R
include 2nd max in table

@@ -55,10 +55,12 @@ specific.genes.isoforms <- function(grouping,min.specificity=0.98,min.max.expres
                                 tracking_id=tracking_id,
                                 tissue=tissue,
                                 gi.max=max,
+                                gi.max_2nd=max_2nd,
                                 specificity=tissue.specificity.index,
                                 log.specificity=tissue.specificity.index.log2,
                                 possibly.specific=possibly.specific,
                                 gi.which.max=tissue_max,
+                                gi.which.max_2nd=tissue_max_2nd,
                                 mean=mean,
                                 mean.log2=mean.log2,
                                 var=var,

### 2017-10-17 manuscript/manuscript.Rnw
update manuscript figures

@@
 \usepackage[noblocks,auth-sc]{authblk}
-\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
-%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
+% \usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
+\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
 % Floats at end for journals
@@
 opts_chunk$set(out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
-opts_chunk$set(dev="cairo_pdf",cache=TRUE, autodep=TRUE)
+opts_chunk$set(dev="CairoPDF",cache=TRUE, autodep=TRUE)
 options(scipen = -1, digits = 2)
@@
 
-<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
+<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","all_interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
 for (a in paste0("../tissue_specific_expression/",
                  c("tissue_specific_markers",
-                   "interesting_gene_reads_wide",
+                   "all_interesting_gene_reads_wide",
                    "combined_star_logs",
@@
            + theme(legend.position = "top",
-                   axis.ticks = element_blank(),
+                   axis.ticks.y = element_blank(),
+                   axis.ticks.x = element_blank(),
                    panel.grid=element_blank(),
@@
                    axis.text.y = element_text(size=10,face="italic"),
-                   axis.text.x = element_text(size=9,angle = 30, hjust = 1, vjust=1))
+                   axis.text.x = element_text(size=9,angle = 40,
+                                              hjust = 1, vjust=1,
+                                              margin=unit(c(0,0,0,0),"cm")))
            )
@@
                               list(Gene=link.gene.names(gene),
-                                   "{Tissue specificity}"=specificity,
-                                   "{Mean Expression}"=gi.max)]
-    setorder(temp,-`{Mean Expression}`)
-    print(xtable(temp,
-                 align=c("r","c","S","S[table-format=1.0]"),
+                                   specificity,
+                                   gi=gi.max)]
+    setorder(temp,-gi)
+    print(xtable(temp[,list(Gene,
+                            "{\\multicolumn{1}{p{1.5cm}}{\\centering Tissue \\\\ specificity }}"=
+                            specificity,
+                            "{\\multicolumn{1}{p{1.5cm}}{\\centering Mean \\\\ Expression }}"=
+                                gi)],
+                 align=c("r","c","S[table.format=1.2]","S[table-format=5.0,round-mode=places,round-precision=0]"),
                  caption=legend,
@@
           include.rownames=FALSE,
+          size="\\footnotesize",
           booktabs=TRUE,
@@
                               legend=
-                                  paste0("Placenta specific genes with expression ",
-                                         "$\\ge 100$ ordered by median expression in placenta."),
-                              min.gi=100
+                                  paste0("Placenta specific genes ($\\tau_g \\ge 0.98$) ",
+                                         "with expression ",
+                                         "$\\ge 300$ ordered by mean expression in placenta."),
+                              min.gi=300
                               )

### 2017-10-17 tissue_specific_expression/R/calculate_grouping_variables.R
add missing comma

@@ -95,7 +95,7 @@ grouping.variables <-
                                entropy=calculate.entropy(mean_FPKM),
                                entropy.log2=calculate.entropy(log2(mean_FPKM+1)),
                                max=max(mean_FPKM,na.rm=TRUE),
-                               max_2nd=-sort.int(-mean_FPKM,partial=2)[2]
+                               max_2nd=-sort.int(-mean_FPKM,partial=2)[2],
                                var=var(mean_FPKM,na.rm=TRUE),
                                var.log2=var(log2(mean_FPKM+1),na.rm=TRUE),
                                mean=mean(mean_FPKM,na.rm=TRUE),
